---
title: "Manuscript-figures"
author: "Kirk Gosik"
date: "6/26/2020"
output: html_document
---

# Manuscript Figures

## Setup

This part is just setup for reading in necessary packages and setting project directories.  It also contains a helper function for reading in zipped files for the genotype datasets.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(readxl)
library(magrittr)
library(qtl2)
library(tidyverse)
library(igraph)
library(Seurat)
library(cowplot)
library(VennDiagram)
library(RColorBrewer)
library(Gviz)
library(QTLRel)
library(snpStats)

project_path <- "/ahg/regevdata/projects/FASI_DOmice/"
if( length(dir(project_path)) == 0 ) project_path <- "/Volumes/ahg_regevdata/projects/FASI_DOmice/"

## assigning project paths
geno_path <- paste0(project_path, "genotype/")
pheno_path <- paste0(project_path, "phenotype/")
my_path <- paste0(project_path, "kirk/")
data_agg_path <- paste0(my_path, "data/aggregate-expression/")

fread_zip <- function(zipfile, skip = 9){
  
  # Create a name for the dir where we'll unzip
  zipdir <- tempfile()
  
  # Create the dir using that name
  dir.create(zipdir)
  
  # Unzip the file into the dir
  unzip(zipfile, exdir = zipdir)
  
  # Get a list of files in the dir
  files <- list.files(zipdir, full.names = TRUE)
  
  dat <- fread(files, skip = skip)
  gc(reset = TRUE)
  return(dat)
}
```

### helpers

More helpers for creating plots.  Uniform color schemes and plotting functions.

```{r my_themes}
# ## Seurat default colors
# default.colors <- c(scales::hue_pal()(length(x = levels(x = group.use))))
#       if (!is.null(x = names(x = group.colors))) {
#         cols <- unname(obj = group.colors[levels(x = group.use)])
#       } else {
#         cols <- group.colors[1:length(x = levels(x = group.use))] %||% default.colors
#         pal.max <- length(x = col.dups) + through
#           cols.extra <- hue_pal()(pal.max)[(through + 1):pal.max]
#           cols[col.dups] <- cols.extra
          
          

vega_20_scanpy = c("#1f77b4","#ff7f0e","#2ca02c","#d62728","#9467bd","#8c564b","#e377c2",  
                   "#bcbd22","#17becf","#aec7e8","#ffbb78","#98df8a","#ff9896","#c5b0d5",
                   "#c49c94","#f7b6d2","#dbdb8d","#9edae5","#ad494a","#8c6d31")  

zeileis_26 = c("#023fa5","#7d87b9","#bec1d4","#d6bcc0","#bb7784","#8e063b","#4a6fe3",
               "#8595e1","#b5bbe3","#e6afb9","#e07b91","#d33f6a","#11c638","#8dd593",
               "#c6dec7","#ead3c6","#f0b98d","#ef9708","#0fcfc0","#9cded6","#d5eae7",
               "#f3e1eb","#f6c4e1","#f79cd4","#7f7f7f","#c7c7c7","#1CE6FF","#336600")

godsnot_64 = c("#FFFF00","#1CE6FF","#FF34FF","#FF4A46","#008941","#006FA6","#A30059",
               "#FFDBE5","#7A4900","#0000A6","#63FFAC","#B79762","#004D43","#8FB0FF",
               "#997D87","#5A0007","#809693","#FEFFE6","#1B4400","#4FC601","#3B5DFF",
               "#4A3B53","#FF2F80","#61615A","#BA0900","#6B7900","#00C2A0","#FFAA92",
               "#FF90C9","#B903AA","#D16100","#DDEFFF","#000035","#7B4F4B","#A1C299",
               "#300018","#0AA6D8","#013349","#00846F","#372101","#FFB500","#C2FFED",
               "#A079BF","#CC0744","#C0B9B2","#C2FF99","#001E09","#00489C","#6F0062",
               "#0CBD66","#EEC3FF","#456D75","#B77B68","#7A87A1","#788D66","#885578",
               "#FAD09F","#FF8A9A","#D157A0","#BEC459","#456648","#0086ED","#886F4C",
               "#34362D","#B4A8BD","#00A6AA","#452C2C","#636375","#A3C8C9","#FF913F",
               "#938A81","#575329","#00FECF","#B05B6F","#8CD0FF","#3B9700","#04F757",
               "#C8A1A1","#1E6E00","#7900D7","#A77500","#6367A9","#A05837","#6B002C",
               "#772600","#D790FF","#9B9700","#549E79","#FFF69F","#201625","#72418F",
               "#BC23FF","#99ADC0","#3A2465","#922329","#5B4534","#FDE8DC","#404E55",
               "#0089A3","#CB7E98","#A4E804","#324E72","#6A3A4C")

my_theme <- theme(
    axis.line=element_blank(),
    axis.text.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks=element_blank(),
    # axis.title.x=element_blank(),
    # axis.title.y=element_blank(),
    # legend.position="none",
    panel.background=element_blank(),
    # panel.border=element_blank(),
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    plot.background=element_blank()
    )

my_theme_nolegened <- theme(
    axis.line=element_blank(),
    axis.text.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks=element_blank(),
    # axis.title.x=element_blank(),
    # axis.title.y=element_blank(),
    legend.position="none",
    panel.background=element_blank(),
    # panel.border=element_blank(),
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    plot.background=element_blank()
    )

my_blank_theme <- theme(
    axis.line=element_blank(),
    axis.text.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks=element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_blank(),
    legend.position="none",
    panel.background=element_blank(),
    # panel.border=element_blank(),
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    plot.background=element_blank()
    )


my_plot <- function( data, column, embedding = "tsne" ) { 
    
    ggplot(data, aes_string(x = paste0("X_", embedding, "1"), 
                            y = paste0("X_", embedding, "2"), 
                            color = column)) +
      geom_point(shape = 46) + 
      # my_theme_nolegened + 
      # scale_color_gradient(low="lightgrey", high="darkblue") +
      scale_color_gradient(low="darkblue", high="lightgreen") +
      labs(title = tools::toTitleCase(column))
  
}

## CHECK PLOT #######
# plot_quantile <- function( data, variable, group ) {
#   
#   AA <- quantile(as.numeric(data[[variable]][df[[group]] == 0]), 50:99 / 100, na.rm = TRUE)
#   Aa <- quantile(as.numeric(data[[variable]][df[[group]] == 1]), 50:99 / 100, na.rm = TRUE)
#   aa <- quantile(as.numeric(data[[variable]][df[[group]] == 2]), 50:99 / 100, na.rm = TRUE)
#   
#   out <- data.frame(quantile = rep(50:99/100, 3), 
#                     genotype = c(rep("AA", 50),  rep("Aa", 50), rep("aa", 50)), 
#                     variable = c(AA, Aa, aa))
#   
#   ggplot(out, aes(quantile, variable, color = genotype)) + 
#     geom_point() + 
#     geom_line() + 
#     labs(title = group)
#   
# }

my_quantile <- function(x) quantile(x, 50:100/100, na.rm=TRUE)

plot_quantile <- function(data, column, names_to) {
  data.frame(
    quantile = 50:100/100,
    AA = my_quantile(data[data[[names_to]]==0, column]), 
    Aa = my_quantile(data[data[[names_to]]==1, column]),
    aa = my_quantile(data[data[[names_to]]==2, column])
    ) %>% 
    pivot_longer(AA:aa, names_to = names_to) %>% 
    ggplot(., aes_string("quantile", "value", color = names_to)) + 
    geom_point() + 
    geom_line() +
    labs(title = column)  + 
    theme(    
      panel.background=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank()
      )
}





## TODO: Update ######
# # load(marker_file) ## load GM_snps
# marker_map <- readRDS(paste0(geno_path, "Regev_map_20171221.rds"))
# marker_map <- readRDS(marker_map_file)

# tf_lods <- tf_lods %>%
#   left_join({
#     data.frame(xpos) %>% rownames_to_column(var = "X1")
#   })

## NEEDS WORK ########
plot_qtl <- function( qtl_output, marker_map, varcolumn ) {
  
  df <- as.data.frame(qtl_output) %>%
    left_join({
      data.frame(xpos) %>% tibble::rownames_to_column(var = "marker")
    })
  
  xpos <- qtl2:::map_to_xpos(marker_map, gap = 25)
  xchrbound <- qtl2:::map_to_boundaries(marker_map, gap = 25)
  chr_breaks <- apply(xchrbound, 2, median)
  
  rectangles <- data.frame(
    xmin = apply(xchrbound, 2, min),
    xmax = apply(xchrbound, 2, max),
    ymin = 0,
    ymax = max(qtl_output[[varcolumn]])
    )[seq(2,20,2), ]
  
  gwas_plot <- ggplot() + 
    geom_rect(data = rectangles, 
              mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
              fill='gray80', alpha=0.8) +
    geom_point(data = qtl_output, 
               mapping = aes_string(x = "xpos", y = varcolumn), shape = 46) +
    geom_hline(yintercept = 11, color = "red", linetype = "dashed") + 
    # geom_label(data = markers_df, 
    #            aes(x = xpos, y = plot_lods, label = marker), 
    #            size = 2) +
    # geom_label_repel(data = snp_lables, 
    #                  aes(label=as.factor(marker), alpha = 0.7), 
    #                  size = 2, force = 1.3)
    scale_x_continuous(breaks = chr_breaks, label = c(as.character(1:19), "X")) + 
    labs(x = "Chromosome of SNP", 
         y = "P-value(-log10 scale)", 
         title = varcolumn) +
    theme(
      axis.line=element_blank(),
      axis.ticks=element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
      )
  
  gwas_plot
  
}



gwas_plot <- function( gwas_data, marker_map, plot_title ) {
  
  xpos <- qtl2:::map_to_xpos(marker_map, gap = 25)
  xchrbound <- qtl2:::map_to_boundaries(marker_map, gap = 25)
  chr_breaks <- apply(xchrbound, 2, median)
  
  rectangles <- data.frame(
    xmin = apply(xchrbound, 2, min),
    xmax = apply(xchrbound, 2, max),
    ymin = 0,
    ymax = max(gwas_data$lods)
    )[seq(2,20,2), ]
  
  ggplot() + 
    geom_rect(data = rectangles, 
              mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
              fill = 'gray80', alpha = 0.8) +
    geom_point(data = gwas_data, 
               mapping = aes(x = xpos, y = lods), shape = 46) +
    geom_hline(yintercept = 11, color = "red", linetype = "dashed") + 
    # geom_label(data = markers_df, 
    #            aes(x = xpos, y = plot_lods, label = marker), 
    #            size = 2) +
    # geom_label_repel(data = snp_lables, 
    #                  aes(label=as.factor(marker), alpha = 0.7), 
    #                  size = 2, force = 1.3)
    scale_x_continuous(breaks = chr_breaks, label = c(as.character(1:19), "X")) + 
    labs(x = "Chromosome of SNP", 
         y = "P-value(-log10 scale)", 
         title = plot_title) +
    theme(
      axis.line=element_blank(),
      axis.ticks=element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
      )
  
}

```



## Load Data

This chunk loads in the bulk of the data used to create the figures.  It can't all be read in at the same time and would need to be broken up for each of the plots they are needed for.  

```{r load_data}
obj <- readRDS(paste0(my_path, "data/seurat/allchannels.rds"))

obs <- fread(paste0(my_path, "data/scanpy/allchannels/obs.csv"), data.table = FALSE)
vars <- fread(paste0(my_path, "data/scanpy/allchannels/vars.csv"), data.table = FALSE)
obsm <- fread(paste0(my_path, "data/scanpy/allchannels/obsm.csv"), data.table = FALSE)
obsm <- data.frame(index = obs$index, obsm)
## rm(obs)

## meta <- fread(paste0(my_path, "data/MetaDataAllChannels.csv"), stringsAsFactors=FALSE)


## read SNP consequence dataset
cat("Reading SNP Consequence file ... \n")
conseq <- fread(paste0(my_path, "data/reference/GM_SNPS_Consequence.csv"), 
                data.table = FALSE)

ensembl <- fread(paste0(my_path, "data/reference/ensembl.Mus_musculus.GRCm38.93.csv")) %>%
  filter(chr %in% c(as.character(1:19), "X")) %>%
  left_join({
    fread(paste0(my_path, "data/reference/GeneID.csv"), 
                 col.names = c("symbol", "gene"), data.table = FALSE)
  })


## Read marker map
marker_map <- readRDS(paste0(geno_path, "Regev_map_20171221.rds"))


## ALREADY RUN ###################
# ## Promoter-like (PLS) 
# ## Proximal enhancer-like (pELS)
# ## Distal enhancer-like (dELS)
# ## CTCF-only
# ## DNase-H3K4me3
# encode <- read_delim(
#   file = paste0(my_path, "data/mm10-ccREs.bed"),
#   delim = "\t", 
#   col_names = c("chromosome", "start", "end", "id1", "id2", "type")) %>%
#   mutate(chr = str_remove(chromosome, "chr"))
# 
# encode_filtered <- lapply(1 : NROW(conseq), function(i) {
#   cat("index:",i)
#   encode %>%
#     filter(chr == conseq$chr[i],
#            start <= conseq$pos[i],
#            end >= conseq$pos[i]) %>%
#     mutate(pos = conseq$pos[i])
#   
#   }) %>%
#   do.call(rbind, .)
# 
# encode_conseq <- conseq %>%
#   left_join(encode_filtered, by = c("chr", "pos"))
# 
# write_csv(encode_conseq, "GM_SNPS_Consequence_cCRE.csv")
ccre <- fread(paste0(my_path, "data/GM_SNPS_Consequence_cCRE.csv"))
ccre[, recomb_rate := round(cM / (pos / 1e6), 4)]



# https://www.bioconductor.org/packages/release/bioc/vignettes/snpStats/inst/doc/ld-vignette.pdf
## Read SNP data
longfile <- paste0(geno_path, "20180706_batch4_G30_foreQTLandImmune/Broad_Inst_Xu_MURGIGV01_20180706/Broad_Inst_Xu_MURGIGV01_20180706_FinalReport.zip")

final <- fread_zip(longfile)

## getting mice ids to use later
mice_ids <- unique(final$`Sample ID`)

## intial QC Filter
final <- final[`GC Score` > 0.5, ]
## assigning genotype with Reference Allele A
final[, genotype := as.numeric(`Allele1 - AB` == "A") + as.numeric(`Allele2 - AB` == "A")]

## assigning NA genotype
# final[`Allele1 - AB` == "-", genotype := NA]

## making genotype dataframe
cat("Making genotype dataframe ... \n")
geno <- final[, .SD,.SDcols = c("SNP Name", "Sample ID", "genotype")] %>%
  # tidyr::spread( `SNP Name`, genotype ) %>%
  tidyr::pivot_wider(names_from = `SNP Name`, 
                     values_from = genotype, 
                     values_fill = list(genotype = 0)) %>%
  dplyr::rename( SampleID = `Sample ID` ) %>%
  as.data.frame()

rownames(geno) <- geno[,1]

## https://bioconductor.org/packages/release/bioc/vignettes/snpStats/inst/doc/data-input-vignette.pdf
# gdata <- read.long(allelesfile,
#                    fields=c(snp=1, sample=2, allele.A=3, allele.B=4, confidence=5),
#                    threshold=0.95)
# gdata
# Create a name for the dir where we'll unzip
zipdir <- tempfile()
# Create the dir using that name
dir.create(zipdir)
# Unzip the file into the dir
unzip(longfile, exdir = zipdir)

writeLines(text = readLines(list.files(zipdir, full.names = TRUE))[-c(1:10)],
           con = paste0(geno_path, "20180706-final-report.txt"))
# cat(readLines(paste0(geno_path, "20180706-final-report.txt")))

gdata <- read.long(paste0(geno_path, "20180706-final-report.txt"),
                   fields = c(snp=1, sample=2, allele.A=3, allele.B=4))
```


### Creating Plotting Data

Combining a lot of the loaded data above and parsing down to the necessary features for the majority of the plots.  Saving that into a dataframe that can be quickly accessed when needed. 

```{r create_plot_data}
## ALREADY RUN #####################
# keep_cols <- c("index", "Channel", "louvain_labels", "nUMI", "percent_mito", "LogProb",
#                "PValue", "Limited", "FDR", "doublet", "S.Score", "G2M.Score", "Phase",
#                "BIOCARTA_GATA3_PATHWAY", "BIOCARTA_HIF_PATHWAY",
#                "BIOCARTA_TCAPOPTOSIS_PATHWAY", "BIOCARTA_P53HYPOXIA_PATHWAY",
#                "BIOCARTA_IL17_PATHWAY", "BIOCARTA_IL3_PATHWAY", "BIOCARTA_IL5_PATHWAY",
#                "BIOCARTA_IL6_PATHWAY", "BIOCARTA_IL22BP_PATHWAY", "BIOCARTA_IL7_PATHWAY",
#                "BIOCARTA_TCRA_PATHWAY", "BIOCARTA_NO2IL12_PATHWAY",
#                "BIOCARTA_NKCELLS_PATHWAY", "BIOCARTA_TH1TH2_PATHWAY",
#                "BIOCARTA_TNFR2_PATHWAY", "PRB.DBL", "BestCall", "scrublet", "Ambient",
#                "called_cell_types", "DTT_used", "dissociation_genes",
#                "louvain_8_subclusters", "called_cell_types_new",
#                paste0("topic", 0:19), "filtered_set")
# 
# mhc2_genes <- grep("^H2",obj@assays$RNA@data@Dimnames[[1]], value = TRUE)
# ccr_ccl_genes <- grep("Ccr|Cxc|Ccl",obj@assays$RNA@data@Dimnames[[1]], value = TRUE)
# 
# plot_df <- data.frame(index = obj@assays$RNA@data@Dimnames[[2]],
#                  Ntm = obj@assays$RNA@data["Ntm", ],
#                  Il4 = obj@assays$RNA@data["Il4", ],
#                  Il5 = obj@assays$RNA@data["Il5", ],
#                  Il6 = obj@assays$RNA@data['Il6', ],
#                  Il13 = obj@assays$RNA@data["Il13", ],
#                  Il22 = obj@assays$RNA@data["Il22", ],
#                  Il25 = obj@assays$RNA@data["Il25", ],
#                  Il33 = obj@assays$RNA@data["Il33", ],
#                  Cd4 = obj@assays$RNA@data["Cd4", ],
#                  Cd6 = obj@assays$RNA@data["Cd6", ],
#                  Tpm2 = obj@assays$RNA@data["Tpm2", ],
#                  Gda = obj@assays$RNA@data["Gda", ],
#                  Stim2 = obj@assays$RNA@data["Stim2", ],
#                  Il17a = obj@assays$RNA@data["Il17a", ],
#                  Il18r1 = obj@assays$RNA@data["Il18r1", ],
#                  Il18rap = obj@assays$RNA@data["Il18rap", ],
#                  Zfp36 = obj@assays$RNA@data["Zfp36", ],
#                  Gata3 = obj@assays$RNA@data["Gata3", ],  ## topic11
#                  Calca = obj@assays$RNA@data["Calca", ], ## topic11
#                  Rora = obj@assays$RNA@data["Rora", ],
#                  Tnf = obj@assays$RNA@data["Tnf", ],
#                  Ifng = obj@assays$RNA@data["Ifng", ],
#                  Klrg1 = obj@assays$RNA@data["Klrg1", ],
#                  Nk1.1 = obj@assays$RNA@data["Klrb1c", ], ## topic0
#                  Ptpn1 = obj@assays$RNA@data["Ptpn1", ], ## topic0
#                  Arhgdib = obj@assays$RNA@data["Arhgdib", ],
#                  Cacna1e = obj@assays$RNA@data["Cacna1e", ],
#                  Rbpj = obj@assays$RNA@data["Rbpj", ],
#                  Nlrp12 = obj@assays$RNA@data["Nlrp12", ],
#                  Slc39a10 = obj@assays$RNA@data["Slc39a10", ],  ## topic8
#                  Slc39a11 = obj@assays$RNA@data["Slc39a11", ],  ## topic8
#                  Sox9 = obj@assays$RNA@data["Sox9", ],  ## topic8
#                  Sstr2 = obj@assays$RNA@data["Sstr2", ], ## topic8
#                  # Sstr4 = obj@assays$RNA@data["Sstr4", ],
#                  Thbd = obj@assays$RNA@data["Thbd", ],
#                  Cd93 = obj@assays$RNA@data["Cd93", ],
#                  Aftph = obj@assays$RNA@data["Aftph", ],  ## topic 4 gene?
#                  Socs2 = obj@assays$RNA@data["Socs2", ],
#                  Plxna4 = obj@assays$RNA@data["Plxna4", ],
#                  Cers6 = obj@assays$RNA@data["Cers6", ],
#                  Cradd= obj@assays$RNA@data["Cradd", ],
#                  Ncr1 = obj@assays$RNA@data["Ncr1", ],
#                  Bcl6 = obj@assays$RNA@data["Bcl6", ],
#                  Rorc = obj@assays$RNA@data["Rorc", ],
#                  Tbet = obj@assays$RNA@data["Tbx21", ],
#                  Ox40l = obj@assays$RNA@data["Tnfsf4", ], # MGI:104511
#                  Id2 = obj@assays$RNA@data["Id2", ],
#                  Nfil3 = obj@assays$RNA@data["Nfil3", ],
#                  Plzf = obj@assays$RNA@data["Zbtb16", ],   ## Zbtb16
#                  Tcf1 = obj@assays$RNA@data["Tcf7", ], ## Tcf7
#                  Tox = obj@assays$RNA@data["Tox", ],
#                  Bcl11b = obj@assays$RNA@data["Bcl11b", ],
#                  Eomes = obj@assays$RNA@data["Eomes", ],  ## tf gene
#                  Crth2 = obj@assays$RNA@data["Ptgdr2", ],  ## tf gene
#                  Areg = obj@assays$RNA@data["Areg", ],  ## tf gene
#                  Ckit = obj@assays$RNA@data["Kit", ],  ## tf gene
#                  Cnn2 = obj@assays$RNA@data["Cnn2", ],  ## eQTL gene ILC1
#                  Hopx = obj@assays$RNA@data["Hopx", ],  ## cis-eQTL gene ILC1
#                  Ppp1r2 = obj@assays$RNA@data["Ppp1r2", ],  ## cis-eQTL gene ILC1
#                  Abcb1b = obj@assays$RNA@data["Abcb1b", ],  ## cis-eQTL gene ILC2
#                  Pdcd5 = obj@assays$RNA@data["Pdcd5", ],  ## cis-eQTL gene ILC2
#                  Sptssa = obj@assays$RNA@data["Sptssa", ],  ## cis-eQTL gene ILC2
#                  Hmha1 = obj@assays$RNA@data["Hmha1", ],  ## cis-eQTL gene ILC3
#                  Lgals3 = obj@assays$RNA@data["Lgals3", ],  ## cis-eQTL gene ILC3
#                  Necap2 = obj@assays$RNA@data["Necap2", ],  ## cis-eQTL gene LTi-like
#                  Tmem132c = obj@assays$RNA@data["Tmem132c", ],  ## ILC Proportions
#                  Lgr6 = obj@assays$RNA@data["Lgr6", ],  ## ILC Proportions
#                  Rrbp1 = obj@assays$RNA@data["Rrbp1", ],  ## ILC Proportions
#                  Il20ra = obj@assays$RNA@data["Il20ra", ],  ## ILC Proportions
#                  Fam21 = obj@assays$RNA@data["Fam21", ],   ## ILC Proportions
#                  Banf2 = obj@assays$RNA@data["Banf2", ],  ## ILC Proportions
#                  Pgm1 = obj@assays$RNA@data["Pgm1", ],  ## ILC Proportions
#                  Atp6ap1l = obj@assays$RNA@data["Atp6ap1l", ],  ## IgE QTL genes
#                  Atg10 = obj@assays$RNA@data["Atg10", ],  ## IgE QTL genes
#                  Acot12 = obj@assays$RNA@data["Acot12", ],  ## IgE QTL genes
#                  Ssbp2 = obj@assays$RNA@data["Ssbp2", ],  ## IgE QTL genes
#                  Zcchc9 = obj@assays$RNA@data["Zcchc9", ],  ## IgE QTL genes
#                  Rasgrf2= obj@assays$RNA@data["Rasgrf2", ],  ## IgE QTL genes
#                  Myh9 = obj@assays$RNA@data["Myh9", ], ## ILC2 eGene IgE QTL
#                  Borcs8 = obj@assays$RNA@data["Borcs8", ], ## ILC2 eGene IgE QTL
#                  Fis1 = obj@assays$RNA@data["Fis1", ], ## ILC2 eGene IgE QTL
#                  Nfkb1 = obj@assays$RNA@data["Nfkb1", ], ## ILC2 eGene IgE QTL
#                  Ahr = obj@assays$RNA@data["Ahr", ], ## Ahr with Ccr6, IL-22, topic3?
#                  Vip = obj@assays$RNA@data["Vip", ], ## ILC2 signal
#                  stringsAsFactors = FALSE) %>%
#   ## left_join(meta[, .SD, .SDcols = keep_cols]) %>%
#   left_join(obs[, keep_cols]) %>%
#   left_join(obsm[,c("index", "X_tsne1", "X_tsne2", "X_umap1", "X_umap2",
#                     "X_diffmap1", "X_diffmap2", "X_diffmap_pca1", "X_diffmap_pca2",
#                     "X_diffmap_sym1", "X_diffmap_sym2", "X_net_fle1", "X_net_fle2",
#                     "X_net_fle_pred1", "X_net_fle_pred2")]) %>%
#   mutate(louvain_labels = factor(louvain_labels),
#          Channel = factor(Channel),
#          activated = ifelse(louvain_labels %in% 4:5, 2, 1),
#          log_umi = log(nUMI)) %>%
#   left_join(data.frame(index = obj@assays$RNA@data@Dimnames[[2]],   ## mhcII genes
#                        t(as.matrix(obj@assays$RNA@data[mhc2_genes,])))) %>%
#   left_join(data.frame(index = obj@assays$RNA@data@Dimnames[[2]],   ## ccr_ccl genes
#                        t(as.matrix(obj@assays$RNA@data[ccr_ccl_genes,]))))
# 
# write_csv(plot_df, paste0(my_path, "data/manuscript-plot-data.csv"))
## plot_df <- read_csv(paste0(my_path, "data/manuscript-plot-data.csv"))
plot_df <- fread(paste0(my_path, "data/manuscript-plot-data.csv"), data.table = FALSE)
```


## Munging Data

Scripts needed to create the data used. 
  - run-uger-eqtl.R
    - results/qtl-plot-lods-ILC1-mean.csv
    - results/qtl-plot-lods-ILC1-cv.csv
    - results/qtl-plot-lods-ILC2-mean.csv
    - results/qtl-plot-lods-ILC2-cv.csv
    - results/qtl-plot-lods-NCR1+ ILC3-mean.csv
    - results/qtl-plot-lods-NCR1+ ILC3-cv.csv
    - results/qtl-plot-lods-Lti ILC3-mean.csv
    - results/qtl-plot-lods-Lti ILC3-cv.csv

```{r ilc1_genes_eqtls}
## ILC1
ilc1_obj <- subset(x = obj, subset = called_cell_types == "ILC1" & filtered_set == 1)

## keep columns with at least 10% expression
ilc1_expressed <- names(which(Matrix::rowMeans(x = ilc1_obj@assays$RNA@data > 0) > 0.1))
rm(ilc1_obj)

## ILC1 eQTL Mean
ilc1_mean <- fread(paste0(my_path, "results/qtl-plot-lods-ILC1-mean.csv")) %>%
  dplyr::left_join({ 
    conseq %>% dplyr::select(marker, marker_pos = pos, ensembl_gene, conseq_clean) }) %>%
  dplyr::left_join({ 
    ensembl %>% dplyr::select(id = gene, gene_start = start, gene_end = end) }) %>%
  dplyr::mutate(cis_effect = as.numeric(marker_chr == gene_chr & 
                                   {marker_pos > (gene_start - 1e6) & 
                                       marker_pos < (gene_end + 1e6)})) %>%
  dplyr::mutate(value_adj = ifelse(cis_effect==1, value + 6, value), cell_type = "ILC1") %>%
  dplyr::filter(value_adj > 10)

## ILC1 eQTL CV
ilc1_cv <- fread(paste0(my_path, "results/qtl-plot-lods-ILC1-cv.csv")) %>%
  dplyr::left_join({ 
    conseq %>% dplyr::select(marker, marker_pos = pos, ensembl_gene, conseq_clean) }) %>%
  dplyr::left_join({ 
    ensembl %>% dplyr::select(id = gene, gene_start = start, gene_end = end) }) %>%
  dplyr::mutate(cis_effect = as.numeric(marker_chr == gene_chr & 
                                   {marker_pos > (gene_start - 1e6) & 
                                       marker_pos < (gene_end + 1e6)})) %>%
  dplyr::mutate(value_adj = ifelse(cis_effect==1, value + 6, value), cell_type = "ILC1") %>%
  dplyr::filter(value_adj > 10)
```


```{r ilc2_genes_eqtls}
## ILC2
ilc2_obj <- subset(x = obj, subset = called_cell_types == "ILC2" & filtered_set == 1)
  
## keep columns with at least 10% expression
ilc2_expressed <- names(which(Matrix::rowMeans(x = ilc2_obj@assays$RNA@data > 0) > 0.1))
rm(ilc2_obj)

## ILC2 eQTL Mean
ilc2_mean <- fread(paste0(my_path, "results/qtl-plot-lods-ILC2-mean.csv")) %>%
  dplyr::left_join({ 
    conseq %>% dplyr::select(marker, marker_pos = pos, ensembl_gene, conseq_clean) }) %>%
  dplyr::left_join({ 
    ensembl %>% dplyr::select(id = gene, gene_start = start, gene_end = end) }) %>%
  dplyr::mutate(cis_effect = as.numeric(marker_chr == gene_chr & 
                                   {marker_pos > (gene_start - 1e6) & 
                                       marker_pos < (gene_end + 1e6)})) %>%
  dplyr::mutate(value_adj = ifelse(cis_effect==1, value + 6, value), cell_type = "ILC2") %>%
  dplyr::filter(value_adj > 10)


## ILC2 eQTL CV
ilc2_cv <- fread(paste0(my_path, "results/qtl-plot-lods-ILC2-cv.csv")) %>%
  dplyr::left_join({ 
    conseq %>% dplyr::select(marker, marker_pos = pos, ensembl_gene, conseq_clean) }) %>%
  dplyr::left_join({ 
    ensembl %>% dplyr::select(id = gene, gene_start = start, gene_end = end) }) %>%
  dplyr::mutate(cis_effect = as.numeric(marker_chr == gene_chr & 
                                   {marker_pos > (gene_start - 1e6) & 
                                       marker_pos < (gene_end + 1e6)})) %>%
  dplyr::mutate(value_adj = ifelse(cis_effect==1, value + 6, value), cell_type = "ILC2") %>%
  dplyr::filter(value_adj > 10)
```


```{r ilc3_genes_eqtls}
## NCR1+ ILC3
ilc3_obj <- subset(x = obj, subset = called_cell_types == "NCR1+ ILC3" & filtered_set == 1)
  
## keep columns with at least 10% expression
ilc3_expressed <- names(which(Matrix::rowMeans(x = ilc3_obj@assays$RNA@data > 0) > 0.1))
rm(ilc3_obj)



## ILC3 eQTL Mean
ilc3_mean <- fread(paste0(my_path, "results/qtl-plot-lods-NCR1+ ILC3-mean.csv")) %>%
  dplyr::left_join({ 
    conseq %>% dplyr::select(marker, marker_pos = pos, ensembl_gene, conseq_clean) }) %>%
  dplyr::left_join({ 
    ensembl %>% dplyr::select(id = gene, gene_start = start, gene_end = end) }) %>%
  dplyr::mutate(cis_effect = as.numeric(marker_chr == gene_chr & 
                                   {marker_pos > (gene_start - 1e6) & 
                                       marker_pos < (gene_end + 1e6)})) %>%
  dplyr::mutate(value_adj = ifelse(cis_effect==1, value + 6, value), cell_type = "ILC3") %>%
  dplyr::filter(value_adj > 10)



## ILC3 eQTL CV
ilc3_cv <- fread(paste0(my_path, "results/qtl-plot-lods-NCR1+ ILC3-cv.csv")) %>%
  dplyr::left_join({ 
    conseq %>% dplyr::select(marker, marker_pos = pos, ensembl_gene, conseq_clean) }) %>%
  dplyr::left_join({ 
    ensembl %>% dplyr::select(id = gene, gene_start = start, gene_end = end) }) %>%
  dplyr::mutate(cis_effect = as.numeric(marker_chr == gene_chr & 
                                   {marker_pos > (gene_start - 1e6) & 
                                       marker_pos < (gene_end + 1e6)})) %>%
  dplyr::mutate(value_adj = ifelse(cis_effect==1, value + 6, value), cell_type = "ILC3") %>%
  dplyr::filter(value_adj > 10)
```



```{r lti_genes_eqtls}
## Lti ILC3
lti_obj <- subset(x = obj, subset = called_cell_types == 'Lti ILC3' & filtered_set == 1)
  
## keep columns with at least 10% expression
lti_expressed <- names(which(Matrix::rowMeans(x = lti_obj@assays$RNA@data > 0) > 0.1))
rm(lti_obj)


## LTi eQTL Mean
lti_mean <- fread(paste0(my_path, "results/qtl-plot-lods-Lti ILC3-mean.csv")) %>%
  dplyr::left_join({ 
    conseq %>% dplyr::select(marker, marker_pos = pos, ensembl_gene, conseq_clean) }) %>%
  dplyr::left_join({ 
    ensembl %>% dplyr::select(id = gene, gene_start = start, gene_end = end) }) %>%
  dplyr::mutate(cis_effect = as.numeric(marker_chr == gene_chr & 
                                   {marker_pos > (gene_start - 1e6) & 
                                       marker_pos < (gene_end + 1e6)})) %>%
  dplyr::mutate(value_adj = ifelse(cis_effect==1, value + 6, value), cell_type = "LTi") %>%
  dplyr::filter(value_adj > 10)


## LTi eQTL CV
lti_cv <- fread(paste0(my_path, "results/qtl-plot-lods-Lti ILC3-cv.csv")) %>%
  dplyr::left_join({ 
    conseq %>% dplyr::select(marker, marker_pos = pos, ensembl_gene, conseq_clean) }) %>%
  dplyr::left_join({ 
    ensembl %>% dplyr::select(id = gene, gene_start = start, gene_end = end) }) %>%
  dplyr::mutate(cis_effect = as.numeric(marker_chr == gene_chr & 
                                   {marker_pos > (gene_start - 1e6) & 
                                       marker_pos < (gene_end + 1e6)})) %>%
  dplyr::mutate(value_adj = ifelse(cis_effect==1, value + 6, value), cell_type = "LTi") %>%
  dplyr::filter(value_adj > 10)


rm(obj)
```


```{r}
ilc1_bygene <- group_by(gene, marker_chr)
```




### Make Gene Table

Make a csv file containing all the genes in the dataset.  Adding this to the scanpy output vars.csv.  Have a column that indicates whether that gene is 

  - expressed each specific cell type (used as phenotype in eQTL analysis)
  - Has SNPs in the Giga MUGA array
  - Was an eGene for each specific cell type
  - had a eQTL for each specific cell type
  - Was differentially expressed in each cell type / Marker gene identified for cell type
  - QTL loci genes for each topic
  - topic genes
  - antibodyQTL loci genes
  - cytokine QTL loci genes
  - cell proportion composition QTL loci genes
  
```{r gene_table, eval = FALSE}
## starting with all ensembl genes
gene_table <- ensembl

# ## identifyig which are in the 10x dataset
# gene_table$in_10x <- 0
# gene_table$in_10x[gene_table$symbol %in% obj@assays$RNA@data@Dimnames[[1]]] <- 1

## Giga MUGA Array Genes with SNPs ######
gene_table$in_gigamuga <- 0
gene_table$in_gigamuga[gene_table$gene %in% conseq$ensembl_gene] <- 1


## eQTL Genes #######

## ILC1 expressed at 10% of the cell type
gene_table$ilc1_expressed <- 0
gene_table$ilc1_expressed[gene_table$symbol %in% names(ilc1_expressed)] <- 1

## Mean #######
## ILC1 eGenes
gene_table$ilc1_egenes_mean <- 0
gene_table$ilc1_egenes_mean[gene_table$symbol %in% unique(ilc1$gene)] <- 1

## ILC1 eQTL Loci genes
gene_table$ilc1_eqtl_loci_genes_mean <- 0
gene_table$ilc1_eqtl_loci_genes_mean[gene_table$gene %in% unique(ilc1$ensembl_gene)] <- 1

## ILC1 eQTL cis-effect associations
gene_table$ilc1_ciseqtl_assoc_mean <- 0
gene_table$ilc1_ciseqtl_assoc_mean[gene_table$symbol %in% unique(ilc1$gene[ilc1$cis_effect==1])] <- 1

## CV #####
## ILC1 eGenes
gene_table$ilc1_egenes_cv <- 0
gene_table$ilc1_egenes_cv[gene_table$symbol %in% unique(ilc1$gene)] <- 1

## ILC1 eQTL Loci genes
gene_table$ilc1_eqtl_loci_genes_cv <- 0
gene_table$ilc1_eqtl_loci_genes_cv[gene_table$gene %in% unique(ilc1$ensembl_gene)] <- 1

## ILC1 eQTL cis-effect associations
gene_table$ilc1_ciseqtl_assoc_cv <- 0
gene_table$ilc1_ciseqtl_assoc_cv[gene_table$symbol %in% unique(ilc1$gene[ilc1$cis_effect==1])] <- 1


## ILC2 expressed at 10% of the cell type
gene_table$ilc2_expressed <- 0
gene_table$ilc2_expressed[gene_table$symbol %in% names(ilc2_expressed)] <- 1

## Mean #######
## ILC2 eGenes
gene_table$ilc2_egenes_mean <- 0
gene_table$ilc2_egenes_mean[gene_table$symbol %in% unique(ilc2$gene)] <- 1

## ILC2 eQTL Loci genes
gene_table$ilc2_eqtl_loci_genes_mean <- 0
gene_table$ilc2_eqtl_loci_genes_mean[gene_table$gene %in% unique(ilc2$ensembl_gene)] <- 1

## ILC2 eQTL cis-effect associations
gene_table$ilc2_ciseqtl_assoc_mean <- 0
gene_table$ilc2_ciseqtl_assoc_mean[gene_table$symbol %in% unique(ilc2$gene[ilc2$cis_effect==1])] <- 1

## CV #####
## ILC2 eGenes
gene_table$ilc2_egenes_cv <- 0
gene_table$ilc2_egenes_cv[gene_table$symbol %in% unique(ilc2$gene)] <- 1

## ILC2 eQTL Loci genes
gene_table$ilc2_eqtl_loci_genes_cv <- 0
gene_table$ilc2_eqtl_loci_genes_cv[gene_table$gene %in% unique(ilc2$ensembl_gene)] <- 1

## ILC2 eQTL cis-effect associations
gene_table$ilc2_ciseqtl_assoc_cv <- 0
gene_table$ilc2_ciseqtl_assoc_cv[gene_table$symbol %in% unique(ilc2$gene[ilc2$cis_effect==1])] <- 1


## ILC3 expressed at 10% of the cell type
gene_table$ilc3_expressed <- 0
gene_table$ilc3_expressed[gene_table$symbol %in% names(ilc3_expressed)] <- 1

## Mean #######
## ILC3 eGenes
gene_table$ilc3_egenes_mean <- 0
gene_table$ilc3_egenes_mean[gene_table$symbol %in% unique(ilc3$gene)] <- 1

## ILC3 eQTL Loci genes
gene_table$ilc3_eqtl_loci_genes_mean <- 0
gene_table$ilc3_eqtl_loci_genes_mean[gene_table$gene %in% unique(ilc3$ensembl_gene)] <- 1

## ILC3 eQTL cis-effect associations
gene_table$ilc3_ciseqtl_assoc_mean <- 0
gene_table$ilc3_ciseqtl_assoc_mean[gene_table$symbol %in% unique(ilc3$gene[ilc3$cis_effect==1])] <- 1

## CV #####
## ILC3 eGenes
gene_table$ilc3_egenes_cv <- 0
gene_table$ilc3_egenes_cv[gene_table$symbol %in% unique(ilc3$gene)] <- 1

## ILC3 eQTL Loci genes
gene_table$ilc3_eqtl_loci_genes_cv <- 0
gene_table$ilc3_eqtl_loci_genes_cv[gene_table$gene %in% unique(ilc3$ensembl_gene)] <- 1

## ILC3 eQTL cis-effect associations
gene_table$ilc3_ciseqtl_assoc_cv <- 0
gene_table$ilc3_ciseqtl_assoc_cv[gene_table$symbol %in% unique(ilc3$gene[ilc3$cis_effect==1])] <- 1



## LTi expressed at 10% of the cell type
gene_table$lti_expressed <- 0
gene_table$lti_expressed[gene_table$symbol %in% names(lti_expressed)] <- 1

## Mean #######
## LTi eGenes
gene_table$lti_egenes_mean <- 0
gene_table$lti_egenes_mean[gene_table$symbol %in% unique(lti$gene)] <- 1

## LTi eQTL Loci genes
gene_table$lti_eqtl_loci_genes_mean <- 0
gene_table$lti_eqtl_loci_genes_mean[gene_table$gene %in% unique(lti$ensembl_gene)] <- 1

## LTi eQTL cis-effect associations
gene_table$lti_ciseqtl_assoc_mean <- 0
gene_table$lti_ciseqtl_assoc_mean[gene_table$symbol %in% unique(lti$gene[lti$cis_effect==1])] <- 1

## CV #####
## LTi eGenes
gene_table$lti_egenes_cv <- 0
gene_table$lti_egenes_cv[gene_table$symbol %in% unique(lti$gene)] <- 1

## LTi eQTL Loci genes
gene_table$lti_eqtl_loci_genes_cv <- 0
gene_table$lti_eqtl_loci_genes_cv[gene_table$gene %in% unique(lti$ensembl_gene)] <- 1

## LTi eQTL cis-effect associations
gene_table$lti_ciseqtl_assoc_cv <- 0
gene_table$lti_ciseqtl_assoc_cv[gene_table$symbol %in% unique(lti$gene[lti$cis_effect==1])] <- 1 



## Topic Genes ######

topic_genes <- fread(paste0(my_path, "results/allchannels/gensim/outputs/allchannels-20-topics-top-genes.csv"))

topic0_genes <- topic_genes$Gene[topic_genes$Topic == 0]
vars$topic0_genes <- 0
vars$topic0_genes[vars$index %in% topic0_genes] <- 1

topic1_genes <- topic_genes$Gene[topic_genes$Topic == 1]
vars$topic1_genes <- 0
vars$topic1_genes[vars$index %in% topic1_genes] <- 1

topic2_genes <- topic_genes$Gene[topic_genes$Topic == 2]
vars$topic2_genes <- 0
vars$topic2_genes[vars$index %in% topic2_genes] <- 1

topic3_genes <- topic_genes$Gene[topic_genes$Topic == 3]
vars$topic3_genes <- 0
vars$topic3_genes[vars$index %in% topic3_genes] <- 1

topic4_genes <- topic_genes$Gene[topic_genes$Topic == 4]
vars$topic4_genes <- 0
vars$topic4_genes[vars$index %in% topic4_genes] <- 1

topic5_genes <- topic_genes$Gene[topic_genes$Topic == 5]
vars$topic5_genes <- 0
vars$topic5_genes[vars$index %in% topic5_genes] <- 1

topic6_genes <- topic_genes$Gene[topic_genes$Topic == 6]
vars$topic6_genes <- 0
vars$topic6_genes[vars$index %in% topic6_genes] <- 1

topic7_genes <- topic_genes$Gene[topic_genes$Topic == 7]
vars$topic7_genes <- 0
vars$topic7_genes[vars$index %in% topic7_genes] <- 1

topic8_genes <- topic_genes$Gene[topic_genes$Topic == 8]
vars$topic8_genes <- 0
vars$topic8_genes[vars$index %in% topic8_genes] <- 1

topic9_genes <- topic_genes$Gene[topic_genes$Topic == 9]
vars$topic9_genes <- 0
vars$topic9_genes[vars$index %in% topic9_genes] <- 1

topic10_genes <- topic_genes$Gene[topic_genes$Topic == 10]
vars$topic10_genes <- 0
vars$topic10_genes[vars$index %in% topic10_genes] <- 1

topic11_genes <- topic_genes$Gene[topic_genes$Topic == 11]
vars$topic11_genes <- 0
vars$topic11_genes[vars$index %in% topic11_genes] <- 1

topic12_genes <- topic_genes$Gene[topic_genes$Topic == 12]
vars$topic12_genes <- 0
vars$topic12_genes[vars$index %in% topic12_genes] <- 1

topic13_genes <- topic_genes$Gene[topic_genes$Topic == 13]
vars$topic13_genes <- 0
vars$topic13_genes[vars$index %in% topic13_genes] <- 1

topic14_genes <- topic_genes$Gene[topic_genes$Topic == 14]
vars$topic14_genes <- 0
vars$topic14_genes[vars$index %in% topic14_genes] <- 1

topic15_genes <- topic_genes$Gene[topic_genes$Topic == 15]
vars$topic15_genes <- 0
vars$topic15_genes[vars$index %in% topic15_genes] <- 1

topic16_genes <- topic_genes$Gene[topic_genes$Topic == 16]
vars$topic16_genes <- 0
vars$topic16_genes[vars$index %in% topic16_genes] <- 1

topic17_genes <- topic_genes$Gene[topic_genes$Topic == 17]
vars$topic17_genes <- 0
vars$topic17_genes[vars$index %in% topic17_genes] <- 1

topic18_genes <- topic_genes$Gene[topic_genes$Topic == 18]
vars$topic18_genes <- 0
vars$topic18_genes[vars$index %in% topic18_genes] <- 1

topic19_genes <- topic_genes$Gene[topic_genes$Topic == 19]
vars$topic19_genes <- 0
vars$topic19_genes[vars$index %in% topic19_genes] <- 1

## Topic QTL Loci Genes ######

topic_qtl <- fread(paste0(my_path, "results/topic-qtl-peak-lods.csv")) %>%
  mutate(pos = pos * 1e6)


topic_gene_list <- lapply(1 : NROW(topic_qtl), function(i) {
  
  na.omit(ensembl$symbol[ensembl$chr == topic_qtl$chr[i] & 
                           {ensembl$start > topic_qtl$pos[i] - 500000 & 
                               ensembl$start < topic_qtl$pos[i] + 500000} |
                           {ensembl$end > topic_qtl$pos[i] - 500000 & 
                               ensembl$end < topic_qtl$pos[i] + 500000}])
  
})

  
topic0_qtl_genes <- unlist(topic_gene_list[1:2])
vars$topic0_qtl_genes <- 0
vars$topic0_qtl_genes[vars$index %in% topic0_qtl_genes] <- 1

topic1_qtl_genes <- c(unlist(topic_gene_list[3:5]), "Sstr4")
vars$topic1_qtl_genes <- 0
vars$topic1_qtl_genes[vars$index %in% topic1_qtl_genes] <- 1

topic3_qtl_genes <- c(unlist(topic_gene_list[[6]]), "Sstr4")
vars$topic3_qtl_genes <- 0
vars$topic3_qtl_genes[vars$index %in% topic3_qtl_genes] <- 1

topic4_qtl_genes <- unlist(topic_gene_list[7:8])
vars$topic4_qtl_genes <- 0
vars$topic4_qtl_genes[vars$index %in% topic4_qtl_genes] <- 1

topic6_qtl_genes <- unlist(topic_gene_list[9:10])
vars$topic6_qtl_genes <- 0
vars$topic6_qtl_genes[vars$index %in% topic6_qtl_genes] <- 1

topic8_qtl_genes <- unlist(topic_gene_list[11:13])
vars$topic8_qtl_genes <- 0
vars$topic8_qtl_genes[vars$index %in% topic8_qtl_genes] <- 1

topic11_qtl_genes <- unlist(topic_gene_list[14])
vars$topic11_qtl_genes <- 0
vars$topic11_qtl_genes[vars$index %in% topic11_qtl_genes] <- 1

topic13_qtl_genes <- unlist(topic_gene_list[15])
vars$topic13_qtl_genes <- 0
vars$topic13_qtl_genes[vars$index %in% topic13_qtl_genes] <- 1

topic14_qtl_genes <- unlist(topic_gene_list[16])
vars$topic14_qtl_genes <- 0
vars$topic14_qtl_genes[vars$index %in% topic14_qtl_genes] <- 1

topic15_qtl_genes <- unlist(topic_gene_list[17:19])
vars$topic15_qtl_genes <- 0
vars$topic15_qtl_genes[vars$index %in% topic15_qtl_genes] <- 1

topic18_qtl_genes <- unlist(topic_gene_list[20:23])
vars$topic18_qtl_genes <- 0
vars$topic18_qtl_genes[vars$index %in% topic18_qtl_genes] <- 1

topic19_qtl_genes <- unlist(topic_gene_list[24:28])
vars$topic19_qtl_genes <- 0
vars$topic19_qtl_genes[vars$index %in% topic19_qtl_genes] <- 1



## ILC Cell Proportions ###############

## ILC3 Activated vs Non-Activated
ilc3_gwas <- fread(paste0(my_path, "results/ILC3_stressed_vs_non_qtl.csv")) %>%
  filter(lods > 200)
ilc3_gwas_genes <- unique(ilc3_gwas$ensembl_gene)
vars$ilc3_activated_qtl_genes <- 0
vars$ilc3_activated_qtl_genes[vars$gene_ids %in% ilc3_gwas_genes] <- 1

## LTi Activated vs Non-Activated
lti_gwas <- fread(paste0(my_path, "results/LTi_stressed_vs_non_qtl.csv")) %>%
  filter(lods > 200)
lti_gwas_genes <- unique(lti_gwas$ensembl_gene)
vars$lti_activated_qtl_genes <- 0
vars$lti_activated_qtl_genes[vars$gene_ids %in% lti_gwas_genes] <- 1


## ILC1 vs ILC2 proportions
ilc1_ilc2 <- fread(paste0(my_path, 'results/gwas-ilc1-ilc2-results.csv')) %>%
  filter(lods > 100)
ilc1_lic2_genes <- unique(ilc1_ilc2$ensembl_gene)
vars$ilc1_ilc2_prop_qtl_genes <- 0
vars$ilc1_ilc2_prop_qtl_genes[vars$gene_ids %in% ilc1_lic2_genes] <- 1


## ILC1 vs ILC3 proportions
ilc1_ilc3 <- fread(paste0(my_path, 'results/gwas-ilc1-ilc3-results.csv')) %>%
  filter(lods > 100)
ilc1_lic3_genes <- unique(ilc1_ilc3$ensembl_gene)
vars$ilc1_ilc3_prop_qtl_genes <- 0
vars$ilc1_ilc3_prop_qtl_genes[vars$gene_ids %in% ilc1_lic3_genes] <- 1


## ILC1 vs LTi proportions
ilc1_lti <- fread(paste0(my_path, 'results/gwas-ilc1-lti-results.csv')) %>%
  filter(lods > 100)
ilc1_lti_genes <- unique(ilc1_lti$ensembl_gene)
vars$ilc1_lti_prop_qtl_genes <- 0
vars$ilc1_lti_prop_qtl_genes[vars$gene_ids %in% ilc1_lti_genes] <- 1


## ILC2 vs ILC3 proportions
ilc2_ilc3 <- fread(paste0(my_path, 'results/gwas-ilc2-ilc3-results.csv')) %>%
  filter(lods > 150)
ilc2_lic3_genes <- unique(ilc2_ilc3$ensembl_gene)
vars$ilc2_ilc3_prop_qtl_genes <- 0
vars$ilc2_ilc3_prop_qtl_genes[vars$gene_ids %in% ilc2_lic3_genes] <- 1


## ILC2 vs LTi proportions
ilc2_lti <- fread(paste0(my_path, 'results/gwas-ilc2-lti-results.csv')) %>%
  filter(lods > 150)
ilc2_lti_genes <- unique(ilc2_lti$ensembl_gene)
vars$ilc2_lti_prop_qtl_genes <- 0
vars$ilc2_lti_prop_qtl_genes[vars$gene_ids %in% ilc2_lti_genes] <- 1


## ILC3 vs LTi
ilc3_lti <- fread(paste0(my_path, 'results/gwas-ilc3-lti-results.csv')) %>%
  filter(lods > 100)
ilc3_lti_genes <- unique(ilc3_lti$ensembl_gene)
vars$ilc3_lti_prop_qtl_genes <- 0
vars$ilc3_lti_prop_qtl_genes[vars$gene_ids %in% ilc3_lti_genes] <- 1


## Antibody QTL #######################

antibody <- fread(paste0(my_path, "results/antibody-qtl-output/Antibody_QTL_BwCorrectionCorrection.csv"))

## IgA
iga_markers <- antibody$V1[antibody$IgA_log > 5.5]
iga_qtl_genes <- unique(conseq$ensembl_gene[conseq$marker %in% iga_markers])
vars$iga_qtl_genes <- 0
vars$iga_qtl_genes[vars$gene_ids %in% iga_qtl_genes] <- 1

## IgE
ige_markers <- antibody$V1[antibody$IgE_log > 5.5]
ige_qtl_genes <- unique(conseq$ensembl_gene[conseq$marker %in% ige_markers])
vars$ige_qtl_genes <- 0
vars$ige_qtl_genes[vars$gene_ids %in% ige_qtl_genes] <- 1


## IgG1
igg1_markers <- antibody$V1[antibody$IgG1_log > 5.5]
igg1_qtl_genes <- unique(conseq$ensembl_gene[conseq$marker %in% igg1_markers])
vars$igg1_qtl_genes <- 0
vars$igg1_qtl_genes[vars$gene_ids %in% igg1_qtl_genes] <- 1


## IgG2
igg2_markers <- antibody$V1[antibody$IgG2_log > 5.5]
igg2_qtl_genes <- unique(conseq$ensembl_gene[conseq$marker %in% igg2_markers])
vars$igg2_qtl_genes <- 0
vars$igg2_qtl_genes[vars$gene_ids %in% igg2_qtl_genes] <- 1


## IgM
igm_markers <- antibody$V1[antibody$IgM_log > 5.5]
igm_qtl_genes <- unique(conseq$ensembl_gene[conseq$marker %in% igm_markers])
vars$igm_qtl_genes <- 0
vars$igm_qtl_genes[vars$gene_ids %in% igm_qtl_genes] <- 1


## Mmcp1
mmcp1_markers <- antibody$V1[antibody$Mmcp1_log > 5.5]
mmcp1_qtl_genes <- unique(conseq$ensembl_gene[conseq$marker %in% mmcp1_markers])
vars$mmcp1_qtl_genes <- 0
vars$mmcp1_qtl_genes[vars$gene_ids %in% mmcp1_qtl_genes] <- 1



## Cytokines ###################

## Steady State
lods_steady <- read.csv(paste0(my_path, "results/qtl-steady-cytokines-lods.csv"))

infg_markers_steady <- lods_steady$marker[lods_steady$IFNg > 5.5]
infg_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% infg_markers_steady])
il5_markers_steady <- lods_steady$marker[lods_steady$IL5 > 5.5]
il5_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il5_markers_steady])
tnf_markers_steady <- lods_steady$marker[lods_steady$TNFa > 5.5]
tnf_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% tnf_markers_steady])
il2_markers_steady <- lods_steady$marker[lods_steady$IL2 > 5.5]
il2_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il2_markers_steady])
il6_markers_steady <- lods_steady$marker[lods_steady$IL6 > 5.5]
il6_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il6_markers_steady])
il4_markers_steady <- lods_steady$marker[lods_steady$IL4 > 5.5]
il4_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il4_markers_steady])
il10_markers_steady <- lods_steady$marker[lods_steady$IL10 > 5.5]
il10_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il10_markers_steady])
il9_markers_steady <- lods_steady$marker[lods_steady$IL9 > 5.5]
il9_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il9_markers_steady])
il17a_markers_steady <- lods_steady$marker[lods_steady$IL17A > 5.5]
il17a_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il17a_markers_steady])
il17f_markers_steady <- lods_steady$marker[lods_steady$IL17F > 5.5]
il17f_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il17f_markers_steady])
il22_markers_steady <- lods_steady$marker[lods_steady$IL22 > 5.5]
il22_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il22_markers_steady])
il13_markers_steady <- lods_steady$marker[lods_steady$IL13 > 5.5]
il13_qtl_genes_steady_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il13_markers_steady])



vars <- vars %>%
  mutate(infg_qtl_genes_steady = as.numeric(gene_ids %in% infg_qtl_genes_steady_vec),
         il5_qtl_genes_steady = as.numeric(gene_ids %in% il5_qtl_genes_steady_vec),
         tnf_qtl_genes_steady = as.numeric(gene_ids %in% tnf_qtl_genes_steady_vec),
         il2_qtl_genes_steady = as.numeric(gene_ids %in% il2_qtl_genes_steady_vec),
         il6_qtl_genes_steady = as.numeric(gene_ids %in% il6_qtl_genes_steady_vec),
         il4_qtl_genes_steady = as.numeric(gene_ids %in% il4_qtl_genes_steady_vec),
         il10_qtl_genes_steady = as.numeric(gene_ids %in% il10_qtl_genes_steady_vec),
         il9_qtl_genes_steady = as.numeric(gene_ids %in% il9_qtl_genes_steady_vec),
         il17a_qtl_genes_steady = as.numeric(gene_ids %in% il17a_qtl_genes_steady_vec),
         il17f_qtl_genes_steady = as.numeric(gene_ids %in% il17f_qtl_genes_steady_vec),
         il22_qtl_genes_steady = as.numeric(gene_ids %in% il22_qtl_genes_steady_vec),
         il13_qtl_genes_steady = as.numeric(gene_ids %in% il13_qtl_genes_steady_vec)
         )

## Allergy
lods_allergy <- read.csv(paste0(my_path, "results/qtl-allergy-cytokines-lods.csv"))

infg_markers_allergy <- lods_allergy$marker[lods_allergy$IFNg > 5.5]
infg_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% infg_markers_allergy])
il5_markers_allergy <- lods_allergy$marker[lods_allergy$IL5 > 5.5]
il5_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il5_markers_allergy])
tnf_markers_allergy <- lods_allergy$marker[lods_allergy$TNFa > 5.5]
tnf_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% tnf_markers_allergy])
il2_markers_allergy <- lods_allergy$marker[lods_allergy$IL2 > 5.5]
il2_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il2_markers_allergy])
il6_markers_allergy <- lods_allergy$marker[lods_allergy$IL6 > 5.5]
il6_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il6_markers_allergy])
il4_markers_allergy <- lods_allergy$marker[lods_allergy$IL4 > 5.5]
il4_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il4_markers_allergy])
il10_markers_allergy <- lods_allergy$marker[lods_allergy$IL10 > 5.5]
il10_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il10_markers_allergy])
il9_markers_allergy <- lods_allergy$marker[lods_allergy$IL9 > 5.5]
il9_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il9_markers_allergy])
il17a_markers_allergy <- lods_allergy$marker[lods_allergy$IL17A > 5.5]
il17a_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il17a_markers_allergy])
il17f_markers_allergy <- lods_allergy$marker[lods_allergy$IL17F > 5.5]
il17f_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il17f_markers_allergy])
il22_markers_allergy <- lods_allergy$marker[lods_allergy$IL22 > 5.5]
il22_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il22_markers_allergy])
il13_markers_allergy <- lods_allergy$marker[lods_allergy$IL13 > 5.5]
il13_qtl_genes_allergy_vec <- unique(conseq$ensembl_gene[conseq$marker %in% il13_markers_allergy])



vars <- vars %>%
  mutate(infg_qtl_genes_allergy = as.numeric(gene_ids %in% infg_qtl_genes_allergy_vec),
         il5_qtl_genes_allergy = as.numeric(gene_ids %in% il5_qtl_genes_allergy_vec),
         tnf_qtl_genes_allergy = as.numeric(gene_ids %in% tnf_qtl_genes_allergy_vec),
         il2_qtl_genes_allergy = as.numeric(gene_ids %in% il2_qtl_genes_allergy_vec),
         il6_qtl_genes_allergy = as.numeric(gene_ids %in% il6_qtl_genes_allergy_vec),
         il4_qtl_genes_allergy = as.numeric(gene_ids %in% il4_qtl_genes_allergy_vec),
         il10_qtl_genes_allergy = as.numeric(gene_ids %in% il10_qtl_genes_allergy_vec),
         il9_qtl_genes_allergy = as.numeric(gene_ids %in% il9_qtl_genes_allergy_vec),
         il17a_qtl_genes_allergy = as.numeric(gene_ids %in% il17a_qtl_genes_allergy_vec),
         il17f_qtl_genes_allergy = as.numeric(gene_ids %in% il17f_qtl_genes_allergy_vec),
         il22_qtl_genes_allergy = as.numeric(gene_ids %in% il22_qtl_genes_allergy_vec),
         il13_qtl_genes_allergy = as.numeric(gene_ids %in% il13_qtl_genes_allergy_vec)
         )


write_csv(vars, paste0(my_path, "data/scanpy/allchannels/vars.csv"))
```


```{r}
intersect_list <- combn(colnames(vars)[c(240:283,287:288,292:315)], 2, function(x) {

  vars$index[((vars[[x[[1]]]] == 1) & (vars[[x[[2]]]] == 1))]
  
}, simplify = FALSE)

names_list <- combn(colnames(vars)[c(240:283,287:288,292:315)], 2, paste0, collapse = "_by_")
names(intersect_list) <- names_list

out_list <- Filter(function(x) length(x) > 0, intersect_list)

mat <- matrix(NA, nrow = 179, ncol = 661)
for( i in seq_along(out_list) ) {
  
  mat[1:length(out_list[[i]]), i] <- out_list[[i]]
  
}

colnames(mat) <- names(out_list)
write.csv(mat, "qtl-genes-overlap.csv")

getwd()
```


### Make Consequence Table

Make a csv file containing all the genes in the dataset.  Have a column that indicates whether that gene is 

  - expressed each specific cell type (used as phenotype in eQTL analysis)
  - Has SNPs in the Giga MUGA array
  - Was an eGene for each specific cell type
  - had a eQTL for each specific cell type
  - Was differentially expressed in each cell type / Marker gene identified for cell type
  - QTL loci genes for each topic
  - topic genes
  - antibodyQTL loci genes
  - cytokine QTL loci genes
  - cell proportion composition QTL loci genes
  
```{r ccre_table, eval = FALSE}
## starting with all ensembl genes
ccre <- conseq

## eQTL Genes #######

## Mean #######
## ILC1 eQTL Loci
ilc1_markers <- unique(ilc1_mean$marker)
ccre <- ccre %>%
  mutate(ilc1_eqtl_loci_mean = as.numeric(marker %in% ilc1_markers))

## ILC1 eQTL cis-effect associations
ilc1_markers_cis <- unique(ilc1_mean$marker[ilc1_mean$cis_effect==1])
ccre <- ccre %>%
  mutate(ilc1_ciseqtl_assoc_mean = as.numeric(marker %in% ilc1_markers_cis))

## CV #####
## ILC1 eQTL Loci
ilc1_markers <- unique(ilc1_cv$marker)
ccre <- ccre %>%
  mutate(ilc1_eqtl_loci_cv = as.numeric(marker %in% ilc1_markers))

## ILC1 eQTL cis-effect associations
ilc1_markers_cis <- unique(ilc1_cv$marker[ilc1_cv$cis_effect==1])
ccre <- ccre %>%
  mutate(ilc1_ciseqtl_assoc_cv = as.numeric(marker %in% ilc1_markers_cis))


## Mean #######
## ILC2 eQTL Loci
ilc2_markers <- unique(ilc2_mean$marker)
ccre <- ccre %>%
  mutate(ilc2_eqtl_loci_mean = as.numeric(marker %in% ilc2_markers))

## ILC2 eQTL cis-effect associations
ilc2_markers_cis <- unique(ilc2_mean$marker[ilc2_mean$cis_effect==1])
ccre <- ccre %>%
  mutate(ilc2_ciseqtl_assoc_mean = as.numeric(marker %in% ilc2_markers_cis))

## CV #####
## ILC2 eQTL Loci
ilc2_markers <- unique(ilc2_cv$marker)
ccre <- ccre %>%
  mutate(ilc2_eqtl_loci_cv = as.numeric(marker %in% ilc2_markers))

## ILC2 eQTL cis-effect associations
ilc2_markers_cis <- unique(ilc2_cv$marker[ilc2_cv$cis_effect==1])
ccre <- ccre %>%
  mutate(ilc2_ciseqtl_assoc_cv = as.numeric(marker %in% ilc2_markers_cis))



## Mean #######
## ILC3 eQTL Loci
ilc3_markers <- unique(ilc3_mean$marker)
ccre <- ccre %>%
  mutate(ilc3_eqtl_loci_mean = as.numeric(marker %in% ilc3_markers))

## ILC3 eQTL cis-effect associations
ilc3_markers_cis <- unique(ilc3_mean$marker[ilc3_mean$cis_effect==1])
ccre <- ccre %>%
  mutate(ilc3_ciseqtl_assoc_mean = as.numeric(marker %in% ilc3_markers_cis))

## CV #####
## ILC3 eQTL Loci
ilc3_markers <- unique(ilc3_cv$marker)
ccre <- ccre %>%
  mutate(ilc3_eqtl_loci_cv = as.numeric(marker %in% ilc3_markers))

## ILC3 eQTL cis-effect associations
ilc3_markers_cis <- unique(ilc3_cv$marker[ilc3_cv$cis_effect==1])
ccre <- ccre %>%
  mutate(ilc3_ciseqtl_assoc_cv = as.numeric(marker %in% ilc3_markers_cis))



## Mean #######
## LTi eQTL Loci
lti_markers <- unique(lti_mean$marker)
ccre <- ccre %>%
  mutate(lti_eqtl_loci_mean = as.numeric(marker %in% lti_markers))

## LTi eQTL cis-effect associations
lti_markers_cis <- unique(lti_mean$marker[lti_mean$cis_effect==1])
ccre <- ccre %>%
  mutate(lti_ciseqtl_assoc_mean = as.numeric(marker %in% lti_markers_cis))

## CV #####
## LTi eQTL Loci
lti_markers <- unique(lti_cv$marker)
ccre <- ccre %>%
  mutate(lti_eqtl_loci_cv = as.numeric(marker %in% lti_markers))

## LTi eQTL cis-effect associations
lti_markers_cis <- unique(lti_cv$marker[lti_cv$cis_effect==1])
ccre <- ccre %>%
  mutate(lti_ciseqtl_assoc_cv = as.numeric(marker %in% lti_markers_cis))



## Topic QTL Loci Genes ######

topic_qtl <- fread(paste0(my_path, "results/topic-qtl-lods.csv"), data.table = FALSE)

topic0_markers <- topic_qtl$marker[topic_qtl$topic0 > 5.5]
topic1_markers <- topic_qtl$marker[topic_qtl$topic1 > 5.5]
topic2_markers <- topic_qtl$marker[topic_qtl$topic2 > 5.5]
topic3_markers <- topic_qtl$marker[topic_qtl$topic3 > 5.5]
topic4_markers <- topic_qtl$marker[topic_qtl$topic4 > 5.5]
topic5_markers <- topic_qtl$marker[topic_qtl$topic5 > 5.5]
topic6_markers <- topic_qtl$marker[topic_qtl$topic6 > 5.5]
topic7_markers <- topic_qtl$marker[topic_qtl$topic7 > 5.5]
topic8_markers <- topic_qtl$marker[topic_qtl$topic8 > 5.5]
topic9_markers <- topic_qtl$marker[topic_qtl$topic9 > 5.5]
topic10_markers <- topic_qtl$marker[topic_qtl$topic10 > 5.5]
topic11_markers <- topic_qtl$marker[topic_qtl$topic11 > 5.5]
topic12_markers <- topic_qtl$marker[topic_qtl$topic12 > 5.5]
topic13_markers <- topic_qtl$marker[topic_qtl$topic13 > 5.5]
topic14_markers <- topic_qtl$marker[topic_qtl$topic14 > 5.5]
topic15_markers <- topic_qtl$marker[topic_qtl$topic15 > 5.5]
topic16_markers <- topic_qtl$marker[topic_qtl$topic16 > 5.5]
topic17_markers <- topic_qtl$marker[topic_qtl$topic17 > 5.5]
topic18_markers <- topic_qtl$marker[topic_qtl$topic18 > 5.5]
topic19_markers <- topic_qtl$marker[topic_qtl$topic19 > 5.5]

ccre <- ccre %>%
  mutate(topic0_qtl = as.numeric(marker %in% topic0_markers),
         topic1_qtl = as.numeric(marker %in% topic1_markers),
         topic2_qtl = as.numeric(marker %in% topic2_markers),
         topic3_qtl = as.numeric(marker %in% topic3_markers),
         topic4_qtl = as.numeric(marker %in% topic4_markers),
         topic5_qtl = as.numeric(marker %in% topic5_markers),
         topic6_qtl = as.numeric(marker %in% topic6_markers),
         topic7_qtl = as.numeric(marker %in% topic7_markers),
         topic8_qtl = as.numeric(marker %in% topic8_markers),
         topic9_qtl = as.numeric(marker %in% topic9_markers),
         topic10_qtl = as.numeric(marker %in% topic10_markers),
         topic11_qtl = as.numeric(marker %in% topic11_markers),
         topic12_qtl = as.numeric(marker %in% topic12_markers),
         topic13_qtl = as.numeric(marker %in% topic13_markers),
         topic14_qtl = as.numeric(marker %in% topic14_markers),
         topic15_qtl = as.numeric(marker %in% topic15_markers),
         topic16_qtl = as.numeric(marker %in% topic16_markers),
         topic17_qtl = as.numeric(marker %in% topic17_markers),
         topic18_qtl = as.numeric(marker %in% topic18_markers),
         topic19_qtl = as.numeric(marker %in% topic19_markers))



## ILC Cell Proportions ###############

## ILC3 Activated vs Non-Activated
ilc3_gwas <- fread(paste0(my_path, "results/ILC3_stressed_vs_non_qtl.csv")) %>%
  filter(lods > 200)
ilc3_gwas_markers <- unique(ilc3_gwas$marker)


## LTi Activated vs Non-Activated
lti_gwas <- fread(paste0(my_path, "results/LTi_stressed_vs_non_qtl.csv")) %>%
  filter(lods > 200)
lti_gwas_markers <- unique(lti_gwas$marker)


## ILC1 vs ILC2 proportions
ilc1_ilc2 <- fread(paste0(my_path, 'results/gwas-ilc1-ilc2-results.csv')) %>%
  filter(lods > 100)
ilc1_lic2_markers <- unique(ilc1_ilc2$marker)


## ILC1 vs ILC3 proportions
ilc1_ilc3 <- fread(paste0(my_path, 'results/gwas-ilc1-ilc3-results.csv')) %>%
  filter(lods > 100)
ilc1_lic3_markers <- unique(ilc1_ilc3$marker)


## ILC2 vs ILC3 proportions
ilc2_ilc3 <- fread(paste0(my_path, 'results/gwas-ilc2-ilc3-results.csv')) %>%
  filter(lods > 100)
ilc2_lic3_markers <- unique(ilc2_ilc3$marker)


## ILC2 vs LTi proportions
ilc2_lti <- fread(paste0(my_path, 'results/gwas-ilc2-lti-results.csv')) %>%
  filter(lods > 100)
ilc2_lti_markers <- unique(ilc2_lti$marker)


## ILC3 vs LTi
ilc3_lti <- fread(paste0(my_path, 'results/gwas-ilc3-lti-results.csv')) %>%
  filter(lods > 100)
ilc3_lti_markers <- unique(ilc3_lti$marker)


ccre <- ccre %>%
  mutate(ilc3_activated_qtl = as.numeric(marker %in% ilc3_gwas_markers),
         lti_activated_qtl = as.numeric(marker %in% lti_gwas_markers),
         ilc1_ilc2_prop_qtl = as.numeric(marker %in% ilc1_lic2_markers),
         ilc1_lic3_prop_qtl = as.numeric(marker %in% ilc1_lic3_markers),
         ilc2_ilc3_prop_qtl = as.numeric(marker %in% ilc2_lic3_markers),
         ilc3_lti_prop_qtl = as.numeric(marker %in% ilc3_lti_markers))

## Antibody QTL #######################

antibody <- fread(paste0(my_path, "results/antibody-qtl-output/Antibody_QTL_BwCorrectionCorrection.csv"))

## IgA
iga_markers <- antibody$V1[antibody$IgA_log > 5.5]

## IgE
ige_markers <- antibody$V1[antibody$IgE_log > 5.5]

## IgG1
igg1_markers <- antibody$V1[antibody$IgG1_log > 5.5]

## IgG2
igg2_markers <- antibody$V1[antibody$IgG2_log > 5.5]

## IgM
igm_markers <- antibody$V1[antibody$IgM_log > 5.5]

## Mmcp1
mmcp1_markers <- antibody$V1[antibody$Mmcp1_log > 5.5]

ccre <- ccre %>%
  mutate(iga_qtl = as.numeric(marker %in% iga_markers),
         ige_qtl = as.numeric(marker %in% ige_markers),
         igg1_qtl = as.numeric(marker %in% igg1_markers),
         igg2_qtl = as.numeric(marker %in% igg2_markers),
         igm_qtl = as.numeric(marker %in% igm_markers),
         mmcp1_qtl = as.numeric(marker %in% mmcp1_markers))

## Cytokines ###################
## Steady State
lods_steady <- read.csv(paste0(my_path, "results/qtl-cytokines-steady-lods.csv"))

ifng_markers_steady <- lods_steady$marker[lods_steady$IFNg > 5.5]
il5_markers_steady <- lods_steady$marker[lods_steady$IL5 > 5.5]
tnf_markers_steady <- lods_steady$marker[lods_steady$TNFa > 5.5]
il2_markers_steady <- lods_steady$marker[lods_steady$IL2 > 5.5]
il6_markers_steady <- lods_steady$marker[lods_steady$IL6 > 5.5]
il4_markers_steady <- lods_steady$marker[lods_steady$IL4 > 5.5]
il10_markers_steady <- lods_steady$marker[lods_steady$IL10 > 5.5]
il9_markers_steady <- lods_steady$marker[lods_steady$IL9 > 5.5]
il17a_markers_steady <- lods_steady$marker[lods_steady$IL17A > 5.5]
il17f_markers_steady <- lods_steady$marker[lods_steady$IL17F > 5.5]
il22_markers_steady <- lods_steady$marker[lods_steady$IL22 > 5.5]
il13_markers_steady <- lods_steady$marker[lods_steady$IL13 > 5.5]



ccre <- ccre %>%
  mutate(
    ifng_qtl_steady = as.numeric(marker %in% ifng_markers_steady),
    il5_qtl_steady = as.numeric(marker %in% il5_markers_steady),
    tnf_qtl_steady = as.numeric(marker %in% tnf_markers_steady),
    il2_qtl_steady = as.numeric(marker %in% il2_markers_steady),
    il6_qtl_steady = as.numeric(marker %in% il6_markers_steady),
    il4_qtl_steady = as.numeric(marker %in% il4_markers_steady),
    il10_qtl_steady = as.numeric(marker %in% il10_markers_steady),
    il9_qtl_steady = as.numeric(marker %in% il9_markers_steady),
    il17a_qtl_steady = as.numeric(marker %in% il17a_markers_steady),
    il17f_qtl_steady = as.numeric(marker %in% il17f_markers_steady),
    il22_qtl_steady = as.numeric(marker %in% il22_markers_steady),
    il13_qtl_steady = as.numeric(marker %in% il13_markers_steady)
  )


## Allergy State
lods_allergy <- read.csv(paste0(my_path, "results/qtl-cytokines-allergy-lods.csv"))

ifng_markers_allergy <- lods_allergy$marker[lods_allergy$IFNg > 5.5]
il5_markers_allergy <- lods_allergy$marker[lods_allergy$IL5 > 5.5]
tnf_markers_allergy <- lods_allergy$marker[lods_allergy$TNFa > 5.5]
il2_markers_allergy <- lods_allergy$marker[lods_allergy$IL2 > 5.5]
il6_markers_allergy <- lods_allergy$marker[lods_allergy$IL6 > 5.5]
il4_markers_allergy <- lods_allergy$marker[lods_allergy$IL4 > 5.5]
il10_markers_allergy <- lods_allergy$marker[lods_allergy$IL10 > 5.5]
il9_markers_allergy <- lods_allergy$marker[lods_allergy$IL9 > 5.5]
il17a_markers_allergy <- lods_allergy$marker[lods_allergy$IL17A > 5.5]
il17f_markers_allergy <- lods_allergy$marker[lods_allergy$IL17F > 5.5]
il22_markers_allergy <- lods_allergy$marker[lods_allergy$IL22 > 5.5]
il13_markers_allergy <- lods_allergy$marker[lods_allergy$IL13 > 5.5]



ccre <- ccre %>%
  mutate(
    ifng_qtl_allergy = as.numeric(marker %in% ifng_markers_allergy),
    il5_qtl_allergy = as.numeric(marker %in% il5_markers_allergy),
    tnf_qtl_allergy = as.numeric(marker %in% tnf_markers_allergy),
    il2_qtl_allergy = as.numeric(marker %in% il2_markers_allergy),
    il6_qtl_allergy = as.numeric(marker %in% il6_markers_allergy),
    il4_qtl_allergy = as.numeric(marker %in% il4_markers_allergy),
    il10_qtl_allergy = as.numeric(marker %in% il10_markers_allergy),
    il9_qtl_allergy = as.numeric(marker %in% il9_markers_allergy),
    il17a_qtl_allergy = as.numeric(marker %in% il17a_markers_allergy),
    il17f_qtl_allergy = as.numeric(marker %in% il17f_markers_allergy),
    il22_qtl_allergy = as.numeric(marker %in% il22_markers_allergy),
    il13_qtl_allergy = as.numeric(marker %in% il13_markers_allergy)
  )

write_csv(ccre, paste0(my_path, "data/GM_SNPS_Consequence_cCRE.csv"))
```


## Figure 1

### Figure 1A


```{r fig1a}
fig1a_plot <- plot_df %>%
  filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
  ggplot(., aes(x = X_tsne1, y = X_tsne2, color = called_cell_types)) + 
           geom_point(shape = 46)
```


### Figure 1B

### Ven Diagram eQTL Genes in loci

Common and cell type specific eQTLs identification

Overlaps of genes located within .5 Mb of significant eQTL loci accross the major lineages of innnate lymphoid cells.

```{r fig1b_eqtl}
myCol <- RColorBrewer::brewer.pal(4, "Pastel2")
# Chart
venn.diagram(
        x = list(na.omit(vars$index[vars$ilc1_eqtl_loci_genes_cv == 1]),
                 na.omit(vars$index[vars$ilc2_eqtl_loci_genes_cv == 1]),
                 na.omit(vars$index[vars$ilc3_eqtl_loci_genes_cv == 1]),
                 na.omit(vars$index[vars$lti_eqtl_loci_genes_cv == 1])),
        category.names = c("ILC1" , "ILC2 " , "ILC3", "LTi"),
        filename = 'eqtl-loci-genes-venn-diagramm.png',
        output=TRUE,
        main = "eQTL Genes in Significant Loci by Cell Type",
        main.cex = 0.5,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 480 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .5,
        fontface = "bold",
        fontfamily = "sans",
        cat.cex = .5
        
)
```


### Ven Diagram eGenes

Common and cell type specific eGenes identification

Overlaps of significant eGenes accross the major lineages of innnate lymphoid cells.

```{r fig1c_egenes}
myCol <- RColorBrewer::brewer.pal(4, "Pastel2")
# Chart
venn.diagram(
        x = list(na.omit(vars$index[vars$ilc1_egenes_cv == 1]),
                 na.omit(vars$index[vars$ilc2_egenes_cv == 1]),
                 na.omit(vars$index[vars$ilc3_egenes_cv == 1]),
                 na.omit(vars$index[vars$lti_egenes_cv == 1])),
        category.names = c("ILC1" , "ILC2 " , "ILC3", "LTi"),
        filename = 'eGenes-genes-venn-diagramm.png',
        output=TRUE,
        main = "eGenes by Cell Type",
        main.cex = 0.5,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 480 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .5,
        fontface = "bold",
        fontfamily = "sans",
        cat.cex = .5
        
)
```


### Figure 1D

```{r fig1d}
## Plot file list
plot_files <- c(
  paste0(my_path, "results/qtl-plot-lods-ILC1-cv.csv"),
  paste0(my_path, "results/qtl-plot-lods-ILC2-cv.csv"),
  paste0(my_path, "results/qtl-plot-lods-NCR1+ ILC3-cv.csv"),
  paste0(my_path, "results/qtl-plot-lods-Lti ILC3-cv.csv")
)


## Loading Marker Map
cat("Loading SNPs...", "\n")
# load(paste0(geno_path, "GM_snps.Rdata"))
GM_snps <- fread(paste0(my_path, "data/reference/GM_SNPS_Consequence.csv"))
marker_map <- readRDS(paste0(geno_path, "Regev_map_20171221.rds"))
xpos <- qtl2:::map_to_xpos(marker_map, gap = 25)
xchrbound <- qtl2:::map_to_boundaries(marker_map, gap = 25)

cat("Loading Genes...", "\n")
ensembl2 <- fread(paste0(my_path, "data/reference/ensembl.Mus_musculus.GRCm38.93.csv"))
ensembl2 <- ensembl2[chr %in% c(as.character(1:19), "X"), ]
geneids <- fread(paste0(my_path, "data/reference/GeneID.csv"), col.names = c("symbol", "gene"))
  

## Creating Gene Map
gene_map <- split(ensembl2, f = ensembl2$chr)
for( nm in names(gene_map) ) {
  vec <- gene_map[[nm]]$start
  names(vec) <- gene_map[[nm]]$gene
  gene_map[[nm]] <- vec
}
gene_map <- gene_map[c(as.character(1:19), "X")]
ypos <- qtl2:::map_to_xpos(gene_map, gap = 25)
ychrbound <- qtl2:::map_to_boundaries(gene_map, gap = 25)

i <- 1
out <- list()
out_lods <- list()
df <- NULL
plot_title <- c("ILC1", "ILC2", "LTi", "ILC3")
## CoefVar: c(36:38,40)
## Mean: c(65:67,69)
for( f in plot_files ) {
  
  cat('Loading file: ', f, '\n')
  lods <- fread(f)
  
  cat('Identifying Cis-effects \n')
  lods <- lods %>%
    left_join(
      {GM_snps %>% 
          mutate(marker_pos = pos) %>% 
          select(marker = marker, 
                 marker_pos, 
                 conseq_clean,
                 type,
                 ensembl_gene)}
    ) %>%
    left_join(
      {ensembl2 %>% select(gene_start = start, gene_end = end, gene = gene)}
    ) %>%
    mutate(cis_effect = as.numeric(marker_chr == gene_chr & 
             {marker_pos > (gene_start - 1e6) & marker_pos < (gene_end + 1e6)})) %>%
    mutate(value_adj = ifelse(cis_effect==1, value + 6, value), 
           cell_type = plot_title[i]) %>%
    left_join(geneids) %>%
    filter(value_adj > 10)

  cat('Plot File: ', plot_title[i], '\n')
  out_lods[[i]] <- lods
  i <- i + 1

}

out_df <- do.call(rbind, out_lods)
table(out_df$symbol[out_df$cis_effect==1], out_df$cell_type[out_df$cis_effect==1])


cat("Plotting LODs...", "\n")
p <- ggplot(out_lods[[1]], aes(x = xpos, y = ypos, color = cis_effect)) +
    # labs(x = "Chromosome of SNP", y = "Chromosome of Gene") +
    labs(x = "Chromosome of SNP", y = "Chromosome of Gene", title = plot_title[i]) + # title = paste0(plot_title, " (", data_subset[i], ")")) +
    geom_point(shape = 20) +
    scale_x_continuous(breaks = apply(xchrbound, 2, median), label = c(as.character(1:19), "X")) + 
    scale_y_continuous(breaks = apply(ychrbound, 2, median), label = c(as.character(1:19), "X")) + 
    scale_color_continuous(low="lightgrey", high="black") +
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position="none")



# conseq_clean, 
p <- ggplot(lods, aes(x = xpos, y = ypos, color=ctcf)) +
  labs(x = "Chromosome of SNP", y = "Chromosome of Gene", title = paste0(plot_title, " (", data_subset, ") ", geneset_name)) +
  # geom_point(shape = c(46, 8)[(lods$geneset=="Yes") + 1]) +
  geom_point(shape = 20) +
  scale_x_continuous(breaks = apply(xchrbound, 2, median), label = c(as.character(1:19), "X")) +
  scale_y_continuous(breaks = apply(ychrbound, 2, median), label = c(as.character(1:19), "X")) +
  theme_bw() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))

```

```{r}
ccre %>% group_by(type.y) %>% summarise_at(vars(ilc1_eqtl_loci_mean:mmcp1_qtl), sum)

```


```{r}
ccre_cor <- round(cor(ccre[,37:84], use = "pairwise.complete.obs"), 3)
ccre_cor[is.na(ccre_cor)] <- 0
ccre_cor[abs(ccre_cor) < 0.15] <- 0
net <- graph_from_adjacency_matrix(ccre_cor, mode = "undirected", diag = FALSE, weighted = TRUE)
visNetwork::visIgraph(net)

```


### Figure 1E

```{r}

```


### Figure 1F

Tbet + Nrc1(NKp46) + Rorc

```{r}
ncr1 <- my_plot(plot_df, "Ncr1") + my_theme
tbet <- my_plot(plot_df, "Tbet") + my_theme
rorc <- my_plot(plot_df, "Rorc") + my_theme
ccr6 <- my_plot(plot_df, "Ccr6") + my_theme

ggsave("tsne-nrc1.png", ncr1, dpi = 300)
ggsave("tsne-tbet.png", tbet, dpi = 300)
ggsave("tsne-rorc.png", rorc, dpi = 300)
ggsave("tsne-ccr6.png", ccr6, dpi = 300)
```


```{r}
## triple positive cells
plot_df$triple_pos <- as.numeric(plot_df$Tbet > 0 & plot_df$Ncr1 > 0 & plot_df$Rorc > 0)

triple_pos <- plot_df %>%
    dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
  my_plot(., "triple_pos", "tsne") +
  my_theme +
  labs(title = "Tbet + Nrc1(NKp46) + Rorc")
ggsave("tsne-ilcs-tbet-nrc1-rorc.png", triple_pos, dpi = 300)
```


```{r}
triple_pos_df <- df[df$triple_pos == 1,]
```

```{r}
triple_pos_obj <- subset(obj, cells = triple_pos_df$index)
Idents(triple_pos_obj) <- "louvain_labels"
triple_pos_de <- FindMarkers(triple_pos_obj, ident.1 = 4, ident.2 = 8)
```


### Figure 1G

```{r}
conseq %>% group_by(conseq_clean) %>% summarise(count = n())
```


### Figure 1: Supplementary Figures


```{r}
plot_df$Mhc2 <- apply(plot_df[,grep("^H2\\.", colnames(plot_df))], 1, mean)
## making topics and pathways a different color
plotgenes <- colnames(plot_df)[c(2:84, 158:246)]
plotpathways <- colnames(plot_df)[c(97:111, 121:140)]

for( g in plotgenes ) {
  
  pout <- plot_df %>%
    dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
    my_plot(., g) + my_blank_theme
  
  ggsave(filename = paste0(my_path, "results/figures/tsne-ilcs-", g, ".png"), 
         plot = pout, 
         dpi = 300)
  
}


for( t in plotpathways ) {
  
  # my_plot(plot_df, "topic3") + scale_color_gradient(low="lightgrey", high="darkgreen")
  pout <- plot_df %>%
    dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
    my_plot(., t) + 
    scale_color_gradient(low = "darkblue", high = "lightgreen") + 
    my_blank_theme
  
  ggsave(filename = paste0(my_path, "results/figures/tsne-ilcs-", t, ".png"), 
         plot = pout, 
         dpi = 300)
}
```



```{r colorcompare}
p1 <- plot_df %>%
    dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
    my_plot(., t) + 
    scale_color_gradient(low = "lightgrey", high = "darkgreen") + 
    my_blank_theme +
  labs(title = paste("low = lightgrey, high = darkgreen"))

p2 <- plot_df %>%
    dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
    my_plot(., t) + 
    scale_color_gradient(low = "lightblue", high = "lightgreen") + 
    my_blank_theme +
  labs(title = paste("low = lightblue, high = lightgreen"))

p3 <- plot_df %>%
    dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
    my_plot(., t) + 
    scale_color_gradient(low = "lightblue", high = "darkgreen") + 
    my_blank_theme +
  labs(title = paste("low = lightblue, high = darkgreen"))

p4 <- plot_df %>%
    dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
    my_plot(., t) + 
    scale_color_gradient(low = "blue", high = "lightgreen") + 
    my_blank_theme +
  labs(title = paste("low = blue, high = lightgreen"))

p5 <- plot_df %>%
    dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
    my_plot(., t) + 
    scale_color_gradient(low = "blue", high = "darkgreen") + 
    my_blank_theme +
  labs(title = paste("low = blue, high = darkgreen"))

p6 <- plot_df %>%
    dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>%
    my_plot(., t) + 
    scale_color_gradient(low = "darkblue", high = "lightgreen") + 
    my_blank_theme +
  labs(title = paste("low = darkblue, high = lightgreen"))


out <- plot_grid(p1,p2,p3,p4,p5,p6, nrow = 2)
```


#### Supplementary Figure 1d

```{r supfig1d_table}

## ILC1
ilc1_genes <- sum(vars$ilc1_expressed == 1, na.rm = TRUE)
ilc1_degenes <- sum(vars$t_qval_8[vars$selected] == 0)
ilc1_egenes <- sum(vars$ilc1_egenes_cv == 1, na.rm = TRUE)
ilc1_eqtl_genes <- sum(vars$ilc1_eqtl_loci_genes_cv == 1, na.rm = TRUE)
ilc1_cis_eqtl_genes <- sum(vars$ilc1_ciseqtl_assoc_cv == 1, na.rm = TRUE)
ilc1_genes_w_snps <- sum(vars$ilc1_expressed == 1 & vars$in_gigamuga == 1, na.rm = TRUE)

## ILC2
ilc2_genes <- sum(vars$ilc2_expressed == 1, na.rm = TRUE)
ilc2_degenes <- sum(vars$t_qval_3[vars$selected] == 0)
ilc2_egenes <- sum(vars$ilc2_egenes_cv == 1, na.rm = TRUE)
ilc2_eqtl_genes <- sum(vars$ilc2_eqtl_loci_genes_cv == 1, na.rm = TRUE)
ilc2_cis_eqtl_genes <- sum(vars$ilc2_ciseqtl_assoc_cv == 1, na.rm = TRUE)
ilc2_genes_w_snps <- sum(vars$ilc2_expressed == 1 & vars$in_gigamuga == 1, na.rm = TRUE)


## ILC3
ilc3_genes <- sum(vars$ilc3_expressed == 1, na.rm = TRUE)
ilc3_degenes <- sum(vars$t_qval_2[vars$selected] == 0 | vars$t_qval_4[vars$selected] == 0)
ilc3_egenes <- sum(vars$ilc3_egenes_cv == 1, na.rm = TRUE)
ilc3_eqtl_genes <- sum(vars$ilc3_eqtl_loci_genes_cv == 1, na.rm = TRUE)
ilc3_cis_eqtl_genes <- sum(vars$ilc3_ciseqtl_assoc_cv == 1, na.rm = TRUE)
ilc3_genes_w_snps <- sum(vars$ilc3_expressed == 1 & vars$in_gigamuga == 1, na.rm = TRUE)


## LTi-Like
lti_genes <- sum(vars$lti_expressed == 1, na.rm = TRUE)
lti_degenes <- sum(vars$t_qval_1[vars$selected] == 0 | vars$t_qval_5[vars$selected] == 0)
lti_egenes <- sum(vars$lti_egenes_cv == 1, na.rm = TRUE)
lti_eqtl_genes <- sum(vars$lti_eqtl_loci_genes_cv == 1, na.rm = TRUE)
lti_cis_eqtl_genes <- sum(vars$lti_ciseqtl_assoc_cv == 1, na.rm = TRUE)
lti_genes_w_snps <- sum(vars$lti_expressed == 1 & vars$in_gigamuga == 1, na.rm = TRUE)


## All ILCs
all_ilcs_genes <- sum({vars$ilc1_expressed == 1 | 
    vars$ilc2_expressed == 1 | 
    vars$ilc3_expressed == 1 | 
    vars$lti_expressed == 1}, na.rm = TRUE)

all_ilcs_degenes <- sum({vars$t_qval_1[vars$selected] == 0 |
    vars$t_qval_2[vars$selected] == 0 |
    vars$t_qval_3[vars$selected] == 0 |
    vars$t_qval_4[vars$selected] == 0 |
    vars$t_qval_5[vars$selected] == 0 |
    vars$t_qval_8[vars$selected] == 0})

all_ilcs_egenes <- sum({vars$ilc1_egenes_cv == 1 | 
    vars$ilc2_egenes_cv == 1 | 
    vars$ilc3_egenes_cv == 1 | 
    vars$lti_egenes_cv == 1}, na.rm = TRUE)

all_ilcs_eqtl_genes <- sum({vars$ilc1_eqtl_loci_genes_cv == 1 |
    vars$ilc2_eqtl_loci_genes_cv == 1 |
    vars$ilc3_eqtl_loci_genes_cv == 1 |
    vars$lti_eqtl_loci_genes_cv == 1}, na.rm = TRUE)

all_ilcs_cis_eqtl_genes <- sum({vars$ilc1_ciseqtl_assoc_cv == 1 |
    vars$ilc2_ciseqtl_assoc_cv == 1 |
    vars$ilc3_ciseqtl_assoc_cv == 1 |
    vars$lti_ciseqtl_assoc_cv == 1}, na.rm = TRUE)

all_ilcs_genes_w_snps <- sum({{vars$ilc1_expressed == 1 & vars$in_gigamuga == 1} |
    {vars$ilc2_expressed == 1 & vars$in_gigamuga == 1} |
    {vars$ilc3_expressed == 1 & vars$in_gigamuga == 1} |
    {vars$lti_expressed == 1 & vars$in_gigamuga == 1}}, na.rm = TRUE)


summary_df <- data.frame(
  genes = c(ilc1_genes, ilc2_genes, ilc3_genes, lti_genes, all_ilcs_genes),
  de_genes = c(ilc1_degenes, ilc2_degenes, ilc3_degenes, lti_degenes, all_ilcs_degenes),
  egenes = c(ilc1_egenes, ilc2_egenes, ilc3_egenes, lti_egenes, all_ilcs_egenes),
  eqtl_genes = c(ilc1_eqtl_genes, ilc2_eqtl_genes, ilc3_eqtl_genes, lti_eqtl_genes, all_ilcs_eqtl_genes),
  cis_eqtl_genes = c(ilc1_cis_eqtl_genes, ilc2_cis_eqtl_genes, ilc3_cis_eqtl_genes, lti_cis_eqtl_genes, all_ilcs_cis_eqtl_genes),
  genes_w_snps = c(ilc1_genes_w_snps, ilc2_genes_w_snps, ilc3_genes_w_snps, lti_genes_w_snps, all_ilcs_genes_w_snps)
  )
```


## Figure 2

### Figure 2A

SNP/eQTL classification into cCRE, enhancer promoter etc


```{r fig2a}

## Pie ggplot function
#Create a custom color scale
# library(RColorBrewer)
myColors <- brewer.pal(9,"Set1")
names(myColors) <- unique(na.omit(ccre$type.y))
# colScale <- scale_colour_manual(name = "grp",values = myColors)
pie_ggplot <- function( data ) {
  
  ggplot(data, aes(x="", y = value, fill = group)) + 
    geom_bar(stat="identity", width=1) +
    coord_polar("y", start = 0) +
    theme_void() +
    scale_colour_manual(name = "group", values = myColors, aesthetics = "fill")
    # geom_text(aes(y = ypos, label = prop), color = "white", size = 6) +
    # scale_fill_brewer(palette = "Set1", values = myColors)
  
}


# # Create Data ILC1 ccre
ilc1_eqtl_ccre <- table(ccre$type.y[ccre$ilc1_eqtl_loci_mean==1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

ilc1_ccre_plot <- pie_ggplot(ilc1_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-ilc1-eqtl.png"), 
       plot = ilc1_ccre_plot, 
       dpi = 320)


## cis-eQTL
ilc1_cis_eqtl_ccre <- table(ccre$type.y[ccre$ilc1_ciseqtl_assoc_mean == 1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

ilc1_cis_ccre_plot <- pie_ggplot(ilc1_cis_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-ilc1-cis-eqtl.png"), 
       plot = ilc1_cis_ccre_plot,
       dpi = 320)


## trans-eQTL
ilc1_trans_eqtl_ccre <- table(ccre$type.y[ccre$ilc1_ciseqtl_assoc_mean == 0 & ccre$ilc1_eqtl_loci_mean == 1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

ilc1_trans_ccre_plot <- pie_ggplot(ilc1_trans_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-ilc1-trans-eqtl.png"),
       plot = ilc1_trans_ccre_plot,
       dpi = 320)



# # Create Data ILC2 ccre
ilc2_eqtl_ccre <- table(ccre$type.y[ccre$ilc2_eqtl_loci_mean==1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

ilc2_ccre_plot <- pie_ggplot(ilc2_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-ilc2-eqtl.png"), 
       plot = ilc2_ccre_plot,
       dpi = 320)


## cis-eQTL
ilc2_cis_eqtl_ccre <- table(ccre$type.y[ccre$ilc2_ciseqtl_assoc_mean == 1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

ilc2_ccre_plot <- pie_ggplot(ilc2_cis_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-ilc2-cis-eqtl.png"), 
       plot = ilc2_ccre_plot,
       dpi = 320)


## trans-eQTL
ilc2_trans_eqtl_ccre <- table(ccre$type.y[ccre$ilc2_ciseqtl_assoc_mean == 0 & ccre$ilc2_eqtl_loci_mean == 1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

ilc2_ccre_plot <- pie_ggplot(ilc2_trans_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-ilc2-trans-eqtl.png"), 
       plot = ilc2_ccre_plot,
       dpi = 320)


# # Create Data ILC3 ccre
ilc3_eqtl_ccre <- table(ccre$type.y[ccre$ilc3_eqtl_loci_mean==1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

ilc3_ccre_plot <- pie_ggplot(ilc3_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-ilc3-eqtl.png"), 
       plot = ilc3_ccre_plot,
       dpi = 320)


## cis-eQTL
ilc3_cis_eqtl_ccre <- table(ccre$type.y[ccre$ilc3_ciseqtl_assoc_mean == 1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

ilc3_ccre_plot <- pie_ggplot(ilc3_cis_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-ilc3-cis-eqtl.png"), 
       plot = ilc3_ccre_plot,
       dpi = 320)


## trans-eQTL
ilc3_trans_eqtl_ccre <- table(ccre$type.y[ccre$ilc3_ciseqtl_assoc_mean == 0 & ccre$ilc3_eqtl_loci_mean == 1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

ilc3_ccre_plot <- pie_ggplot(ilc3_trans_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-ilc3-trans-eqtl.png"), 
       plot = ilc3_ccre_plot,
       dpi = 320)


# # Create Data LTi ccre
lti_eqtl_ccre <- table(ccre$type.y[ccre$lti_eqtl_loci_mean==1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

lti_ccre_plot <- pie_ggplot(lti_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-lti-eqtl.png"), 
       plot = lti_ccre_plot,
       dpi = 320)


## cis-eQTL
lti_cis_eqtl_ccre <- table(ccre$type.y[ccre$lti_ciseqtl_assoc_mean == 1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

lti_ccre_plot <- pie_ggplot(lti_cis_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-lti-cis-eqtl.png"), 
       plot = lti_ccre_plot,
       dpi = 320)


## trans-eQTL
lti_trans_eqtl_ccre <- table(ccre$type.y[ccre$lti_ciseqtl_assoc_mean == 0 & ccre$lti_eqtl_loci_mean == 1]) %>%
  as.data.frame() %>%
  dplyr::rename(value = Freq, group = Var1) %>%
  dplyr::arrange(desc(group)) %>%
  dplyr::mutate(prop = round(value / sum(value) * 100, 1)) %>%
  dplyr::mutate(ypos = cumsum(prop) - 0.5 * prop )

lti_ccre_plot <- pie_ggplot(lti_trans_eqtl_ccre)
ggsave(filename = paste0(my_path, "results/figures/pie-ccre-lti-trans-eqtl.png"),
       plot = lti_ccre_plot,
       dpi = 320)

```


### Figure 2B


### Polygenic score?
```{r fig2b_prepqtl}
## CRTH2 = Ptgdr2,  TCF-1 = Tcf7, CKIT = KIT, CGRP = Calca
key_genes <- c("Tbx21", "Gata3", "Rorc", "Ncr1", "Areg", 
               "Ptgdr2", "Gfi1", "Tcf7", "Ahr", "Kit", "Calca")

tf_df <- data.frame(index = obj@assays$RNA@data@Dimnames[[2]],
           t(as.matrix(obj@assays$RNA@data[key_genes,]))) %>%
  # mutate(Tbet_rankz = DOQTL::rankZ(Tbx21),
  #        Gata3_rankz = DOQTL::rankZ(Gata3),
  #        Rorc_rankz = DOQTL::rankZ(Rorc),
  #        Ncr1_rankz = DOQTL::rankZ(Ncr1)) %>%
  left_join({ 
    obs %>% dplyr::select( index,
                           louvain_labels, 
                           called_cell_types, 
                           called_cell_types_new, 
                           BestCall,
                           filtered_set )
    })


pheno <- tf_df %>%
  dplyr::filter(called_cell_types %in% c("Lti ILC3", "ILC1", "ILC2", "NCR1+ ILC3")) %>%
  dplyr::filter(filtered_set == 1, BestCall !="AMB", BestCall != "") %>%
  dplyr::select(BestCall, called_cell_types, Tbx21:Kit) %>%
  dplyr::group_by(BestCall, called_cell_types) %>%
  dplyr::summarise_all(mean) %>%
  tidyr::pivot_longer(Tbx21:Kit) %>%
  tidyr::unite("name", called_cell_types:name, remove = TRUE) %>%
  tidyr::pivot_wider( names_from =  name, 
                      values_from = value) %>%
  as.data.frame



marker_file <- paste0(project_path, "genotype/GM_snps.Rdata")
marker_map_file <- paste0(project_path, "genotype/Regev_map_20171221.rds")
gene_map_file <- paste0(my_path, "data/reference/ensembl.Mus_musculus.GRCm38.93.csv")
cores <- 2

## Load Genotypes (Allele Probs)
cat("Reading allele probs data ...", "\n")
load(paste0(geno_path, "alleleprobs.Rdata"))
cat("Loaded allele probs data ...", "\n")
nms <- rownames(aprobs[[1]])

## subbset to DO mice for single cell data
samples <- grep("20180706", nms, value = T)
for( chr in names(aprobs) ) {
  aprobs[[chr]] <- aprobs[[chr]][samples, , ]
}


nms <- rownames(aprobs[[1]])
nms <- str_extract(nms, "C[0-9]{1,2}_[0-9]")

## rename sample ids to match phenotype data
for( chr in names(aprobs) ) {
  rownames(aprobs[[chr]]) <- nms
}

## pheno <- ilc2 %>% group_by(BestCall)  %>% summarise_all(funs(mean)) %>% as.data.frame()
## pheno <- topics %>% group_by(BestCall)  %>% summarise_all(funs(mean)) %>% as.data.frame()
## Load Phenotypes (all females)
cat("Reading phenotype data ...", "\n")
# pheno <- fread(pheno_file, data.table = FALSE)
rownames(pheno) <- pheno[,1]
pheno <- pheno[,-1]


## Sync sample names
cat("Sync sample names ...", "\n")
samples <- intersect(rownames(pheno), rownames(aprobs[[1]]))

pheno <- pheno[samples, ,drop = FALSE]
for( chr in names(aprobs) ) {
  aprobs[[chr]] <- aprobs[[chr]][samples, , ]
}


# Creating kinship matrix
cat("Creating kinship matrix ... \n")
K <- calc_kinship(probs = aprobs, type = "loco", cores = cores)


```

```{r fig2b_runqtl}
cat("Running QTL analysis...", "\n")
qtl <- scan1(genoprobs = aprobs,
             pheno = pheno, # [, var_col, drop = FALSE],
             kinship = K,
             cores = cores)
```

```{r fig2b_perm_test}
var_col <- "Tbet_rankz"
operm <- scan1perm(genoprobs = aprobs[, '2'],
             pheno = pheno[, var_col, drop = FALSE],
             kinship = K[['2']],
             cores = cores,
             n_perm = 100)
```



```{r plot_perm}

for( gene in colnames(qtl) ) {
  png(paste0("qtl-eqtl-", gene, "-gwa.png"))

  ## var_col <- paste0("topic", topic_num)
  ## thr <- quantile(operm, 0.95)
  thr <- 6
  plot(x = qtl, map = marker_map, lodcolumn = gene, main = gene)
  abline(h = thr, col = 'red')

  dev.off()

}

lods <- as.data.frame(qtl)
lods$marker <- rownames(lods)
write.csv(lods, "lods-eqtl-ilc-tfgenes.csv")
```



```{r fig2b_varex}
library(QTLRel)

# pheno
AA <- calc_kinship(probs = aprobs, type = "overall", cores = cores)
v <- list(A = AA)

gdtmp <- geno[,-1] + 1
rownames(gdtmp) <- geno[,1]
gdtmp <- replace(gdtmp, is.na(gdtmp), 0)
ii <- match(rownames(pheno), rownames(gdtmp))
gdtmp <- gdtmp[ii, ]


gmap <- GM_snps %>% 
  dplyr::select(snp = marker, chr = chr, dist = cM, pos = pos) %>%
  dplyr::mutate(pos = pos * 1e6)
gmap <- gmap[!{duplicated(gmap$pos)}, ]
gmap <- gmap[!{duplicated(gmap$dist)}, ]

keep_markers <- intersect(colnames(gdtmp), gmap$snp)
gdtmp <- gdtmp[, keep_markers]
gmap <- gmap %>% dplyr::filter(snp %in% keep_markers)

# rung 'genoProb'
prDat<- genoProb(gdat = gdtmp, 
                 gmap = gmap, 
                 step = Inf,
                 gr = 8,
                 pos = NULL,
                 method = "Haldane", 
                 msg = TRUE)


scanone_list <- sapply( colnames(pheno), function(gene) {
  
  # estimate variance components
  o <- tryCatch({
    estVC(y = pheno[[gene]], v = v)
    }, error = function(e) NULL)
  
  if( !is.null(o) ) {
    
    # genome scan
    out <- scanOne(y = pheno[[gene]], prdat = prDat, vc = o, test = "LRT")
# }, error = function(e) list(parameters = data.frame(Intercept=0,a=0,d=0), pval = 0, v = 0))
      
    data.frame(out$parameters, 
               pval = out$pval, 
               lod = out$pval/(2*log(10)),
               v = out$v) %>%
      dplyr::rename_all( .funs = list(~paste0(., "_", gene))) %>%
      tibble::rownames_to_column(var = "marker")
    }
    
}, USE.NAMES = TRUE, simplify = FALSE)


# Get the LOD scores and the proportion of variance explained by each SNP. 
# gwscan$lod <- out$p/(2*log(10))
# gwscan$pve <- out$v/100
  
# run 'qtlVar'
qtlvar_out <- Map(function(l) {
  
  qef <- as.data.frame(l$parameters[, c("a", "d")])
  qtlVar(qef, prDat$pr)

}, scanone_list)
# ## End(Not run)

qtlvar_df <- Reduce(merge, scanone_list)
write.csv(qtlvar_df, paste0(my_path, "results/variance-explained-rel-eqtl.csv"))
```


```{r}
tf_lods <- read_csv(paste0(my_path, "results/lods-eqtl-ilc-tfgenes.csv"))
tf_var_explained <- read_csv(paste0(my_path, "results/variance-explained-eqtl.csv"))
```

```{r}

## a: -1, 0, 1    d: 0, 1
check <- as.matrix(gdtmp) %*% as.matrix(qtlvar_df[, 3:4])

phenonames <- colnames(pheno)

polygenic_effect <- sapply(phenonames, function(p) {
  
  sapply(colnames(gdtmp), function(g) {
    
    (gdtmp[,g] - 2) * qtlvar_df[g, paste0("a_", p)] + 
      as.numeric(gdtmp[, g] == 2) * qtlvar_df[g, paste0("d_", p)]
    
    }, USE.NAMES = TRUE, simplify = FALSE)
  
}, USE.NAMES = TRUE, simplify = FALSE)
```


```{r}

## a: -1, 0, 1    d: 0, 1
check <- as.matrix(gdtmp) %*% as.matrix(qtlvar_df[, 3:4])

phenonames <- colnames(pheno)

polygenic_effect <- sapply(phenonames, function(p) {
  
  lapply(rownames(gdtmp), function(m) {
    
    aeffect <- qtlvar_df[, paste0("a_", p), drop = FALSE] %*% as.numeric(gdtmp[m, ] - 2)
    deffect <- qtlvar_df[, paste0("d_", p), drop = FALSE] %*% as.numeric(gdtmp[m, ] == 2)
    
    }) %>%
    do.call(., rbind)
  
}, USE.NAMES = TRUE, simplify = FALSE)
```


```{r}
## ILC1
# number of eQTLs per gene
ilc1_eqtl_gene_count <- sort(sapply(split(ilc1_mean, ilc1_mean$gene), function(d) length(unique(d$ensembl_gene))))
ilc1_eqtl_gene_count <- sort(sapply(split(ilc1_cv, ilc1_cv$gene), function(d) length(unique(d$ensembl_gene))))


## ILC2
ilc2_eqtl_gene_count <- sort(sapply(split(ilc2_mean, ilc2_mean$gene), function(d) length(unique(d$ensembl_gene))))
ilc2_eqtl_gene_count <- sort(sapply(split(ilc2_cv, ilc2_cv$gene), function(d) length(unique(d$ensembl_gene))))


## ILC3
ilc3_eqtl_gene_count <- sort(sapply(split(ilc3_mean, ilc3_mean$gene), function(d) length(unique(d$ensembl_gene))))
ilc3_eqtl_gene_count <- sort(sapply(split(ilc3_cv, ilc3_cv$gene), function(d) length(unique(d$ensembl_gene))))


## LTi
lti_eqtl_gene_count <- sort(sapply(split(lti_mean, lti_mean$gene), function(d) length(unique(d$ensembl_gene))))
lti_eqtl_gene_count <- sort(sapply(split(lti_cv, lti_cv$gene), function(d) length(unique(d$ensembl_gene))))


ilc1_eqtl_genecount_df <- data.frame(ilc1_eqtl_gene_count) %>% 
  tibble::rownames_to_column(var = "gene")
ilc2_eqtl_genecount_df <- data.frame(ilc2_eqtl_gene_count) %>% 
  tibble::rownames_to_column(var = "gene")
ilc3_eqtl_genecount_df <- data.frame(ilc3_eqtl_gene_count) %>% 
  tibble::rownames_to_column(var = "gene")
lti_eqtl_genecount_df <- data.frame(lti_eqtl_gene_count) %>% 
  tibble::rownames_to_column(var = "gene")


eqtl_genecount_df <- ilc1_eqtl_genecount_df %>%
  dplyr::full_join(ilc2_eqtl_genecount_df) %>%
  dplyr::full_join(ilc3_eqtl_genecount_df) %>%
  dplyr::full_join(lti_eqtl_genecount_df) %>%
  tidyr::replace_na(replace = list(ilc1_eqtl_genecount_df = 0,
                                   ilc2_eqtl_genecount_df = 0,
                                   ilc3_eqtl_genecount_df = 0,
                                   lti_eqtl_genecount_df = 0))

```


```{r}
## ILC1
# number of eGenes per eQTL gene
ilc1_egene_count <- sort(sapply(split(ilc1_mean, ilc1_mean$ensembl_gene), function(d) length(unique(d$gene))))
ilc1_egene_count <- sort(sapply(split(ilc1_cv, ilc1_cv$ensembl_gene), function(d) length(unique(d$gene))))


## ILC2
ilc2_egene_count <- sort(sapply(split(ilc2_mean, ilc2_mean$ensembl_gene), function(d) length(unique(d$gene))))
ilc2_egene_count <- sort(sapply(split(ilc2_cv, ilc2_cv$ensembl_gene), function(d) length(unique(d$gene))))


## ILC3
ilc3_egene_count <- sort(sapply(split(ilc3_mean, ilc3_mean$ensembl_gene), function(d) length(unique(d$gene))))
ilc3_egene_count <- sort(sapply(split(ilc3_cv, ilc3_cv$ensembl_gene), function(d) length(unique(d$gene))))


## LTi
lti_egene_count <- sort(sapply(split(lti_mean, lti_mean$ensembl_gene), function(d) length(unique(d$gene))))
lti_egene_count <- sort(sapply(split(lti_cv, lti_cv$ensembl_gene), function(d) length(unique(d$gene))))


ilc1_egenecount_df <- data.frame(ilc1_egene_count) %>% 
  tibble::rownames_to_column(var = "gene")
ilc2_egenecount_df <- data.frame(ilc2_egene_count) %>% 
  tibble::rownames_to_column(var = "gene")
ilc3_egenecount_df <- data.frame(ilc3_egene_count) %>% 
  tibble::rownames_to_column(var = "gene")
lti_egenecount_df <- data.frame(lti_egene_count) %>% 
  tibble::rownames_to_column(var = "gene")


egenecount_df <- ilc1_egenecount_df %>%
  dplyr::full_join(ilc2_egenecount_df) %>%
  dplyr::full_join(ilc3_egenecount_df) %>%
  dplyr::full_join(lti_egenecount_df) %>%
  tidyr::replace_na(replace = list(ilc1_egene_count = 0,
                                   ilc2_egene_count = 0,
                                   ilc3_egene_count = 0,
                                   lti_egene_count = 0))

```


### Figure 2C


#### Ven Diagram


```{r}
myCol <- RColorBrewer::brewer.pal(4, "Pastel2")
# Chart
venn.diagram(
        x = list(na.omit(vars$index[vars$ilc1_egenes_cv == 1]),
                 na.omit(vars$index[vars$ilc3_egenes_cv == 1]),
                 na.omit(vars$index[vars$ilc1_se == 1]),
                 na.omit(vars$index[vars$ilc3_se == 1])),
        category.names = c("ILC1 eGenes" , "ILC3 eGenes" , "ILC1 SE", "ILC3 SE"),
        filename = 'egenes-overlap-segenes-venn-diagramm.png',
        output=TRUE,
        main = "Super Enhancer Overlap with eGenes",
        main.cex = 0.5,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 480 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .5,
        fontface = "bold",
        fontfamily = "sans",
        cat.cex = .25
        
)
```


Overlap with ILC genes found in this paper (Figure 4D)
https://pubmed.ncbi.nlm.nih.gov/27156452/


```{r fig2c}
library(ggseqlogo)
library(Gviz)
library(GenomicRanges)
library(GenomicFeatures)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(BSgenome.Mmusculus.UCSC.mm10)
library(RMariaDB)

## Rbpj SNPs - chr5 53553396 - 53661165
start_irange <- 53000000
end_irange <- 54000000
chr_str <- "chr5"
chr_num <- "5"
gen <- "mm10"

strack <- SequenceTrack(Mmusculus, chromosome = chr_str)

txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode = "coarse")

gff_file <- paste0(my_path, "data/reference/Mus_musculus.GRCm38.100.gff3.gz")
gff <- readGFF(gff_file)
rbpj <- subset(gff, Parent=="gene:ENSMUSG00000039191")
gff_gr <- readGFFAsGRanges(gff_file)
txdb <- makeTxDbFromGFF(gff_file, format="gff3")
byexons <- exonsBy(txdb, by="gene")

# txdb <- makeTxDbFromBiomart(biomart="ensembl", dataset = "mmusculus_gene_ensembl")
# keepStandardChromosomes(txdb, pruning.mode="coarse")
# 
# txdb <- makeTxDbFromEnsembl("Mus musculus", server="useastdb.ensembl.org")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)


grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

start_lti_gwas <- end_lti_gwas <- lti_gwas %>%
  arrange(pos) %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_lti_gwas <- lti_gwas %>%
  arrange(pos) %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% lods

lti_gwas_gr <- GRanges(
  seqnames = Rle(chr_str),
  ranges = IRanges(start = start_lti_gwas,
                   end = end_lti_gwas)
  )
##data track
dtrack <- DataTrack(data = data_lti_gwas, 
                    start = start_lti_gwas,
                    end = end_lti_gwas+1, 
                    chromosome = chr, 
                    genome = gen,
                    name = "Uniform")

#highlight
# ht <- HighlightTrack(trackList = list(atrack, grtrack,dtrack),
#                      start = c(26705000, 26720000), width = 7000, chromosome = 7)


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack),
           from = start_irange, to = end_irange, cex = 0.8)


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")
```


### Rbpj

```{r}
## Rbpj SNPs - chr5 53553396 - 53661165
start_irange <- 53000000
end_irange <- 54000000
chr_str <- "chr5"
chr_num <- "5"
gen <- "mm10"

lti_gwas <- fread(paste0(my_path, "results/LTi_stressed_vs_non_qtl.csv"))
ilc2_ilc3_gwas <- fread(paste0(my_path, 'results/gwas-ilc2-ilc3-results.csv'))
ilc2_lti_gwas <- fread(paste0(my_path, 'results/gwas-ilc2-lti-results.csv'))


start_lti_gwas <- end_lti_gwas <- lti_gwas %>%
  arrange(pos) %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_lti_gwas <- lti_gwas %>%
  arrange(pos) %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% lods

lti_dtrack <- DataTrack(data = data_lti_gwas, 
                    start = start_lti_gwas,
                    end = end_lti_gwas+1, 
                    chromosome = chr_num, 
                    genome = gen,
                    name = "LTi Activated")


start_ilc2_ilc3_gwas <- end_ilc2_ilc3_gwas <- ilc2_ilc3_gwas %>%
  arrange(pos) %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_ilc2_ilc3_gwas <- ilc2_ilc3_gwas %>%
  arrange(pos) %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% lods

ilc2_ilc3_dtrack <- DataTrack(data = data_ilc2_ilc3_gwas, 
                    start = start_ilc2_ilc3_gwas,
                    end = end_ilc2_ilc3_gwas, 
                    chromosome = chr_num, 
                    genome = gen,
                    name = "ILC2 vs ILC3")



start_ilc2_lti_gwas <- end_ilc2_lti_gwas <- ilc2_lti_gwas %>%
  arrange(pos) %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_ilc2_lti_gwas <- ilc2_lti_gwas %>%
  arrange(pos) %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% lods

ilc2_lti_dtrack <- DataTrack(data = data_ilc2_lti_gwas, 
                    start = start_ilc2_lti_gwas,
                    end = end_ilc2_lti_gwas, 
                    chromosome = chr_num, 
                    genome = gen,
                    name = "ILC2 vs LTi-Like")

plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack,
                lti_dtrack, ilc2_ilc3_dtrack, ilc2_lti_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")
```



### STIM2

```{r stim2_gviz}
# LTi Activated
## Stim2 SNPs - chr5 53553396 - 53661165
start_irange <- 53500000
end_irange <- 54500000
chr_str <- "chr5"
chr_num <- "5"
gen <- "mm10"

## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

lti_gwas <- fread(paste0(my_path, "results/LTi_stressed_vs_non_qtl.csv"))
ilc2_ilc3_gwas <- fread(paste0(my_path, 'results/gwas-ilc2-ilc3-results.csv'))
ilc2_lti_gwas <- fread(paste0(my_path, 'results/gwas-ilc2-lti-results.csv'))

start_lti_gwas <- end_lti_gwas <- lti_gwas %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_lti_gwas <- lti_gwas %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% lods

lti_dtrack <- DataTrack(data = data_lti_gwas, 
                    start = start_lti_gwas,
                    end = end_lti_gwas+1, 
                    chromosome = chr_num, 
                    genome = gen,
                    name = "LTi Activated")


start_ilc2_ilc3_gwas <- end_ilc2_ilc3_gwas <- ilc2_ilc3_gwas %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_ilc2_ilc3_gwas <- ilc2_ilc3_gwas %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% lods

ilc2_ilc3_dtrack <- DataTrack(data = data_ilc2_ilc3_gwas, 
                    start = start_ilc2_ilc3_gwas,
                    end = end_ilc2_ilc3_gwas, 
                    chromosome = chr_num, 
                    genome = gen,
                    name = "ILC2 vs ILC3")



start_ilc2_lti_gwas <- end_ilc2_lti_gwas <- ilc2_lti_gwas %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_ilc2_lti_gwas <- ilc2_lti_gwas %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% lods

ilc2_lti_dtrack <- DataTrack(data = data_ilc2_lti_gwas, 
                    start = start_ilc2_lti_gwas,
                    end = end_ilc2_lti_gwas, 
                    chromosome = chr_num, 
                    genome = gen,
                    name = "ILC2 vs LTi-Like")

plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack,
                lti_dtrack, ilc2_ilc3_dtrack, ilc2_lti_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")

```




### Slc39a10


```{r slc39a10_gviz}
# ILC3 Activated
## Slc39a10 SNPs - chr1 46807544 - 46892852
start_irange <- 46500000
end_irange <- 47500000
chr_str <- "chr1"
chr_num <- "1"
gen <- "mm10"

## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

ilc3_gwas <- fread(paste0(my_path, "results/ILC3_stressed_vs_non_qtl.csv"))
topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv")) 

start_ilc3_gwas <- end_ilc3_gwas <- ilc3_gwas %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_ilc3_gwas <- ilc3_gwas %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% lods

ilc3_dtrack <- DataTrack(data = data_ilc3_gwas, 
                    start = start_ilc3_gwas,
                    end = end_ilc3_gwas+1, 
                    chromosome = chr_num, 
                    genome = gen,
                    name = "ILC3 Activated")



plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, ilc3_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")

```

### Sstr4

```{r sstr_gviz}
## topic 3
# topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))
## Sstr4 - Chr2:148395377-148396764
start_irange <- 148000000
end_irange <- 149000000
chr_str <- "chr2"
chr_num <- "2"
gen <- "mm10"


## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

## topic1, topic3, topic6
topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))

start_topic3 <- end_topic3 <- topics %>%
  dplyr::left_join(ccre, by = "marker") %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_topic3 <- topics %>%
  dplyr::left_join(ccre, by = "marker") %>%
  dplyr::arrange(pos) %>% 
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %>%
  dplyr::select(topic1, topic3) 
  # dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% topic3

topic3_dtrack <- DataTrack(data = data_topic3,
                    start = start_topic3,
                    end = end_topic3,
                    chromosome = chr_num,
                    groups = c("Topic 1", "Topic 3"),
                    genome = gen,
                    name = "Topic")


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, 
                topic3_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")

```


### Sox9

```{r sox9_gviz}
## topic 8
# topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))
## Sox9 - Chr11:112782224-112787760
start_irange <- 112500000
end_irange <- 113250000
chr_str <- "chr11"
chr_num <- "11"
gen <- "mm10"


## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

## topic8
topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))

start_topic <- end_topic <- topics %>%
  dplyr::left_join(ccre, by = "marker") %>%
  dplyr::arrange(pos) %>%
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos

data_topic8 <- topics %>%
  dplyr::left_join(ccre, by = "marker") %>%
  dplyr::arrange(pos) %>% 
  dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% topic8

topic_dtrack <- DataTrack(data = data_topic8,
                    start = start_topic,
                    end = end_topic,
                    chromosome = chr_num,
                    genome = gen,
                    name = "Topic 8")


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, 
                topic_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")


```



### NLRP12
(ILC1-topic0)

Many NLRs play inhibitory roles in regulating the signaling of TLRs including NLRP12 which targets MAPK and NF-B activation contribute to a cross-linked and finely tuned network of PRR signaling in response to the large repertoire of PAMPs and ensures the most effective and proper outcomes of innate immune responses(Liu and Cao, 2016). NLRP12 is a checkpoint of noncanonical NF-B, inflammation and tumorigenesis suppressing Colon Inflammation and Tumorigenesis through the Negative Regulation of Non-canonical NF-B(Allen et al., 2012). Heightened inflammatory response of IECs was associated with significant intrinsic activation of NF-B in IECs and increased frequency of ILC1, ILC3, and myeloid osteoclast progenitors. IEC-specific NF-B activation plays the central role in the cellular, inflammatory, and osteolytic phenotypes observed in the chemically induced colitis and conditional deletion of IKK2 from IECs significantly attenuated inflammation and bone loss in DSS-induced colitis(Ke et al., 2019). Further, while the dysregulation of IEC could influence ILC1 directly or  promote phagocytes to produce cytokines like IFNg to stimulate ILC1 development and function(n.d.). Our observation pinpoints the IEC-ILC1 axis in intestinal inflammation

```{r nlrp12_gviz}
## cis-eQTL for all major lineages
## Nlrp12 - Chr7:3218784-3249740  ENSMUSG00000078817 
start_irange <- 3218784
end_irange <- 3249740
chr_str <- "chr7"
chr_num <- "7"
gen <- "mm10"


## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

# ## topic8
# topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))
# 
# start_topic <- end_topic <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>%
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
# 
# data_topic8 <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>% 
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% topic8
# 
# topic_dtrack <- DataTrack(data = data_topic8,
#                     start = start_topic,
#                     end = end_topic,
#                     chromosome = chr_num,
#                     genome = gen,
#                     name = "Topic 8")


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, 
                topic_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")


```


### SOCS2


### CERS6


### TMEM132c

5 proteins of the human TMEM132 family (TMEM132A, B, C, D and E). These are genes in which variants are enriched for individuals with hearing loss, panic disorder or cancer. Common variants within the TMEM132E gene are associated with insomnia symptoms(Li et al., 2014) common and rare variants near TMEM132D gene are robustly associated with panic disorder(Haaker et al., 2014); and variants near TMEM132B are associated with excessive daytime sleepiness (Lane et al., 2017). In healthy individuals, some of the TMEM132D non-coding variants exhibit higher anxiety scores and larger volumetric estimates of the amygdala and hippocampus, key neural structures associated with fear and anxiety(Haaker et al., 2014). In the mouse, anterior cingulate cortex TMEM132D expression correlates with anxiety-related behavior (Erhardt et al., 2010). Finally, mutations in TMEM132D are unusually frequent in small-cell lung cancer(Iwakawa et al., 2015)and in pancreatic cancer(Forbes et al., 2014). Disease mutations near, or functions of, TMEM132A or C have yet to be identified. Recent study has shown TMEM132c is highly expressed in human colorectal cancer. Here we found TEME132c is a SNP-gene with eQTL effect associated with ILC1/ILC3 relative composition in mouse and its expression correlates with
Suggest the potential involvement of TEME protein in mediating the disease susceptibility to host defense and inflammation.


```{r tmem132c_gviz}
## cis-eQTL for all major lineages
## Tmem132c Chr5:127241826-127565790  ENSMUSG00000034324 
start_irange <- 127000000
end_irange <- 128000000
chr_str <- "chr5"
chr_num <- "5"
gen <- "mm10"

ilc1_ilc3_gwas <- fread(paste0(my_path, 'results/gwas-ilc2-ilc3-results.csv'))

## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

# ## topic8
# topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))
# 
# start_topic <- end_topic <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>%
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
# 
# data_topic8 <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>% 
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% topic8
# 
# topic_dtrack <- DataTrack(data = data_topic8,
#                     start = start_topic,
#                     end = end_topic,
#                     chromosome = chr_num,
#                     genome = gen,
#                     name = "Topic 8")


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, 
                topic_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")


```



### Lgr6

the 7-transmembrane receptor, Lgr5 has recently gained prominence as a marker of Wnt-regulated adult stem cell populations in the hair-follicle, intestine and stomach. A closely-related protein, Lgr6 marks adult stem cells responsible for fueling the renewal of the sebaceous gland and skin(Leushacke and Barker, 2011). Lgr6 as a marker of adult stem cells in the skin(Barker et al., 2013). Lgr6 is expressed by airway smooth muscle cells (aSMCs) lining the bronchiolar epithelium. Secretion of Fgf10 by Lgr6 + aSMCs promotes differentiation of airway epithelial cells(Leung et al., 2018).While Igr6 is well implicated in colon cancer as an intestinal stem cell marker. We found the eQTL for this gene implicated here with the relative proportion of ILC1/ILC3 which further emphasize the importance of epithelium cell for the development of ILCs.


```{r lgr6_gviz}
## cis-eQTL for all major lineages
## Lgr6 Chr1:134983301-135105276  ENSMUSG00000042793 
start_irange <- 134983301
end_irange <- 135105276
chr_str <- "chr1"
chr_num <- "1"
gen <- "mm10"


## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

# ## topic8
# topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))
# 
# start_topic <- end_topic <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>%
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
# 
# data_topic8 <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>% 
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% topic8
# 
# topic_dtrack <- DataTrack(data = data_topic8,
#                     start = start_topic,
#                     end = end_topic,
#                     chromosome = chr_num,
#                     genome = gen,
#                     name = "Topic 8")


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, 
                topic_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")


```



### Rrbp1

Rrbp1 is highly expressed in epithelium cells in small intestine. Ribosome-binding protein 1 (RRBP1) has been implicated in the regulation of unfolded protein response, which is involved in almost every aspect of cancer development. Further, Endoplasmic reticulum ribosome-binding protein 1, RRBP1, promotes progression of colorectal cancer and predicts an unfavourable prognosis(Pan et al., 2015). over the last decade have shown that the UPR plays a critical role in shaping immunity and inflammation, resulting in the recognition of the UPR as a key player in pathological
processes including complex inflammatory, autoimmune and neoplastic diseases. Intestinal epithelium, with its many highly secretory cells, forms an important barrier and
messenger between the luminal environment and the host immune system. It is not
surprising, that numerous studies have associated ER stress and the UPR with intestinal diseases such as inflammatory bowel disease (IBD) and colorectal cancer (CRC). The UPR in immune cells including B cell, T cells, macrophage and epithelium cells, all of which has great crosstalk with innate lymphoid cells(Coleman and Haller, 2019). We detected the eQTLs for Rrbp1 gene which accounts for ILC1/ILC3 composition. Our observation again emphasized the significance of UPR in regulating innate immune response in terms of ILC1 and ILC3 composition in small intestine.


```{r rrbp1_gviz}
## cis-eQTL for all major lineages
## Rrbp1 Chr2:143947395-144011263  ENSMUSG00000027422 
start_irange <- 143947395
end_irange <- 144011263
chr_str <- "chr2"
chr_num <- "2"
gen <- "mm10"


## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

# ## topic8
# topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))
# 
# start_topic <- end_topic <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>%
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
# 
# data_topic8 <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>% 
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% topic8
# 
# topic_dtrack <- DataTrack(data = data_topic8,
#                     start = start_topic,
#                     end = end_topic,
#                     chromosome = chr_num,
#                     genome = gen,
#                     name = "Topic 8")


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, 
                topic_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")


```



### IL-20RA

The interleukin-20-receptor I complex (IL-20-RI) is composed of two chains, IL20RA and IL20RB. Its ligands are the three members of the IL19 subfamily of cytokines, IL-19, IL-20 and IL-24. These cytokines are important in the manifestation of psoriatic lesions and, recently, an association of polymorphisms of IL20 with psoriasis has been described(Kingo et al., 2008). Capture Hi-C Identifies a Novel Causal Gene, IL20RA, in the Pan-Autoimmune Genetic Susceptibility Region 6q23. Interleukin 20 receptor, alpha subunit is a subunit for the interleukin-20 receptor. The risk allele of the most likely causal SNP, rs6927172, is correlated with both a higher frequency of interactions and increased expression of IL20RA, along with a stronger binding of both the NFB transcription factor and chromatin marks characteristic of active enhancers in T-cells(McGovern et al., 2016). Here, we found the eQTL for IL-20RA is responsible for distinguish between ILC3 and LTi-like ILC3.

```{r il20ra_gviz}
## cis-eQTL for all major lineages
## Il20RA Chr10:19712570-19760053  ENSMUSG00000020007 
start_irange <- 19712570
end_irange <- 19760053
chr_str <- "chr10"
chr_num <- "10"
gen <- "mm10"


## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

# ## topic8
# topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))
# 
# start_topic <- end_topic <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>%
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
# 
# data_topic8 <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>% 
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% topic8
# 
# topic_dtrack <- DataTrack(data = data_topic8,
#                     start = start_topic,
#                     end = end_topic,
#                     chromosome = chr_num,
#                     genome = gen,
#                     name = "Topic 8")


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, 
                topic_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")


```



### Fam21

WASH exists in a multiprotein complex containing FAM21, which links WASH to endosomes and is required for WASH-dependent retromer-mediated sorting(Gomez and Billadeau, 2009; Jia et al., 2012). WASH deletion impairs the cell pool of NKp46+ILC3s. In NKp46+ ILC3s, WASH recruits Arid1a to the Ahr promoter thus activating AHR expression, indicating WASH-mediated AHR expression has a critical function in the maintenance of NKp46+ILC3s(Xia et al., 2017). While NKp46+ is also expressed in ILC1 and here in our data we found eQTLs for FAM21 tips the balance of relative proportion of IL1/ILC2 indicating its role in regulating ILC development.

```{r fam21_gviz}
## cis-eQTL for all major lineages
## Washc2 (Fam21) Chr6:116208038-116262686  ENSMUSG00000024104 
start_irange <- 116208038
end_irange <- 116262686
chr_str <- "chr6"
chr_num <- "6"
gen <- "mm10"


## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

# ## topic8
# topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))
# 
# start_topic <- end_topic <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>%
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
# 
# data_topic8 <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>% 
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% topic8
# 
# topic_dtrack <- DataTrack(data = data_topic8,
#                     start = start_topic,
#                     end = end_topic,
#                     chromosome = chr_num,
#                     genome = gen,
#                     name = "Topic 8")


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, 
                topic_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")


```




### BANF2

BANF2 is termed for Barrier to autointegration factor 2, which has been associated with the abnormality of immune system physiology phenotype in GWAS datasets from the GWASdb SNP-Phenotype Associations dataset. This immune disease implicated SNP-gene is here to identified to be a eQTL-gene which influence the ILC development by tipping the balance of ILC2/ILC3 relative proportion.

```{r banf2_gviz}
## cis-eQTL for all major lineages
## Banf2 Chr2:144033102-144073979  ENSMUSG00000037307 
start_irange <- 144033102
end_irange <- 144073979
chr_str <- "chr2"
chr_num <- "2"
gen <- "mm10"


## Sequence and Ideogram
strack <- SequenceTrack(Mmusculus, chromosome = chr_str)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")
gtrack <- GenomeAxisTrack()
itrack <- IdeogramTrack(genome = gen, chromosome = chr_str)

## Genome Track
grtrack <- GeneRegionTrack(txdb, 
                           genome = gen,
                           chromosome = chr_str, 
                           name = "Gene Model",
                           geneAnnotation = "symbol")

pos_snps <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
names_snps <-  ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% marker

snps_atrack <- AnnotationTrack(start = pos_snps,
                               width = rep(1, length(pos_snps)),
                               chromosome = chr_str,
                               group = names_snps,
                               genome = gen,
                               name = "SNPs")

start_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% start
end_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% end
names_ccre <- ccre %>%
  filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(type.y)) %$% type.y

ccre_atrack <- AnnotationTrack(start = start_ccre,
                               width = end_ccre - start_ccre,
                               chromosome = chr_str,
                               group = names_ccre,
                               genome = gen,
                               name = "cCREs")

# ## topic8
# topics <- fread(paste0(my_path, "results/topic-qtl-lods.csv"))
# 
# start_topic <- end_topic <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>%
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
# 
# data_topic8 <- topics %>%
#   dplyr::left_join(ccre, by = "marker") %>%
#   dplyr::arrange(pos) %>% 
#   dplyr::filter(chr == chr_num, between(pos, start_irange, end_irange), !is.na(pos)) %$% topic8
# 
# topic_dtrack <- DataTrack(data = data_topic8,
#                     start = start_topic,
#                     end = end_topic,
#                     chromosome = chr_num,
#                     genome = gen,
#                     name = "Topic 8")


plotTracks(list(itrack, gtrack, snps_atrack, grtrack, strack, ccre_atrack, 
                topic_dtrack),
           from = start_irange, to = end_irange, cex = 0.8, type = "b")


```



### GVIZ Function

```{r}

txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
keepStandardChromosomes(txdb, pruning.mode="coarse")


gene_gviz <- function( chr, start, end, genome = "mm10" ) {
  
  ## Reference Sequence Track
  strack <- SequenceTrack(Mmusculus, chromosome = chr)
  ## Genome Axis
  gtrack <- GenomeAxisTrack()
  ## ideogram
  itrack <- IdeogramTrack(genome = genome, chromosome = chr)
  
  ## SNPs
  pos_snps <- ccre$pos[ccre$pos >= start_irange & ccre$pos <= start]
  names_snps <- ccre$marker[ccre$pos >= start_irange & ccre$pos <= end]
  snps_gr <- GRanges(
    seqnames = Rle("chr5"),
    ranges = IRanges(start = pos_snps, 
                     end = pos_snps,
                     names = names_snps)
    )
  snps_atrack <- AnnotationTrack(snps_gr, name = "SNPs")
    
  ## cCREs
  start_ccre <- na.omit(ccre$start[ccre$pos >= start & ccre$pos <= end])
  end_ccre <- na.omit(ccre$end[ccre$pos >= start & ccre$pos <= end])
  names_ccre <- na.omit(ccre$type.y[ccre$pos >= start & ccre$pos <= end])
  ccre_gr <- GRanges(
    seqnames = Rle("chr5"),
    ranges = IRanges(start = start_ccre, 
                     end = end_ccre,
                     names = names_ccre)
    )
  ccre_atrack <- AnnotationTrack(ccre_gr, name = "cCREs")

  
#   start_lti_gwas <- end_lti_gwas <- lti_gwas %>%
#   arrange(pos) %>%
#   filter(chr == "5", between(pos, start_irange, end_irange), !is.na(pos)) %$% pos
# 
# data_lti_gwas <- lti_gwas %>%
#   arrange(pos) %>%
#   filter(chr == "5", between(pos, start_irange, end_irange), !is.na(pos)) %$% lods
# 
# lti_gwas_gr <- GRanges(
#   seqnames = Rle("chr5"),
#   ranges = IRanges(start = start_lti_gwas,
#                    end = end_lti_gwas)
#   )
# ##data track
# dtrack <- DataTrack(data = data_lti_gwas, 
#                     start = start_lti_gwas,
#                     end = end_lti_gwas+1, 
#                     chromosome = chr, 
#                     genome = gen,
#                     name = "Uniform")
}
```


### Figure 2D

Gata3 QTL comparison
Do similar for TBET (Tbx21)
DO Similar for Ror gamma-t (Rorc)

```{r}
gata3 <- my_plot(plot_df, "Gata3") + my_theme
tbet <- my_plot(plot_df, "Tbet") + my_theme
rorc <- my_plot(plot_df, "Rorc") + my_theme
eomes <- my_plot(plot_df, "Eomes") + my_theme

ggsave("tsne-gata3.png", gata3, dpi = 300)
ggsave("tsne-tbet.png", tbet, dpi = 300)
ggsave("tsne-rorc.png", rorc, dpi = 300)
ggsave("tsne-eomes.png", eomes, dpi = 300)
```



```{r fig2d}
df <- df %>%
  left_join({
    geno %>% 
      select(BestCall = SampleID, 
             UNC20336327, UNCHS036777, UNCHS014430, UNCHS022710, UNC20419420)
  })

## Gata3 most significant eQTL for Gata3 - UNCHS036777
## UNCHS014430 - topic11 
gata3 <- my_plot(df, "Gata3")
topic3 <- my_plot(df, "topic11")
gata3_biocarta <- my_plot(df, "BIOCARTA_GATA3_PATHWAY")

biocarta_gata3_quant <- plot_quantile(df, "BIOCARTA_GATA3_PATHWAY", "UNCHS014430")
topic11_quant <- plot_quantile(df, "topic11", "UNCHS014430")
gata3_quant <- plot_quantile(df, "Gata3", "UNCHS014430")

## Tbet significant eQTL - UNC20336327
tbet <- my_plot(df, "Tbx21")
plot_quantile(df, "Tbet", "UNC20336327")


## Rorc most significant eQTL - UNCHS022710, UNC20419420
rorc <- my_plot(df, "Rorc")
plot_quantile(df, "Rorc", "UNCHS022710")
## need cumplot by SNP level for significant eQTL/topic QTL 
```



## Figure 3

Ex: ILC3 contribute to the progression and aggravation of IBD while both IL-22 and IL-17, along with IFN-, are overexpressed by the dysregulation of NCR ILC3 or NCR+ ILC3 function and the bias of NCR+ ILC3 towards ILC1 as well as regulatory ILC dysfunction in the pathological state
 IL-25 has been reported to have the capability to give rise to ILC3-like IL-17 producers, although in nave mice or upon helminth infection, they appear to default to a more conventional and less plastic ILC2 phenotype
Ex: The genetic makeup of the host is an important determinant of the pattern of T cell differentiation. Some inbred strains of mice develop Th2 responses to the same microbes that stimulate Th1 differentiation in most other strains. Strains of mice that develop Th2-dominant responses are susceptible to infections by intracellular microbes. It is possible, although not proven, that people differ in their propensity to mount Th1, Th2, or Th17 responses based on inherited genes
ILC2s and NKp44 ILC3s but not NKp44+ ILC3s are present in the peripheral blood and skin of healthy people, whereas the blood and inflamed skin of those with psoriasis contained NKp44+ ILC3s, indicating that dynamic changes in ILC population structure are associated with psoriatic inflammation
In the setting of malnutrition elicited by vitamin A deficiency in mice, an imbalance in ILC2 and ILC3 responses was evident, resulting in defective ILC3dependent antibacterial immunity but enhanced ILC2-dependent antihelminth immunity

Based on their transcriptional regulation and cytokine profiles, ILCs can be categorized into five subsets with defined phenotypes and functional profiles, but they also have the ability to adapt to local environmental cues by changing these profiles.


### Figure 3A

#### Cell Proportion Significant Loci
#### Ven Diagram Cell Proportion Significant Loci Genes


```{r fig3a_cellcomp_venn}
myCol <- RColorBrewer::brewer.pal(6, "Pastel2")
# Chart
# x = list(na.omit(vars$index[vars$ilc1_ilc2_prop_qtl_genes==1]),
#                  na.omit(vars$index[vars$ilc1_ilc3_prop_qtl_genes==1]),
#                  na.omit(vars$index[vars$ilc1_lti_prop_qtl_genes==1]),
#                  na.omit(vars$index[vars$ilc2_ilc3_prop_qtl_genes==1]),
#                  na.omit(vars$index[vars$ilc2_lti_prop_qtl_genes==1]),
#                  na.omit(vars$index[vars$ilc3_lti_prop_qtl_genes==1]))[1:5],
#         category.names = c("ILC1 vs ILC2" , "ILC1 vs ILC3 " , "ILC1 vs LTi", 
#                            "ILC2 vs ILC3", "ILC2 vs LTi", "ILC3 vs LTi")[1:5],
#         filename = 'cellcomposition-loci-genes-venn-diagramm.png',
#         main = "Cell Type Composition QTL",

upset(data = vars, 
      sets = c("ilc1_ilc2_prop_qtl_genes", "ilc1_ilc3_prop_qtl_genes", 
               "ilc1_lti_prop_qtl_genes", "ilc2_ilc3_prop_qtl_genes", 
               "ilc2_lti_prop_qtl_genes", "ilc3_lti_prop_qtl_genes"), 
      sets.bar.color = "#56B4E9",
      order.by = "freq", 
      empty.intersections = "on")
```


```{r}
upset(data = ccre, 
      sets = c("ilc3_activated_qtl", "lti_activated_qtl",
               "ilc1_ilc2_prop_qtl", "ilc1_lic3_prop_qtl",
               "ilc2_ilc3_prop_qtl", "ilc3_lti_prop_qtl"), 
      sets.bar.color = "#56B4E9",
      order.by = "freq", 
      empty.intersections = "on")
#      boxplot.summary = c("recomb_rate"))
```



Need to create a founder genotype data frame containing the significant marker genotypes for the significant loci affecting ILC proportions.  

```{r}
load(paste0(geno_path, 'genoprobs_rename.Rdata'))

## get significant markers here
prop_markers <- unique(c(ccre$marker[ccre$ilc1_ilc2_prop_qtl == 1],
                  ccre$marker[ccre$ilc1_ilc3_prop_qtl == 1],
                  ccre$marker[ccre$ilc1_lti_prop_qtl == 1],
                  ccre$marker[ccre$ilc2_ilc3_prop_qtl == 1],
                  ccre$marker[ccre$ilc2_lti_prop_qtl == 1],
                  ccre$marker[ccre$ilc3_lti_prop_qtl == 1]))

probs_markers <- sapply(names(probs), function(i) dimnames(probs[[i]])[[3]], simplify = T)
probs_markers <- unique(unlist(probs_markers))

prop_markers <- intersect(prop_markers, probs_markers)

samples <- grep("20180706", rownames(probs[[1]]), value = T)
nms <- str_extract(samples, "C[0-9]{1,2}_[0-9]")
cell_prop_geno_df <- data.frame(BestCall = nms)


for( m in prop_markers[4312:4586] ) {
  # i <- 1
  
  chr <- as.character(ccre$chr[ccre$marker == m])[1]
  
  cell_prop_geno <- round(apply(probs[[ chr ]][samples, , m], c(1,2), mean), 3)
  cell_prop_geno_df[[ m ]] <- colnames(cell_prop_geno)[apply(cell_prop_geno, 1, which.max)]
  # i <- i+1
  
}

write.csv(cell_prop_geno_df, "cell-proportions-qtl-genotypes.csv")
```


```{r}
## look at cCRE and cell_prop_geno_df to make plots for different proportions

# ccre
# plot_df

## get significant markers here
# prop_markers <- unique(c(ccre$marker[ccre$ilc1_ilc2_prop_qtl == 1],
#                   ccre$marker[ccre$ilc1_ilc3_prop_qtl == 1],
#                   ccre$marker[ccre$ilc1_lti_prop_qtl == 1],
#                   ccre$marker[ccre$ilc2_ilc3_prop_qtl == 1],
#                   ccre$marker[ccre$ilc2_lti_prop_qtl == 1],
#                   ccre$marker[ccre$ilc3_lti_prop_qtl == 1]))

cell_prop_geno_df <- read.csv(paste0(my_path, "results/cell-proportions-qtl-genotypes.csv"))


## Rbpj significant marker - UNCHS014402
## Stim2 significant marker - UNCHS014413
## Fam21 (Washc2) significant marker - UNCHS018751
## Rrbp1 significant marker - UNCHS007079
## Tmem132c - significant marker - UNCHS015849
## Il20ra - significant marker - UNC17535571

cell_prop_order <- plot_df %>%
  dplyr::left_join({ 
    cell_prop_geno_df %>% dplyr::select(BestCall, founder_genotype = UNCHS014413)
    }) %>%
  filter(louvain_labels %in% c(1,2,3,4,5,8), !is.na(founder_genotype)) %>% 
  group_by(founder_genotype, called_cell_types_new) %>% 
  summarise(n = n()) %>% 
  mutate(total = sum(n), prop = n / sum(n)) %>%
  filter(called_cell_types_new == "ILC3(LTi-like)") %>%
  arrange(prop) %$%
  # filter(called_cell_types_new == "ILC1") %>%
  # arrange(desc(prop)) %$%
  founder_genotype %>%
  as.character()

celltype_order <- c("NK", "ILC1", "ILC2", "ILC3", "ILC3(LTi-like)")
plot_df <- plot_df %>%
  dplyr::left_join({ 
    cell_prop_geno_df %>% dplyr::select(BestCall, founder_genotype = UNCHS014413)
    }) %>%
  dplyr::mutate(founder_genotype = factor(founder_genotype, levels = cell_prop_order),
         called_cell_types_new = factor(called_cell_types_new, levels = celltype_order))


plot_df %>% 
  filter(louvain_labels %in% c(1,2,3,4,5,8), !is.na(founder_genotype)) %>% 
  ggplot(., aes(founder_genotype, 
                color = called_cell_types_new, 
                fill = called_cell_types_new)) + 
  geom_bar(position = "fill") + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Cell Proportion at Marker UNCHS014413 (Stim2)")
```



### Figure 3B

```{r}

```


### Figure 3C

Super Enhancer Genes and relation to the eQTLs.

```{r ilc_se}
ilc_se <- read_csv(paste0(my_path, "ILC1-vs-ILC3-SE-NIHMS776743-supplement-9.csv"))
cnames <-  c("chr", "start", "end", "Symbol")
ilc1_se <- read_csv(paste0(my_path, "ilc1-se-genelist.csv"), 
                      col_names = cnames)
ilc3_se <- read_csv(paste0(my_path, "ilc3-se-genelist.csv"), 
                      col_names = cnames)

homologs <- read_delim(file = paste0(my_path, "data/reference/HOM_MouseHumanSequence.txt"), 
                       delim = "\t") %>%
  dplyr::select(`HomoloGene ID`, `Common Organism Name`, `Symbol`)
  # dplyr::distinct() %>%
  # pivot_wider(id_cols = "HomoloGene ID", 
  #             names_from = "Common Organism Name", 
  #             values_from = "Symbol")
mouse_hom <- homologs %>% dplyr::filter(`Common Organism Name` == "mouse, laboratory")
human_hom <- homologs %>% dplyr::filter(`Common Organism Name` == "human")

homologs_tab <- mouse_hom %>% left_join(human_hom, by = "HomoloGene ID")

ilc_se <- ilc_se %>%
  left_join({ homologs_tab %>% dplyr::select(Symbol = Symbol.y, index = Symbol.x) })

ilc1_se <- ilc1_se %>%
  left_join({ homologs_tab %>% dplyr::select(Symbol = Symbol.y, index = Symbol.x) })

ilc3_se <- ilc3_se %>%
  left_join({ homologs_tab %>% dplyr::select(Symbol = Symbol.y, index = Symbol.x) })

vars <- vars %>%
  dplyr::left_join({ ilc_se %>% select(index, ilc_se = `Cell Type`) }) %>%
  dplyr::mutate(ilc1_se = as.numeric(index %in% ilc1_se$index),
                ilc3_se = as.numeric(index %in% ilc3_se$index))


write.csv(vars, paste0(my_path, "data/scanpy/allchannels/vars.csv"))
```


```{r}

ilc1_se_egenes <- vars$index[vars$ilc1_egenes_cv == 1 & vars$ilc1_se==1]
ilc1_se_egenes <- vars$index[vars$ilc1_egenes_mean == 1 & vars$ilc1_se==1]

ilc3_se_egenes <- vars$index[vars$ilc3_egenes_cv == 1 & vars$ilc3_se==1]
ilc3_se_egenes <- vars$index[vars$ilc3_egenes_mean == 1 & vars$ilc3_se==1]

```

#### GO Enrichment

```{r go_enrich}
library(topGO)
library(GO.db)
library(biomaRt)
library(Rgraphviz)

bg_genes <- ensembl$gene

# create GO db for genes to be used using biomaRt - please note that this takes a while
# mmusculus_gene_ensembl, hsapiens_gene_ensembl
db <- useMart('ENSEMBL_MART_ENSEMBL', 
              dataset='mmusculus_gene_ensembl', 
              host="www.ensembl.org")
go_ids <- getBM(attributes=c('go_id', 
                             'external_gene_name', 
                             'ensembl_gene_id'), 
                filters='ensembl_gene_id', 
                values = bg_genes, 
                mart = db)

## candidate_list <- grep(pattern = "ENSMUSG", net_groups[[6]], value = TRUE)
## candidate_list <- vars$gene_ids[vars$ilc1_egenes_cv == 1 & vars$ilc1_se==1]
candidate_list <- vars$gene_ids[vars$ilc1_ilc2_prop_qtl_genes == 1]


# build the gene 2 GO annotation list (needed to create topGO object)
gene_2_GO <- unstack(go_ids[,c(1,4)])
affyLib <- 
# remove any candidate genes without GO annotation
keep <- candidate_list %in% go_ids[,4]
keep <- which(keep == TRUE)
candidate_list <- candidate_list[keep]
 
# make named factor showing which genes are of interest
geneList = factor(as.integer(bg_genes %in% candidate_list))
names(geneList) <- bg_genes

GOdata <- new("topGOdata", 
              description = "Simple session", 
              ontology = "BP",
              allGenes = geneList, 
              geneSel = topDiffGenes,
              nodeSize = 10,
              annot = annFUN.db, 
              affyLib = gene_2_GO)

classic_fisher_result <- runTest(GOdata, algorithm='classic', statistic='fisher')
# define test using the weight01 algorithm (default) with fisher
weight_fisher_result <- runTest(GOdata, algorithm='weight01', statistic='fisher') 
 
# generate a table of results: we can use the GenTable function to generate a summary table with the results from tests applied to the topGOdata object.
allGO <- usedGO(GOdata)
all_res <- GenTable(GOdata, 
                    weightFisher = weight_fisher_result, 
                    orderBy = 'weightFisher', 
                    topNodes = length(allGO))

```


```{r gsea_enrich}
# https://bioconductor.org/packages/release/bioc/vignettes/fgsea/inst/doc/fgsea-tutorial.html
library(fgsea) 
library(msigdbr)

# markers <- read.csv("tbet-nrc1-rorc-de-genes-ilc1-ilc3.csv")
# ranks <- -log10(markers$p_val_adj)[1:100]
# names(ranks) <- markers$X[1:100]

ranks <- rep(1, length(ilc1_se_egenes))
names(ranks) <- ilc1_se_egenes

m_df <- msigdbr(species = "Mus musculus", category = "H")
pathways <- m_df %>% split(x = .$gene_symbol, f = .$gs_name)
# pathways <- reactomePathways(names(exampleRanks))
# fgseaRes <- fgsea(pathways, exampleRanks, maxSize=500)
# head(fgseaRes)
# Run GSEA.
fgseaRes <- fgsea(pathways = pathways, 
                  stats = ranks, 
                  minSize = 10, maxSize = 500, nperm = 1000)
# Visualize GSEA results.
head(fgseaRes[order(pval), ])
sum(fgseaRes[, padj < 0.01])
plotEnrichment(examplePathways[["5990980_Cell_Cycle"]],
               exampleRanks) + labs(title="Cell Cycle")
topPathwaysUp <- fgseaRes[ES > 0][head(order(pval), n=10), pathway]
topPathwaysDown <- fgseaRes[ES < 0][head(order(pval), n=10), pathway]
topPathways <- c(topPathwaysUp, rev(topPathwaysDown))
plotGseaTable(examplePathways[topPathways], exampleRanks, fgseaRes, 
              gseaParam = 0.5)
```


```{r se_venn}
myCol <- RColorBrewer::brewer.pal(4, "Pastel2")
# Chart
venn.diagram(
        x = list(na.omit(vars$index[vars$ilc1_egenes_cv==1]),
                 na.omit(vars$index[vars$ilc1_se==1]),
                 na.omit(vars$index[vars$ilc3_egenes_cv==1]),
                 na.omit(vars$index[vars$ilc3_se==1])),
        category.names = c("ILC1 eGenes" , "ILC1 SE " , "ILC3 eGenes", "ILC3 SE"),
        filename = 'ilc1-ilc3-se-genes-venn-diagramm.png',
        output=TRUE,
        
        # Output features
        imagetype="png" ,
        height = 480 , 
        width = 480 , 
        resolution = 300,
        compression = "lzw",
        
        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,
        
        # Numbers
        cex = .5,
        fontface = "bold",
        fontfamily = "sans",
        cat.cex = .5
        
)
```


### Figure 3D

```{r}
marker_map <- readRDS(paste0(geno_path, "Regev_map_20171221.rds"))
xpos <- qtl2:::map_to_xpos(marker_map, gap = 25)
xchrbound <- qtl2:::map_to_boundaries(marker_map, gap = 25)

chr_breaks <- apply(xchrbound, 2, median)

rectangles <- data.frame(
  xmin = apply(xchrbound, 2, min),
  xmax = apply(xchrbound, 2, max),
  ymin = 0,
  ymax = max(lti_gwas$plot_lods) 
)[seq(2,20,2), ]

markers_df <- lti_gwas[chr == "5" & pos > 53000000 & pos < 54000000 & plot_lods > 10,]


gwas_plot <- ggplot() + 
  geom_rect(data = rectangles, 
            mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
            fill='gray80', alpha=0.8) +
  geom_point(data = lti_gwas, 
            mapping = aes(x = xpos, y = plot_lods), shape = 46) +
  geom_hline(yintercept = 11, color = "red", linetype = "dashed") + 
  geom_label(data = markers_df, aes(x = xpos, y = plot_lods, label = marker), size = 2) +
  scale_x_continuous(breaks = chr_breaks, label = c(as.character(1:19), "X")) + 
  labs(x = "Chromosome of SNP", y = "P-value(-log10 scale)", title = "Activated/Non-Activated LTi GWAS") +
  theme(
    axis.line=element_blank(),
    axis.ticks=element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    )

ggsave("plot-gwas-lti-stressed.png", gwas_plot, dpi = 300)
```



## Figure 4

### Figure 4A

```{r}

```



### Figure 4B

```{r}

```


### Figure 4C

```{r}

```


### Figure 4D

```{r}
ccr6 <- my_plot(plot_df, "Ccr6") + my_theme
gata3 <- my_plot(plot_df, "Gata3") + my_theme
tbet <- my_plot(plot_df, "Tbet") + my_theme
rorc <- my_plot(plot_df, "Rorc") + my_theme

topic3 <- my_plot(plot_df, "topic3") + my_theme + labs(title = "Topic 3")
topic8 <- my_plot(plot_df, "topic8") + my_theme + labs(title = "Topic 8")
topic10 <- my_plot(plot_df, "topic10") + my_theme + labs(title = "Topic 10")
topic11 <- my_plot(plot_df, "topic11") + my_theme + labs(title = "Topic 11")

ggsave("tsne-topic3.png", topic3, dpi = 300)
ggsave("tsne-topic8.png", topic8, dpi = 300)
ggsave("tsne-topic10.png", topic10, dpi = 300)
ggsave("tsne-topic11.png", topic11, dpi = 300)

## need to replot
topic3_qtl
topic8_qtl
topic10_qtl
topic11_qtl


```

```{r}
marker_map <- readRDS(paste0(geno_path, "Regev_map_20171221.rds"))
xpos <- qtl2:::map_to_xpos(marker_map, gap = 25)
xchrbound <- qtl2:::map_to_boundaries(marker_map, gap = 25)

## Make Marker Reference Dataset
MarkerRef <- data.frame(marker = names(xpos), xpos = xpos, stringsAsFactors = FALSE) %>%
  left_join(
    {conseq %>% select(marker, chr)}
  )

topic_lods <- lods %>% left_join(conseq) %>% left_join(MarkerRef)

rectangles <- data.frame(
  xmin = apply(xchrbound, 2, min),
  xmax = apply(xchrbound, 2, max),
  ymin = 0,
  ymax = 10 
)[seq(2,20,2), ]


chr2_markers <- conseq$marker[conseq$chr == "2"]
chr5_markers <- conseq$marker[conseq$chr == "5"]
chr11_markers <- conseq$marker[conseq$chr == "11"]

p1 <- ggplot() + 
  geom_rect(data = rectangles, 
            mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
            fill='gray80', alpha=0.8) +
  geom_line(data = topic_lods, 
            mapping = aes(x = xpos, y = topic3)) +
  geom_hline(yintercept = 6, color = "red", linetype = "dashed") + 
  scale_x_continuous(breaks = apply(xchrbound, 2, median), label = c(as.character(1:19), "X")) + labs(x = "Chromosome of SNP", y = "LOD", title = "Topic 3", ymin = 0, ymax = 8) +
  theme(
    axis.line=element_blank(),
    axis.ticks=element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    )


p2 <- ggplot() +
  geom_line(data = topic_lods[topic_lods$marker %in% chr2_markers,], 
            mapping = aes(x = xpos, y = topic3)) +
  geom_hline(yintercept = 6, color = "red", linetype = "dashed") + 
  scale_x_continuous(breaks = apply(xchrbound, 2, median), label = c(as.character(1:19), "X")) + 
  labs(x = "Chromosome of SNP", y = "LOD", title = "Topic 3") +
  ylim(ymin = 0, ymax = 8) + 
  theme(
    axis.line=element_blank(),
    axis.ticks=element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    )

ggsave("qtl-plot-topic3-chr2.png", p2, dpi = 300)

```



## Figure 5

### Figure 5A

```{r}
ilc1_cis_markers <- ilc1_cv %>% 
  dplyr::filter(cis_effect==1, value > 8) %>% 
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::group_by(gene) %>%
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::slice(1) %>%
  ungroup

ilc1_trans_markers <- ilc1_cv %>% 
  dplyr::filter(cis_effect!=1, value > 16) %>% 
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::group_by(gene) %>%
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::slice(1) %>%
  ungroup


ilc2_cis_markers <- ilc2_cv %>%
  dplyr::filter(cis_effect==1, value > 7) %>% 
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::group_by(gene) %>%
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::slice(1) %>%
  ungroup

ilc2_trans_markers <- ilc2_cv %>%
  dplyr::filter(cis_effect!=1, value > 16) %>% 
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::group_by(gene) %>%
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::slice(1) %>%
  ungroup



ilc3_cis_markers <- ilc3_cv %>% 
  dplyr::filter(cis_effect==1, value > 8) %>% 
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::group_by(gene) %>%
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::slice(1) %>%
  ungroup

ilc3_trans_markers <- ilc3_cv %>% 
  dplyr::filter(cis_effect!=1, value > 16) %>% 
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::group_by(gene) %>%
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::slice(1) %>%
  ungroup



lti_cis_markers <- lti_cv %>% 
  dplyr::filter(cis_effect==1, value > 8) %>% 
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::group_by(gene) %>%
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::slice(1) %>%
  ungroup

lti_trans_markers <- lti_cv %>% 
  dplyr::filter(cis_effect!=1, value > 16) %>% 
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::group_by(gene) %>%
  dplyr::arrange(desc(value_adj)) %>%
  dplyr::slice(1) %>%
  ungroup

eqtl_peaks <- do.call(rbind, 
                      list(ilc1_cis_markers, ilc2_cis_markers, 
                           ilc3_cis_markers, lti_cis_markers,
                           ilc1_trans_markers, ilc2_trans_markers, 
                           ilc3_trans_markers, lti_trans_markers))

write.csv(eqtl_peaks, "eqtl-peaks.csv")
```



```{r}

load(paste0(geno_path, 'genoprobs_rename.Rdata'))
eqtl_peaks <- read.csv(paste0(my_path, "eqtl-peaks.csv"))

samples <- grep("20180706", rownames(probs[[1]]), value = T)
nms <- str_extract(samples, "C[0-9]{1,2}_[0-9]")
eqtl_geno_df <- data.frame(BestCall = nms)

for( i in 1 : NROW(eqtl_peaks) ) {
  # i <- 1
  
  chr <- eqtl_peaks$marker_chr[i]
  marker <- eqtl_peaks$marker[i]
  celltype <- eqtl_peaks$cell_type[i]
  lodcol <- paste0(celltype, "_", eqtl_peaks$gene[i])
  
  eqtl_geno <- round(apply(probs[[chr]][samples,, marker], c(1,2), mean), 3)
  eqtl_geno_df[[paste0(lodcol, "_", marker)]] <- colnames(eqtl_geno)[apply(eqtl_geno, 1, which.max)]
  # i <- i+1
  
}

write.csv(eqtl_geno_df, paste0(my_path, "results/eqtl-genotypes.csv"))
```


```{r}
eqtl_peaks <- read.csv(paste0(my_path, "eqtl-peaks.csv"))
eqtl_geno_df <- read.csv(paste0(my_path, "results/eqtl-genotypes.csv"))
plot_df <- plot_df %>%
  dplyr::left_join(eqtl_geno_df, by = "BestCall")

plotgenes <- eqtl_peaks$gene[eqtl_peaks$cis_effect==1]
plotgenes <- as.character(gsub("-", ".", plotgenes))

for( g in plotgenes ) {
  
  eqtl_markers <- grep(g, colnames(eqtl_geno_df), value = TRUE)
  
  if( !is.null(eqtl_markers) ) {
    
    for( m in eqtl_markers ) {
      plot_out <- plot_df %>%
        dplyr::filter(!is.na(.data[[ m ]]),
                      !grepl("Y", .data[[ m ]]),
                      louvain_labels %in% c(1,2,3,4,5,8)) %>%
        ggplot(., aes_string("X_tsne1", "X_tsne2", color = g)) + 
          geom_point(shape = 46) + 
          scale_color_gradient(low = "lightgrey", high = "darkblue") +
          my_theme +
          facet_wrap( m ) + 
        labs(title = m)
      
      ggsave(filename = paste0(my_path, "results/figures/tsne-ilcs-bygenotype-", m, ".png"), 
             plot = plot_out, 
             dpi = 300)
    }
    
  }
  
}
```



```{r}

plotgenes <- eqtl_peaks$gene[eqtl_peaks$cis_effect==1]
plotgenes <- as.character(gsub("-", ".", plotgenes))

for( g in plotgenes ) {
  
  eqtl_markers <- grep(g, colnames(eqtl_geno_df), value = TRUE)
  
  if( !is.null(eqtl_markers) ) {
    
    for( m in eqtl_markers ) {
      
      genotype_order <- plot_df %>%
        dplyr::filter(!is.na(.data[[ m ]]),
                      !grepl("Y", .data[[ m ]])) %>%
        dplyr::group_by( BestCall, .data[[ m ]] ) %>%
        dplyr::summarise( topic = mean(.data[[ g ]] )) %>% 
        dplyr::group_by( .data[[ m ]] ) %>% 
        dplyr::summarise( topic = max(topic) ) %>% 
        dplyr::arrange(topic) %>% 
        .[[ m ]] %>%
        as.character
      
      plot_df <- plot_df %>%
        dplyr::mutate(founder_genotype = factor(.data[[ m ]], levels = genotype_order))
      
      plot_out <- plot_df %>%
        dplyr::filter(!is.na(.data[[ m ]]),
                      !grepl("Y", .data[[ m ]])) %>%
        dplyr::group_by(BestCall, founder_genotype) %>% 
        dplyr::summarise(topic_score = mean(.data[[ g ]])) %>%
        ggplot(., aes(founder_genotype, topic_score)) + 
          geom_point() +
          theme(
            panel.grid.minor=element_blank()
            ) +
          labs(title = paste(toupper(g), "at marker", strsplit(m, "_")[[1]][2]))
      
      ggsave(filename = paste0(my_path, "results/figures/scatter-egene-bygenotype-", m, ".png"), 
             plot = plot_out, 
             dpi = 300)

    }
    
  }
  
}
```


### Figure 5C

topic0 founder plot

```{r}

## 

```


### Figure 5D

topic3 founder plot

```{r}

```



## Figure 6

### Figure 6A

Cytokine heatmap

```{r cytokine_heatmap_prep}
cnames <- c("batch", "cage", "well", "IFNg", "IL5", "TNFa", "IL2", "IL6", "IL4", "IL10",
            "IL9", "IL17A", "IL17F", "IL22", "IL13", "coding", "mouse_id", "remove")
pheno_v1 <- lapply(1:4, function(i) {
  read_excel(path = paste0(pheno_path, "cytokine detection (version 1).xlsx"),
             sheet = i) %>% 
    na.omit
}) %>%
  do.call(rbind, .)
colnames(pheno_v1) <- cnames

# sample_map_df <- NULL
# ## batch3 - Broad_Inst_Xu_MURGIGV01_20180510 or Broad_Inst_Xu_MURGIGV01_20180607
batch3_samples <- pheno_v1$mouse_id[pheno_v1$batch == "Batch 3"]

# ## batch4 - Broad_Inst_Xu_MURGIGV01_20180706
batch4_samples <- pheno_v1$mouse_id[pheno_v1$batch == "Batch 4"]

pheno_bw <- read.csv(paste0(pheno_path, "pheno_bodyweight_processed.csv")) %>%
  dplyr::filter({ ID %in% batch3_samples & 
      genotyping_date %in% c("Broad_Inst_Xu_MURGIGV01_20180510", "Broad_Inst_Xu_MURGIGV01_20180607") } |
                {ID %in% batch4_samples & 
                    genotyping_date %in% c("Broad_Inst_Xu_MURGIGV01_20180706")}) %>%
  dplyr::rename(mouse_id = ID)


sample_map_df <- readr::read_csv(paste0(geno_path, "reference-data/DODB_GM_inventory_20180816.csv")) 
sample_map_df_batch3 <- sample_map_df %>% 
  filter(Project == 'XBI_Set3', Original.Mouse.ID %in% batch3_samples)
sample_map_df_batch4 <- sample_map_df  %>% 
  filter(Original.Mouse.ID %in% batch4_samples)

sample_df <- rbind(sample_map_df_batch3, sample_map_df_batch4) %>%
  dplyr::select(mouse_id = Original.Mouse.ID, aprobs_sample = Unique.Sample.ID) %>%
  dplyr::left_join( pheno_bw , by = "mouse_id")

pheno <- pheno_v1 %>%
  dplyr::left_join(sample_df) %>%
  dplyr::filter(remove == 0, !is.na(aprobs_sample)) %>%
  dplyr::mutate(IL13 = ifelse(is.na(as.numeric(IL13)), 41, as.numeric(IL13))) %>%
  as.data.frame
```

```{r cytokine_heatmap}
# Load latest version of heatmap.3 function
devtools::source_url("https://raw.githubusercontent.com/obigriffith/biostar-tutorials/master/Heatmaps/heatmap.3.R")

# pheno <- read.csv(paste0(pheno_path, "pheno_batch4_processed.csv"))
# pheno$Batches <- factor(pheno$Batches)
# cnames <- colnames(pheno)[c(3:14,16,20, 22:33)]
# options(na.action='na.pass')
# dat_na_removed <- model.matrix(object = as.formula(paste("~ 0 +", paste(cnames, collapse = "+"))), 
#                                  data = pheno,
#                                  contrasts.arg = list(Batches = contrasts(pheno$Batches, contrasts = F)))
# 
#   
# ## removing exluded mice
# pheno <- pheno %>% filter(excluded == 0)

# plotdata <- scale(pheno[, c("IgA_boxcox", "IgE_boxcox", "IgG1_boxcox", "IgG2_boxcox", "IgM_boxcox", "Mmcp1_boxcox")])
# rownames(plotdata) <- pheno$ID

plotdata <- scale(pheno[, 4:15])
plotdata <- pheno[, 4:15]
rownames(plotdata) <- pheno$mouse_id


# myheatmap(plotdata, "Boxcox Data")

bw_colors <- sort(unique(pheno$BW_1st))
names(bw_colors) <- colorRampPalette(brewer.pal(8, "Oranges"))(15)
  
mouse_color <- c("black","grey", "white")[pheno$Color]
mouse_batch <- c("Red", "Blue", "Green", "Purple")[pheno$Batches]
mouse_bw_colors <- names(bw_colors)[pheno$BW_1st - 12]

rowlabs <- t(cbind(mouse_color, mouse_batch, mouse_bw_colors))
rownames(rowlabs) <- c("Color", "Batch", "Weight")

rowlabs <- t(cbind(mouse_bw_colors))
rownames(rowlabs) <- c("Weight")

main_title = "Cytokine"
par(cex.main=1)
heatmap.3(x = plotdata, 
          na.rm = TRUE, 
          scale="none", 
          dendrogram = "both", 
          margins = c(6,12),
          Rowv = TRUE, 
          Colv = TRUE, 
          RowSideColors = rowlabs, 
          symbreaks = FALSE, 
          key = TRUE, 
          symkey = FALSE,
          density.info = "none", 
          trace = "none", 
          main = main_title, 
          col = rev(colorRampPalette(brewer.pal(8, "RdYlGn"))(25)), 
          labCol = cnames[4:15],
          labRow = FALSE,
          RowSideColorsSize = 2, 
          KeyValueName = "Cytokine")
legend("topright",
       legend = c("Batch1", "Batch2", "Batch3", "Batch4", "", "black mouse","white mouse", "grey mouse", "", "Low Weight", "Medium Weight", "High Weight"),
      fill = c("Red", "Blue", "Green", "Purple", "white", "black", "white", "grey", "white", names(bw_colors)[c(2,8,17)]), 
         border = FALSE, 
         bty = "n", 
         y.intersp = 0.7, 
         cex = 0.7)
```



 
### Figure 6B

 Cytokine Co-localization

```{r}
## IL-22
ccre[il22_qtl == 1, lapply(.SD, sum), .SDcols = 37:96]

il22_ilc3_markers <- ccre[il22_qtl==1 & chr == 1 & ilc3_eqtl_loci_mean==1, marker]
table(ilc3_mean$gene[ilc3_mean$marker %in% il22_ilc3_markers])
table(lti_mean$gene[lti_mean$marker %in% il22_ilc3_markers])



topic19_marker <- ccre[il22_qtl==1 & chr == 8 & {ilc2_ilc3_prop_qtl==1 | topic19_qtl==1},marker]

table(ilc1_cv$gene[ilc1_cv$marker %in% topic19_marker])
table(ilc2_cv$gene[ilc2_cv$marker %in% topic19_marker])
table(ilc3_cv$gene[ilc3_cv$marker %in% topic19_marker])
table(lti_cv$gene[lti_cv$marker %in% topic19_marker])



## TNFa
ccre[tnf_qtl == 1, lapply(.SD, sum), .SDcols = 37:96]

tnf_chr5_markers_loci1 <- ccre[tnf_qtl==1 & chr == 5 & pos < 115000000, marker]
table(ilc3_mean$gene[ilc3_mean$marker %in% tnf_chr5_markers_loci1])


tnf_chr5_markers_loci2 <- ccre[tnf_qtl==1 & chr == 5 & pos > 115000000, marker]





```


### Figure 6C

```{r}

```


### Figure 6D

```{r}

```


### Figure 6: Supplementary



```{r}
# Load latest version of heatmap.3 function
devtools::source_url("https://raw.githubusercontent.com/obigriffith/biostar-tutorials/master/Heatmaps/heatmap.3.R")

pheno <- read.csv(paste0(pheno_path, "pheno_batch4_processed.csv"))
pheno$Batches <- factor(pheno$Batches)
cnames <- colnames(pheno)[c(3:14,16,20, 22:33)]
options(na.action='na.pass')
dat_na_removed <- model.matrix(object = as.formula(paste("~ 0 +", paste(cnames, collapse = "+"))), 
                                 data = pheno,
                                 contrasts.arg = list(Batches = contrasts(pheno$Batches, contrasts = F)))

  
## removing exluded mice
pheno <- pheno %>% filter(excluded == 0)

# plotdata <- scale(pheno[, c("IgA_boxcox", "IgE_boxcox", "IgG1_boxcox", "IgG2_boxcox", "IgM_boxcox", "Mmcp1_boxcox")])
# rownames(plotdata) <- pheno$ID

plotdata <- scale(pheno[, c("IgA_boxcox", "IgE_boxcox", "IgG1_boxcox", "IgG2_boxcox", "IgM_boxcox")])
rownames(plotdata) <- pheno$ID


# myheatmap(plotdata, "Boxcox Data")

bw_colors <- sort(unique(pheno$BW_1st))
names(bw_colors) <- colorRampPalette(brewer.pal(8, "Oranges"))(17)
  
mouse_color <- c("black","grey", "white")[pheno$Color]
mouse_batch <- c("Red", "Blue", "Green", "Purple")[pheno$Batches]
mouse_bw_colors <- names(bw_colors)[pheno$BW_1st - 12]

# rowlabs <- t(cbind(mouse_color, mouse_batch, mouse_bw_colors))
# rownames(rowlabs) <- c("Color", "Batch", "Weight")

rowlabs <- t(cbind(mouse_bw_colors))
rownames(rowlabs) <- c("Weight")

main_title = "Antibody"
par(cex.main=1)
heatmap.3(x = plotdata, 
          na.rm = TRUE, 
          scale="none", 
          dendrogram = "both", 
          margins = c(6,12),
          Rowv = TRUE, 
          Colv = TRUE, 
          RowSideColors = rowlabs, 
          symbreaks = FALSE, 
          key = TRUE, 
          symkey = FALSE,
          density.info = "none", 
          trace = "none", 
          main = main_title, 
          col = rev(colorRampPalette(brewer.pal(8, "RdYlGn"))(25)), 
          labCol = sapply(c("IgA", "IgE", "IgG1", "IgG2", "IgM", "Mmcp1"), function(p) na.omit(str_extract(colnames(plotdata), p))),
          labRow = FALSE,
          RowSideColorsSize = 2, 
          KeyValueName = "Antibody")
legend("topright",
       legend = c("Batch1", "Batch2", "Batch3", "Batch4", "", "black mouse","white mouse", "grey mouse", "", "Low Weight", "Medium Weight", "High Weight"),
      fill = c("Red", "Blue", "Green", "Purple", "white", "black", "white", "grey", "white", names(bw_colors)[c(2,8,17)]), 
         border = FALSE, 
         bty = "n", 
         y.intersp = 0.7, 
         cex = 0.7)
```



### Figure 6Bs

  - antibody-lod-heatmap.Rmd

```{r}

## QTL data
qtl_data <- read_csv(paste0(my_path, "results/Antibody_QTL_wBatchCorrection.csv"))
## Marker Map
load(paste0(geno_path, "Regev_map_20171221.Rdata"))
## SNP Info
load(paste0(my_path, "data/GM_snps.Rdata"))
## Genotype Data
# load(paste0(geno))

indexes <- qtl2:::map_to_index(map)
xpos <- qtl2:::map_to_xpos(map, gap = 25)
xpos <- data.frame(marker = names(xpos), xpos = xpos)
chrbound <- qtl2:::map_to_boundaries(map, gap = 25)


evens <- seq(2, length(map), by = 2)

xpos_median <- xpos[mapply(median, indexes), ]
xpos_min <- xpos[mapply(min, indexes), ]
xpos_max <- xpos[mapply(max, indexes), ]


plotdata <- qtl_data %>% 
  #tibble::rownames_to_column %>%
  rename(marker = X1) %>%
  data.table::melt("marker") %>%
  left_join(GM_snps) %>%
  left_join(xpos) %>%
  mutate(value = as.numeric(value))


dat <- as.matrix(qtl_data[,-1])
rownames(dat) <- qtl_data$X1

cnames <- c("IgA_boxcox", "IgE_boxcox", "IgG1_boxcox", "IgG2_boxcox", "IgM_boxcox", "Mmcp1_boxcox", "BW_1st")
distout <- dist(t(dat[,cnames]))

 hc <- hclust(distout)
 hc$order


(p <- ggplot(plotdata[plotdata$variable %in% cnames, ], aes(x = xpos, y = variable)) + 
  geom_tile(aes(color = value)) + 
  scale_color_gradientn(colours = brewer.pal(11, "RdBu")[8:1]) +
  labs(title= "Antibody LOD Heatmap", x = "Chromosome", y = "Antibody", color = "LOD") + 
  theme(panel.background = element_rect(fill = "white"), 
        panel.grid.minor = element_line(color = "black", size = 500, linetype = 8)) + 
  scale_x_discrete(limits = xpos_median$xpos, 
                   labels = paste0(c(1:19, "X"))) + 
  scale_y_discrete(limits = cnames[hc$order], 
                   labels = c("IgA", "IgE", "IgG1", "IgG2", "IgM", "Mmcp1", "BW")[hc$order]))

ggsave("lod_heatmap.png", p)
```



## Figure 7


### Figure 7A

```{r}
founders1 <- fread(paste0(my_path, 'data/founders1-louvain_labels-labels-transfer/obs.csv'),
                   data.table = FALSE)
founders2 <- fread(paste0(my_path, 'data/founders2-louvain_labels-labels-transfer/obs.csv'),
                   data.table = FALSE)


founders <- rbind(founders1, founders2)
```



```{r}
for( t in paste0("topic", 0:19) ) {
  
  # t <- "topic0"
  
  genotype_order <- founders %>%
    dplyr::filter(best_clean != "AMB", best_clean != "nan") %>%
    dplyr::group_by( best_clean ) %>%
    dplyr::summarise( topic = mean(.data[[ t ]]) ) %>% 
    dplyr::arrange(topic) %>% 
    .$best_clean %>%
    as.character
  
  founders <- founders %>%
    dplyr::mutate(founder_genotype = factor(best_clean, levels = genotype_order))
  
  
  plot_out <- founders %>%
    dplyr::filter(best_clean != "AMB", best_clean != "nan") %>%
    dplyr::mutate(topic_score = .data[[ t ]]) %>%
    # dplyr::group_by(best_clean, founder_genotype) %>% 
    # dplyr::summarise(topic_score = mean(.data[[ t ]])) %>%
    ggplot(., aes(founder_genotype, topic_score)) + 
      geom_point() +
      theme(
        panel.grid.minor=element_blank()
        ) + 
    stat_summary(fun.data=mean_sdl, fun.args = list(mult = 2),
                 geom = "errorbar", color = "red", width = 0.2) +
    stat_summary(fun=mean, geom="point", color="red") + 
    ylim(0, NA) + 
    labs(title = paste(tools::toTitleCase(t), "by Founder Mouse"), 
         x = "Founder Genotype", 
         y = "Topic Score")
  
  ggsave(
    filename = paste0(my_path, "results/figures/scatter-topicscore-byfounder-",t , ".png"), 
    plot = plot_out, 
    dpi = 300)
  
}

```


```{r}
founders1X <- fread(paste0(my_path, 'data/founders1-louvain_labels-labels-transfer/X.csv'),
                   data.table = FALSE)
founders2X <- fread(paste0(my_path, 'data/founders2-louvain_labels-labels-transfer/X.csv'),
                   data.table = FALSE)

cnames <- intersect(colnames(founders1X), colnames(founders2X))
foundersX <- rbind(founders1X[, cnames], founders2X[, cnames])
colnames(foundersX) <- gsub("-", ".", colnames(foundersX)) 
```

```{r}
founders <- founders %>%
  dplyr::left_join(foundersX)
```


```{r}

#  [1] "scatter-egene-bygenotype-ILC1_Cnn2_UNCHS028864.png"        
#  [2] "scatter-egene-bygenotype-ILC1_Hopx_UNCHS014802.png"        
#  [3] "scatter-egene-bygenotype-ILC1_Ppp1r2_JAX00417146.png"      
#  [4] "scatter-egene-bygenotype-ILC2_Abcb1b_UNCHS013491.png"      
#  [5] "scatter-egene-bygenotype-ILC2_Pdcd5_UNC12624590.png"       
#  [6] "scatter-egene-bygenotype-ILC2_Sptssa_JAX00334445.png"      
#  [7] "scatter-egene-bygenotype-ILC3_H2.Q4_UNCHS044197.png"       
#  [8] "scatter-egene-bygenotype-ILC3_Hmha1_UNCHS028841.png"       
#  [9] "scatter-egene-bygenotype-ILC3_Lgals3_UNC23954465.png"      
# [10] "scatter-egene-bygenotype-ILC3_Necap2_JAX00568909.png"      
# [11] "scatter-egene-bygenotype-LTi_H2.T23_backupUNC170076641.png"
# [12] "scatter-egene-bygenotype-LTi_Lgals3_UNCHS038151.png" 

ilc1_genelist <- c("Cnn2", "Hopx", "Ppp1r2")  # louvain_labels - 8
ilc2_genelist <- c("Abcb1b", "Pdcd5", "Sptssa") # louvain_labels - 3
ilc3_genelist <- c("H2.Q4", "Lgals3", "Necap2") # louvain_labels - 2,4
lti_genelist <- c("H2.Q4", "Lgals3", "H2.T23") # louvain_labels - 1,5

## maybe put cell type in plot name
for( g in lti_genelist ) {
  
  # g <- "Hopx"
  
  genotype_order <- founders %>%
    dplyr::filter(best_clean != "AMB", best_clean != "nan") %>%
    dplyr::filter(louvain_labels %in% c(2,4)) %>%
    dplyr::group_by( best_clean ) %>%
    dplyr::summarise( gene = mean(.data[[ g ]]) ) %>% 
    dplyr::arrange(gene) %>% 
    .$best_clean %>%
    as.character
  
  founders <- founders %>%
    dplyr::mutate(founder_genotype = factor(best_clean, levels = genotype_order))
  
  
  plot_out <- founders %>%
    dplyr::filter(best_clean != "AMB", best_clean != "nan") %>%
    dplyr::mutate(gene = .data[[g]]) %>%
    # dplyr::group_by(best_clean, founder_genotype) %>% 
    # dplyr::summarise(topic_score = mean(.data[[ t ]])) %>%
    ggplot(., aes(founder_genotype, gene)) + 
      geom_point() +
      theme(
        panel.grid.minor=element_blank()
        ) +
    stat_summary(fun.data=mean_sdl, fun.args = list(mult = 2),
                 geom = "errorbar", color = "red", width = 0.2) +
    stat_summary(fun=mean, geom="point", color="red") + 
    ylim(0, NA) + 
    labs(title = paste(tools::toTitleCase(g), "by Founder Mouse"), 
         x = "Founder Genotype", 
         y = "Gene Expression")
  
  ggsave(
    filename = paste0(my_path, "results/figures/scatter-expression-byfounder-", g , ".png"), 
    plot = plot_out, 
    dpi = 300)
  
}

```



## Allergy and Proportion Overlap

```{r}
library(UpSetR)


vars <- fread(paste0(my_path, "data/scanpy/allchannels/vars.csv"), data.table = FALSE)
check_vars <- grep("prop|_steady|_allergy", colnames(vars), value = T)


gene_list <- combn(check_vars, 2, function(x) {
  list(vars$index[vars[[x[1]]] == 1], vars$index[vars[[x[2]]] == 1])
}, simplify = FALSE)
```


### Lineage Tracing?

lineage markers were poorly, if at all expressed, and did not differ between WT and Tox/ cells. Genes Itga4 and Itgb7,

```{r}
id2plot <- plot_df %>% 
  dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>% 
  my_plot(., "Id2", embedding = "net_fle") + theme_minimal()

toxplot <- plot_df %>% 
  dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>% 
  my_plot(., "Tox", embedding = "net_fle") + theme_minimal()

rbpjplot <- plot_df %>% 
  dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>% 
  my_plot(., "Rbpj", embedding = "net_fle") + theme_minimal()

cxcr6plot <- plot_df %>% 
  dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>% 
  my_plot(., "Cxcr6", embedding = "net_fle") + theme_minimal()

rorcplot <- plot_df %>% 
  dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>% 
  my_plot(., "Rorc", embedding = "net_fle") + theme_minimal()

gata3plot <- plot_df %>% 
  dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>% 
  my_plot(., "Gata3", embedding = "net_fle") + theme_minimal()

tbetplot <- plot_df %>% 
  dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>% 
  my_plot(., "Tbet", embedding = "net_fle") + theme_minimal()

ncr1plot <- plot_df %>% 
  dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>% 
  my_plot(., "Ncr1", embedding = "net_fle") + theme_minimal()

ccr6plot <- plot_df %>% 
  dplyr::filter(louvain_labels %in% c(1,2,3,4,5,8)) %>% 
  my_plot(., "Ccr6", embedding = "net_fle") + theme_minimal()




plot_grid(rbpjplot, toxplot, id2plot, 
          cxcr6plot, rorcplot, gata3plot, 
          tbetplot, ncr1plot, ccr6plot, nrow=3)

```

