---
title: "cytokine-founder-alleles"
format: html
editor: visual
---

## Cytokine Loci by Founder Allele

```{r setup}
library(qtl2)
library(data.table)
library(readr)
library(dplyr)
library(ggplot2)

domice_dir <- "/workspace/fasi-domice/"
data_dir <- paste0(domice_dir, "data/")
results_dir <- paste0(domice_dir, "results/")
figure_dir <- paste0(domice_dir, "figures/")
```

```{r config_reference}
# project_path <- "/ahg/regevdata/projects/FASI_DOmice/"
# if( length(dir(project_path)) == 0 ) project_path <- "/Volumes/ahg_regevdata/projects/FASI_DOmice/"
source(paste0(domice_dir, "src/helpers.R"))

load(paste0(data_dir, "genotype/GM_snps.Rdata"))
## founder key
founderdf <- data.frame(founder_letter = LETTERS[1:8], 
                        founder_color = qtl2::CCcolors, 
                        founder_domice = names(qtl2::CCcolors),
                        founder_cytokine_name = c("A","B6","129s",
                                                  "NOD", "NZO",
                                                  "CAST","PWK","WSB"),
                        best_clean = c("A_J", "C57BL_6NJ",
                                       "129S1_SvImJ", "NOD_ShiLtJ",
                                       "NZO_HlLtJ", "CAST_EiJ", 
                                       "PWK_PhJ", "WSB_EiJ"),
                        stringsAsFactors = FALSE)
```

```{r lods_df}
lods <- fread(paste0(results_dir,"cytokines/qtl-steady-cytokines-lods.csv.gz"))
```

```{r alleleprobs}
load(paste0(data_dir, "genotype/alleleprobs.Rdata"))
antibody <- read_csv(paste0(data_dir, "phenotype/pheno_batch4_processed.csv"))
cytokines <- readxl::read_excel(paste0(data_dir, "phenotype/cytokine production -steady state.xlsx"))

```

```{r alleleprobs_filtering}
aprob_rnames <- rownames(aprobs[[1]])

name_vec <- sapply(cytokines$mouse_id, function(x) grep(x, aprob_rnames, value=T))

name_map <- data.frame(full_id = sapply(name_vec, last)) %>% 
  tibble::rownames_to_column("mouse_id") %>% 
  dplyr::filter(!is.na(full_id))
```

```{r format_cytokine_df}
cytokines_qtl_df <- cytokines %>%
  dplyr::left_join(name_map) %>%
  dplyr::filter(!is.na(full_id))
```

```{r select_ids}
keep_ids <- intersect(aprob_rnames, cytokines_qtl_df$full_id)
```

```{r qtl_df}
cytokines_qtl_df <- cytokines_qtl_df %>%
  filter(full_id %in% keep_ids)
  
```

```{r kinship_matrix}
cores <- 2
## Load Phenotypes (all females)
pheno <- cytokines_qtl_df %>% as.data.frame()
rownames(pheno) <- pheno$full_id


## Sync sample names
length(keep_ids)

pheno <- pheno[keep_ids, ,drop = FALSE]
for( chr in names(aprobs) ) {
  aprobs[[chr]] <- aprobs[[chr]][keep_ids, , ]
}


# Creating kinship matrix
# kinship_loco
K <- calc_kinship(probs = aprobs, type = "loco", cores = cores)
```

```{r}
ccre <- read_csv(paste0(data_dir, "references/GM_SNPS_Consequence_cCRE.csv"))
marker_map <- readRDS(paste0(data_dir, "genotype/Regev_map_20171221.rds"))
xpos <- qtl2:::map_to_xpos(marker_map, gap = 25)
xchrbound <- qtl2:::map_to_boundaries(marker_map, gap = 25)

## Make Marker Reference Dataset
MarkerRef <- data.frame(marker = names(xpos), xpos = xpos, stringsAsFactors = FALSE) %>%
  left_join(
    {ccre %>% dplyr::select(marker, chr)}
  )

```

## IgE

```{r}
ige <- pheno$IgE
names(ige) <- pheno$full_id
# IgE - https://www.informatics.jax.org/marker/MGI:2685746
# Chromosome 12: 113,269,260-113,273,248 reverse strand.
c2blup <- scan1blup(aprobs[,"12"], ige, K[["12"]])
coef_df <- as.data.frame(c2blup)
```

```{r}
p1 <- coef_df %>%
  tibble::rownames_to_column("X") %>%
  dplyr::rename(marker = X) %>%
  left_join(data.frame(marker = names(xpos), 
                       xpos = xpos, stringsAsFactors = FALSE)) %>%
  dplyr::select(-intercept) %>%
  pivot_longer(cols = A:H, names_to = "founder_letter") %>%
  left_join(founderdf, by = "founder_letter") %>%
  filter(xpos > 2009, xpos < 2010) %>%
  ggplot(., aes(x = xpos, y = value, color = founder)) +
    geom_line() + 
    scale_color_manual(values = qtl2::CCcolors) +
    scale_x_continuous(breaks = apply(xchrbound, 2, median), label = c(as.character(1:19), "X")) +
    # geom_vline(xintercept = 2009)
    labs(x = "Chromosome of SNP", y = "LOD", title = "IgE") +
    theme(
      axis.line=element_blank(),
      axis.ticks=element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    ) +
    theme_minimal()
```

## Do Mice Cytokines

### Reference Dataframe

This part is the reference data from specific to cytokine location of the cytokines for the mm10 (GRCm38 reference) positions.

```{r ref_data}
cytokine_loci <- data.frame(
  Cytokine = c("IFNg", "IL-5", "TNFa", "IL-2", "IL-6",
               "IL-4", "IL-10", "IL-9", "IL-17A", "IL-17F",
               "IL-22", "IL-13"),
  Chromosome = c("10", "11", "17", "3", "5",
                 "11", "1", "13", "1", "1",
                 "10", "11"),
  Genomic_Position = c(
    "118441046-118445892",
    "53720794-53725106",
    "35199381-35202007", 
    "37120523-37125959",
    "30013114-30019981",
    "53602982-53618669",
    "131019845-131024974",
    "56479277-56482246",
    "20730905-20734496",
    "20777146-20790617",
    "118204942-118210047",
    "53631324-53634702"
  ),
  stringsAsFactors = FALSE
)

cnames <- cytokine_loci$Cytokine
```

This next chunk is to specify the cytokine to focus on and take the above reference positions to grab the makrers in the GigaMUGA array that correspond to the cytokine location.

```{r marker_list}
cname <- "IL-2"

extract_marker_list <- function(cname) {
  
  pvec <- pheno[,cname]
  names(pvec) <- pheno$full_id
  pchr <- cytokine_loci %>% 
    filter(Cytokine == cname) %>% 
    pull(Chromosome)
  genomic_position <- cytokine_loci %>% 
    filter(Cytokine == cname) %>% 
    pull(Genomic_Position)
  
  pos_start <- stringr::str_split(genomic_position,'-',simplify = T)[1]
  pos_start <- floor(as.numeric(pos_start)/10e4)*10e4
  pos_end <- stringr::str_split(genomic_position,'-',simplify = T)[2]
  pos_end <- ceiling(as.numeric(pos_end)/10e4)*10e4
  marker_list <- ccre %>% 
    filter(pos > pos_start, 
           pos < pos_end,
           chr == pchr) %>% 
    pull(marker)
  
  marker_list
  
}
```

This part extracts the cytokine region markers from the allele probs array and averages them across the DO mice and predicted founder allele probabilities. This outputs a dataframe where the rows are the DO mouse, the columns are the Founder Allele (A-H) and the entries are the average probability across all the markers at the cytokine location.

```{r mean_df}
aprob_chr10_markers <- intersect(dimnames(aprobs[["10"]])[[3]], marker_list)
out_df <- aprobs[["10"]][,,aprob_chr10_markers]

# Compute mean across all data frames
mean_df <- apply(out_df,c(1,2),mean) %>% as.data.frame() %>%
  tibble::rownames_to_column("full_id")
```

```{r}
l <- sapply(cnames, function(cname) {
  pchr <- cytokine_loci %>% 
    filter(Cytokine == cname) %>% 
    pull(Chromosome)
  
  marker_list <- extract_marker_list(cname)
  aprob_markers <- intersect(dimnames(aprobs[[pchr]])[[3]], marker_list)
  out_df <- aprobs[[pchr]][,,aprob_markers]
  
  # Compute mean across all data frames
  mean_df <- apply(out_df,c(1,2),mean) %>% 
    as.data.frame() %>%
    tibble::rownames_to_column("full_id")
  
  mean_df
  
}, USE.NAMES = TRUE, simplify = FALSE)
```

```{r}
get_columns_above_threshold <- function(df, threshold = 0.4) {

  # For each row, return column names with value == 1 and mean > threshold
  apply(df, 1, function(row) {
    cols <- colnames(df)[row > threshold]
    if (length(cols) == 1) {
      paste0(rep(cols,2),  collapse="") }
    else paste0(cols,  collapse="")
    
  })
}

```

This part takes the maximum probability and imputes that as the founder allele. Maybe think of how to collapse to genoprobs (e.g two founder alleles per mouse) or upload the genoprobs matrix also

```{r}

create_founder_call <- function(l, cname) {
  
  # Add a column with the name of the column that has the max value in each row
  df_with_max_col <- l[[cname]] %>%
    rowwise() %>%
    mutate(founder_letter = names(.)[-1][which.max(c_across(A:H))]) %>%
    ungroup() %>%
    dplyr::left_join(dplyr::select(pheno, full_id, !!cname)) %>%
    dplyr::left_join(dplyr::rename(founderdf, founder = founder_domice))
  
  df_with_max_col$founder_geno <- get_columns_above_threshold(l[[cname]][,2:9])
  df_with_max_col
  
}
```

Averaging across the imputed founder allele to see cytokine levels for each allele. This can be used to match with the founder cytokine data provided.

```{r}

summarize_by_allele <- function(df_with_max_col, cname) {
  
  domice_df <- df_with_max_col %>% 
    group_by(founder = best_clean) %>% 
    summarize(across(all_of(cname), ~mean(.x, na.rm = TRUE)))
  
  domice_df
  
}
```

```{r}
cname <-  names(l)[1]
cat(cname,'\n')
df_with_max_col <- create_founder_call(l, cname)
plot_df_domice <- summarize_by_allele(df_with_max_col, cname)
```

### Founder Correlations

Geno Calc Correlations:
IFNg: -0.579258
IL-5: 0.02514933
TNFa: -0.248
IL-2: 0.678
IL-6: -0.575
IL-4: -0.23
IL-10: NA
IL-9: 0.267
IL-17A: -0.425
IL-17F : NA
IL-22: NA
IL-13: NA

```{r genos_cor}
cor_df <- df_with_max_col %>%
  dplyr::mutate(across(A:H, function(x) as.numeric(x>0.4))) %>%
  dplyr::select(sym(cname), A:H) %>%
  tidyr::pivot_longer(!sym(cname)) %>%
  dplyr::filter(value==1) %>%
  dplyr::group_by(name) %>%
  dplyr::summarise(domice_cytokine = mean(.data[[cname]], na.rm=T)) %>%
  dplyr::left_join(dplyr::select(founderdf, name = founder_letter,
                                 founder = best_clean)) %>% 
  left_join({
    broad_founder_all %>%
      mutate(founder_cytokine = as.numeric(.data[[cname]])) %>% 
      group_by(founder = best_clean) %>% 
      summarise(founder_cytokine = mean(founder_cytokine, na.rm=TRUE)) 
    })

cor(cor_df[,c('domice_cytokine','founder_cytokine')], use='pairwise.complete')
```


```{r}

df <- plot_df_domice %>% 
  left_join({
    broad_founder_all %>%
      mutate(cytokine = as.numeric(`IFNg`)) %>% 
      group_by(founder = best_clean) %>% 
      summarise(cytokine = mean(cytokine, na.rm=TRUE)) 
    })

cor(df[,2:3], use='pairwise.complete')
plot(df[,2:3])
```

Plotting DO mice imputed alleles by cytokine levels.

```{r domice_founder_plot}
founder_order <- c("A_J", "C57BL_6NJ", "129S1_SvImJ", "NOD_ShiLtJ", 
                   "NZO_HlLtJ", "CAST_EiJ", "PWK_PhJ", "WSB_EiJ")

founder_colors <- founderdf$founder_color
names(founder_colors) <- founderdf$best_clean

df_with_max_col %>%
  mutate(best_clean = factor(best_clean, levels = founder_order)) %>%
  ggplot(aes(best_clean, .data[[cname]], color = best_clean)) + 
    geom_violin() +
    geom_jitter() +
    scale_color_manual(values = founder_colors) +
    theme_minimal() +
    labs(title = "Cytokine Profile by Predicted Founder Allele",
         x = "Founder (Predicted)",
         y = paste0(cname, " (pg/ml)"),
         color = "Founder")

```

### QTL Analysis

Running QTL analysis to get coefficient of predicted founder alleles impact at each cytokine location for that cytokine region. (cis-cytokine predictions.)

```{r run_qtl_coef}
for( cname in cnames) {
  pvec <- pheno[,cname]
  names(pvec) <- pheno$full_id
  pchr <- cytokine_loci %>% 
    filter(Cytokine == cname) %>% 
    pull(Chromosome)
  genomic_position <- cytokine_loci %>% 
    filter(Cytokine == cname) %>% 
    pull(Genomic_Position)
  
  pos_start <- stringr::str_split(genomic_position,'-',simplify = T)[1]
  pos_start <- floor(as.numeric(pos_start)/10e4)*10e4
  pos_end <- stringr::str_split(genomic_position,'-',simplify = T)[2]
  pos_end <- ceiling(as.numeric(pos_end)/10e4)*10e4
  marker_list <- ccre %>% 
    filter(pos > pos_start, 
           pos < pos_end,
           chr == pchr) %>% 
    pull(marker)
  
  pxpos <- MarkerRef %>% 
    filter(marker %in% marker_list, chr == pchr) %>%
    pull(xpos) %>% 
    round %>% 
    first
  
  # IgE - https://www.informatics.jax.org/marker/MGI:2685746
  # Chromosome 12: 113,269,260-113,273,248 reverse strand.
  c2blup <- scan1blup(aprobs[,pchr], pvec, K[[pchr]])
  coef_df <- as.data.frame(c2blup)
  write.csv(coef_df, paste0(results_dir, "cytokines/founder_coef_", cname, ".csv"))
  
  ## plot results
  p1 <- coef_df %>%
    tibble::rownames_to_column("X") %>%
    dplyr::rename(marker = X) %>%
    dplyr::left_join(data.frame(marker = names(xpos), 
                                xpos = xpos, stringsAsFactors = FALSE)) %>%
    dplyr::select(-intercept) %>%
    tidyr::pivot_longer(cols = A:H, names_to = "founder_letter") %>%
    dplyr::left_join(founderdf, by = "founder_letter") %>%
    dplyr::filter(xpos > (pxpos-1), xpos < (pxpos+1)) %>%
    ggplot(., aes(x = xpos, y = value, color = founder)) +
      geom_line() + 
      scale_color_manual(values = qtl2::CCcolors) +
      scale_x_continuous(breaks = apply(xchrbound, 2, median), label = c(as.character(1:19), "X")) +
    # geom_vline(xintercept = 2009)
      labs(x = "Chromosome of SNP", y = "LOD", title = cname) +
      theme(
        axis.line=element_blank(),
        axis.ticks=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      ) +
      theme_minimal()
  ggsave(paste0("founder_loci_plot_", cname, ".png"), p1)

}
```

## Founder Cytokines

### Previous Publications

Whereas cellular composition was shaped by interactions between genotype and environment, cytokine response heterogeneity including IFNγ concentrations was primarily driven by genotype with consequence on worm burden

https://pmc.ncbi.nlm.nih.gov/articles/PMC11224019/#MOESM29

```{r}
founder_map <- data.frame(
  Genotype = c("C57BL/6","129S1",  "PWK/PhJ"),
  founder = c("C57BL_6NJ", "129S1_SvImJ","PWK_PhJ")
)

public_cytokines <- readxl::read_excel(paste0(data_dir, "cytokines/41590_2024_1862_MOESM10_ESM.xlsx"), skip = 1) %>%
  rename(IFNg = `IFN-gamma`,
         TNFa = `TNF-a`) %>%
  dplyr::left_join(founder_map)
  
```

```{r public_plot}
#founder_order <- c("C57BL_6NJ", "129S1_SvImJ", "PWK_PhJ")

public_cytokines %>%
  dplyr::mutate(Sample = factor(PrincetonID),
                founder = factor(founder, levels = founder_order)) %>%
  dplyr::filter(InfectionStatus == "Uninfected") %>%
  ggplot(aes(Sample, .data[[cname]], fill = founder)) +
    geom_col(position='dodge') +
    geom_point(aes(Sample, .data[[cname]]), color = "black") +
    scale_fill_manual(values = founder_colors) +
    facet_wrap(~founder, scales = 'free_x') +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Cytokine Profile by Founder (Oyesola 2024)",
         x = "Founder",
         y = paste0(cname, " (pg/ml)"),
         color = "Founder")
```

### Broad Generation

This all needs processing to get it in an analyzable format. Things to consider - Figure out the concentration calculations used - Match / reproduce what is in the slide deck - think of a way of standardizing cytokine concentraions across the disparate experiments for comparison. - maybe just report raw correlations as the metric (ignore scaling)

```{r read_founder_cytokine}
## read legened
legend_df = read.csv(paste0(data_dir,"cytokines/founders/Legendplex_plate_maps_05292025 - Legend plex plate maps.csv"))

## plate1 - standard curve info
founder_plate1_cytokines_standard <- readxl::read_excel(paste0(data_dir,"cytokines/founders/TMD_Plate1_report_2025-05-30.xlsx"), sheet = 1)

## plate1 - predicted concentrations
founder_plate1_cytokines <- readxl::read_excel(paste0(data_dir,"cytokines/founders/TMD_Plate1_report_2025-05-30.xlsx"), sheet = 2)

## plate1 - raw output readings
founder_plate1_cytokines_raw <- readxl::read_excel(paste0(data_dir,"cytokines/founders/TMD_Plate1_report_2025-05-30.xlsx"), sheet = 3)


## plate2 - standard curve info
founder_plate2_cytokines_standard <- readxl::read_excel(paste0(data_dir,"cytokines/founders/TMD_Plate2_report_2025-05-30.xlsx"), sheet = 1)

## plate2 - predicted concentrations
founder_plate2_cytokines <- readxl::read_excel(paste0(data_dir,"cytokines/founders/TMD_Plate2_report_2025-05-30.xlsx"), sheet = 2)

## plate2 - raw output readings
founder_plate2_cytokines_raw <- readxl::read_excel(paste0(data_dir,"cytokines/founders/TMD_Plate2_report_2025-05-30.xlsx"), sheet = 3)

```

```{r founder_dataframe}

broad_founder_all <- rbind(founder_plate1_cytokines,founder_plate2_cytokines) %>%
  dplyr::rename(fcs = well) %>%
  dplyr::left_join(legend_df) %>%
  dplyr::select(fcs, Sample, 
                IFNg = `IFN-γ (A4)`,
                `IL-5` = `IL-5 (A5)`,
                TNFa = `TNF-α (A6)`,
                `IL-2` = `IL-2 (A7)`,
                `IL-6` = `IL-6 (A8)`,
                `IL-4` = `IL-4 (A10)`,
                `IL-10` = `IL-10 (B2)`,
                `IL-9` = `IL-9 (B3)`,
                `IL-17A` = `IL-17A (B4)`,
                `IL-17F` = `IL-17F (B5)`,
                `IL-22` = `IL-22 (B7)`,
                `IL-13` = `IL-13 (B9)`) %>%
  dplyr::mutate(founder_cytokine_name = stringr::str_split(Sample, '-', simplify = TRUE)[,1]) %>%
  dplyr::left_join(founderdf)
```

```{r broad_plot}
sample_order <- c(
  'A-J-30002', 'A-J-30003', 'A-J-30004', 'A-J-30005', 
  'A-J-30006', 'A-J-30007','A-J-30008', 
  'B6-30009', 'B6-30010', 'B6-30011', 'B6-30012', 
  '129s-30013', '129s-30014', '129s-30015', '129s-30016', '129s-30017',
  '129s-30018', '129s-30019', '129s-30020', '129s-30021', '129s-30022', 
  'NOD-30023', 'NOD-30024', 'NOD-30025', 'NOD-30026', 
  'NOD-30027', 'NOD-30028', 'NOD-30029', 'NOD-30030', 'NOD-30031',
  'NZO-30032', 'NZO-30033', 
  'CAST-30035', 'CAST-30036', 'CAST-30037', 'CAST-30038', 'CAST-30039',
  'CAST-30044', 'CAST-30045', 'CAST-30046', 'CAST-30047', 
  'PWK-30040', 'PWK-30041', 'PWK-30042', 'PWK-30043', 
  'WSB-30048', 'WSB-30049', 'WSB-30050', 'WSB-30051', 
  'WSB-30052', 'WSB-30053', 'WSB-30054', 'WSB-30055')

broad_founder_all %>% 
  dplyr::mutate(Sample = factor(Sample, levels=sample_order),
                cytokine = as.numeric(.data[[cname]]),
                best_clean = factor(best_clean, levels = founder_order)) %>% 
  tidyr::replace_na(replace = list(cytokine = 0)) %>%
  dplyr::group_by(Sample) %>%
  dplyr::mutate(cytokine_sample_mean = mean(cytokine)) %>%
  dplyr::filter(!is.na(best_clean)) %>%
  ggplot(aes(Sample, cytokine_sample_mean, fill = best_clean)) +
    geom_col(position='dodge') +
    scale_fill_manual(values = founder_colors) +
    geom_point(aes(Sample, cytokine), color = "black") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Cytokine Profile by Founder (Broad)",
         x = "Founder",
         y = paste0(cname, " (pg/ml)"),
         fill = "Founder")

```

```{r}
ifng_founder_df <- ifng_founder_all %>%
  dplyr::filter(!is.na(best_clean)) %>%
  group_by(founder = best_clean) %>%
  summarize(IFNg = mean(IFNg))
```
