---
title: "Untitled"
format: html
editor: visual
---

## COMMENT:
In the revised manuscript, we will:
  (1) provide analysis of differentially expressed genes 
between these two populations; 
  (2) quantify ILC3 signature genes between Rorc high and 
Rorc low populations, to show the appropriate signature in both; and 
  (3) check the existence of the population of NKp46-CCR6-DN ILC3s in our dataset.

## Set Paths

# Load necessary libraries
# If you don't have these installed, run:
# install.packages("Seurat")
# install.packages("dplyr")
# install.packages("ggplot2")
# BiocManager::install("GSVA") # For Gene Set Variation Analysis (optional, for more advanced scoring)

```{r setup}
library(Seurat)
library(dplyr)
library(ggplot2)
# library(GSVA) # Uncomment if you plan to use GSVA for gene set scoring

domice_dir <- "/workspace/fasi-domice/"
data_dir <- paste0(domice_dir, "data/")
results_dir <- paste0(domice_dir, "results/")
figure_dir <- paste0(domice_dir, "figures/")
```


## Config

```{r config}
source(paste0(domice_dir, "src/helpers.R"))
```


## Load Data

```{r load_data}
obj <- readRDS(paste0(data_dir, "allchannels/ilc-seurat-object.rds"))
```

```{r}
# Set the identity of cells based on Rorc_status for DEG analysis
Idents(your_seurat_object) <- "Rorc_status"
```




```{r}
print("Performing differential expression analysis between Rorc_high and Rorc_low populations...")
deg_results <- FindMarkers(
  your_seurat_object,
  ident.1 = "Rorc_high",
  ident.2 = "Rorc_low",
  logfc.threshold = 0.25, # Minimum log2 fold-change
  min.pct = 0.1,          # Minimum percentage of cells expressing the gene
  test.use = "wilcox"     # Statistical test to use
)
```




```{r}
# 1.3 View and save DEG results
print("Differential Expression Results (top 10):")
print(head(deg_results, 10))
```


```{r}
# You can filter for significant genes (e.g., adjusted p-value < 0.05)
significant_degs <- deg_results %>%
  filter(p_val_adj < 0.05)

print(paste("Number of significant DEGs (adj. p-value < 0.05):", nrow(significant_degs)))

# Save the full DEG results to a CSV file
write.csv(deg_results, "DEGs_Rorc_high_vs_Rorc_low.csv", row.names = TRUE)
print("Full DEG results saved to 'DEGs_Rorc_high_vs_Rorc_low.csv'")

```


```{r}
# To create a volcano plot, you'll need the EnhancedVolcano package or custom ggplot2 code.
# install.packages("EnhancedVolcano")
# library(EnhancedVolcano)
# EnhancedVolcano(deg_results,
#                 lab = rownames(deg_results),
#                 x = 'avg_log2FC',
#                 y = 'p_val_adj',
#                 title = 'Rorc_high vs Rorc_low DEGs',
#                 pCutoff = 0.05,
#                 FCcutoff = 0.25)


# --- Reviewer Question 2: Quantify ILC3 signature genes ---
# Goal: Calculate and compare an ILC3 gene signature score between Rorc high and Rorc low populations.

# 2.1 Define ILC3 signature genes
# Replace with your actual list of ILC3 signature genes.
# Ensure these gene names match the gene names in your Seurat object.
ILC3_signature_genes <- c("Rorc", "Il7r", "Cd127", "Il2ra", "Gata3", "Id2", "Tcf7", "Ahr", "Stat3", "Il1rl1")

# Filter the signature gene list to only include genes present in your dataset
ILC3_signature_genes_present <- ILC3_signature_genes[ILC3_signature_genes %in% rownames(your_seurat_object)]
if (length(ILC3_signature_genes_present) == 0) {
    stop("None of the specified ILC3 signature genes were found in your dataset. Please check gene names.")
} else if (length(ILC3_signature_genes_present) < length(ILC3_signature_genes)) {
    print(paste("Warning: Not all ILC3 signature genes were found in your dataset. Using only:",
                paste(ILC3_signature_genes_present, collapse = ", ")))
}
```




```{r}
# 2.2 Calculate gene signature score
# Seurat's AddModuleScore is a good way to do this. It calculates the average expression
# of features in a gene set, subtracted by the aggregated expression of control feature sets.
print("Calculating ILC3 gene signature score...")
your_seurat_object <- AddModuleScore(
    object = your_seurat_object,
    features = list(ILC3_signature_genes_present),
    ctrl = 100, # Number of control genes to randomly select
    name = 'ILC3_Signature' # Name of the new metadata column for the score
)

# The scores will be added to the metadata slot, e.g., 'ILC3_Signature1'
# Rename for clarity
obj$ILC3_Score <- abj$ILC3_Signature1


```


```{r}
# 2.3 Quantify and visualize the scores
# Use violin plots or box plots to compare scores between Rorc high and Rorc low populations.
print("Visualizing ILC3 signature scores...")
p_ilc3_score <- VlnPlot(obj,
                        features = "ILC3_Score",
                        group.by = "Rorc_status",
                        pt.size = 0.1, # Adjust point size for visualization
                        cols = c("lightblue", "lightcoral")) + # Customize colors
    ggtitle("ILC3 Signature Score by Rorc Status") +
    ylab("ILC3 Signature Score") +
    xlab("Rorc Status") +
    theme_minimal()

print(p_ilc3_score)
```



```{r}
# Save the plot
ggsave("ILC3_Signature_Score_ViolinPlot.png", plot = p_ilc3_score, width = 6, height = 5, dpi = 300)
print("ILC3 signature score violin plot saved to 'ILC3_Signature_Score_ViolinPlot.png'")

# Calculate mean and standard deviation of ILC3 scores for each group
ILC3_score_summary <- your_seurat_object@meta.data %>%
    group_by(Rorc_status) %>%
    summarise(
        Mean_ILC3_Score = mean(ILC3_Score),
        SD_ILC3_Score = sd(ILC3_Score),
        Median_ILC3_Score = median(ILC3_Score),
        IQR_ILC3_Score = IQR(ILC3_Score)
    )
print("\nSummary of ILC3 Signature Scores:")
print(ILC3_score_summary)

# Optional: Perform a statistical test (e.g., Wilcoxon) to compare scores
wilcox_test_ilc3 <- wilcox.test(ILC3_Score ~ Rorc_status, data = your_seurat_object@meta.data)
print("\nWilcoxon Rank Sum test for ILC3 Score difference:")
print(wilcox_test_ilc3)
```




```{r}
# --- Reviewer Question 3: Check existence of NKp46-CCR6-DN ILC3s ---
# Goal: Identify and quantify cells with low/negative expression of NKp46 and CCR6,
# while retaining ILC3 characteristics (e.g., Rorc high).

# 3.1 Define gene names for NKp46 and CCR6
# Replace with the actual gene names in your dataset (case-sensitive).
NKp46_gene <- "Ncr1" # Often referred to as Ncr1 in mouse, NKp46 in human
CCR6_gene <- "Ccr6"

# Check if these genes exist in your object
if (!all(c(NKp46_gene, CCR6_gene) %in% rownames(your_seurat_object))) {
    stop(paste("Genes", NKp46_gene, "or", CCR6_gene, "not found in your Seurat object. Please check gene names."))
}

# 3.2 Extract expression data for relevant genes
# Use normalized data ('data' slot) for thresholds or visualization.
ncr1_expression <- GetAssayData(your_seurat_object, slot = "data")[NKp46_gene, ]
ccr6_expression <- GetAssayData(your_seurat_object, slot = "data")[CCR6_gene, ]

# Add expression to metadata for easier plotting/subsetting
your_seurat_object$NKp46_expression <- ncr1_expression
your_seurat_object$CCR6_expression <- ccr6_expression

# 3.3 Define thresholds for "negative" expression
# This is crucial and depends on your data and biological context.
# A common approach is to consider "negative" as expression below a certain percentile,
# or below a minimal detectable level after normalization.
# For example, cells with expression values close to 0 or below a low quantile.
# Here, we'll use a simple threshold (e.g., < 0.1 for normalized data, adjust as needed).
# You might want to visualize the distribution of these genes first to pick a threshold:
# hist(your_seurat_object$NKp46_expression[your_seurat_object$Rorc_status == "Rorc_high"])
# hist(your_seurat_object$CCR6_expression[your_seurat_object$Rorc_status == "Rorc_high"])

threshold_ncr1_negative <- 0.1 # Example threshold
threshold_ccr6_negative <- 0.1 # Example threshold

# 3.4 Identify NKp46-CCR6-DN ILC3s
# These are cells that are Rorc_high AND have low NKp46 AND low CCR6 expression.
your_seurat_object$ILC3_subtype <- "Other ILC3" # Default category

your_seurat_object@meta.data <- your_seurat_object@meta.data %>%
    mutate(
        ILC3_subtype = case_when(
            Rorc_status == "Rorc_high" & NKp46_expression < threshold_ncr1_negative & CCR6_expression < threshold_ccr6_negative ~ "NKp46-CCR6-DN ILC3",
            TRUE ~ ILC3_subtype # Keep existing "Other ILC3" or other categories
        )
    )

# 3.5 Quantify the proportion of this population
dn_ilc3_counts <- table(your_seurat_object$ILC3_subtype)
print("\nCounts of ILC3 subtypes:")
print(dn_ilc3_counts)

proportion_dn_ilc3 <- dn_ilc3_counts["NKp46-CCR6-DN ILC3"] / sum(dn_ilc3_counts)
print(paste0("Proportion of NKp46-CCR6-DN ILC3s among all ILC3-related cells: ", round(proportion_dn_ilc3 * 100, 2), "%"))

# 3.6 Visualize the existence of this population on a UMAP/t-SNE
# Assuming you have a UMAP reduction already computed in your Seurat object.
# If not, run:
# your_seurat_object <- RunUMAP(your_seurat_object, dims = 1:30) # Adjust dims as needed

print("Visualizing NKp46-CCR6-DN ILC3 population on UMAP...")
p_dn_ilc3_umap <- DimPlot(your_seurat_object,
                          reduction = "umap",
                          group.by = "ILC3_subtype",
                          label = TRUE,
                          repel = TRUE,
                          cols = c("NKp46-CCR6-DN ILC3" = "darkgreen", "Other ILC3" = "grey")) +
    ggtitle("ILC3 Subpopulations on UMAP") +
    theme_minimal()

print(p_dn_ilc3_umap)

# Save the plot
ggsave("NKp46_CCR6_DN_ILC3_UMAP.png", plot = p_dn_ilc3_umap, width = 8, height = 7, dpi = 300)
print("UMAP plot showing NKp46-CCR6-DN ILC3 population saved to 'NKp46_CCR6_DN_ILC3_UMAP.png'")

# Optional: Feature plots to show expression of NKp46 and CCR6
p_ncr1_feature <- FeaturePlot(your_seurat_object, features = NKp46_gene, reduction = "umap") +
    ggtitle(paste0("Expression of ", NKp46_gene, " on UMAP"))
print(p_ncr1_feature)

p_ccr6_feature <- FeaturePlot(your_seurat_object, features = CCR6_gene, reduction = "umap") +
    ggtitle(paste0("Expression of ", CCR6_gene, " on UMAP"))
print(p_ccr6_feature)

# Combine and save feature plots (requires 'patchwork' package)
# install.packages("patchwork")
# library(patchwork)
# combined_feature_plots <- p_ncr1_feature + p_ccr6_feature
# ggsave("NKp46_CCR6_FeaturePlots.png", plot = combined_feature_plots, width = 12, height = 6, dpi = 300)


print("\n--- Script Finished ---")
print("Remember to review the generated CSV files and plots, and adjust thresholds or parameters as needed for your specific dataset.")

```



