---
title: "Untitled"
format: html
editor: visual
---

## Reviewer Comments

First, to clarify, in the analysis in Fig. 2 we showed DO mice. Each DO mouse has a different genetics and mixture as the mouse are outbred (not inbred strains) and each is its own "strain".

```{r setup}
library(qtl2)
library(data.table)
library(readr)
library(dplyr)
library(ggplot2)

domice_dir <- "/workspace/fasi-domice/"
data_dir <- paste0(domice_dir, "data/")
results_dir <- paste0(domice_dir, "results/")
figure_dir <- paste0(domice_dir, "figures/")
```

```{r alleleprobs}
load(paste0(data_dir, "genotype/alleleprobs.Rdata"))
```


## DO Mice Founder Allele Contributions

In the revised manuscript, we will provide a new analysis showing the percent of each founder's haplotype that contributes to each of the DO mice genetic makeup.

TODO: Edit the following code. Count/Average allele prevalence of each founder strain and plot per DO mouse.

Running `founder-comparison.R` as a background job


```{r read_allele_df}
allele_df <- read_csv("founder_allele_counts.csv")
```





## Founder Mice ILC3 Comparisons

Second, we have profiled by scRNA-seq the 8 founder strains, and in the revised manuscript, we will show the difference between ILC3s from NOD and C57BL/6 mice.

TODO: Look at the scRNAseq founder data and the ILC3 cell type calls to quantify this.


```{r}
obs_f1 <- read_csv(paste0(data_dir, "founders1-louvain_labels-labels-transfer/obs.csv"))
```

```{r}
obs_f1 %>% 
  filter(best_clean %in% c("C57BL_6NJ","NOD_ShiLtJ")) %>%
  group_by(best_clean, called_cell_types_new) %>%
  count() %>%
  group_by(best_clean) %>%
  mutate(proportion = n / sum(n)) %>%
  filter(called_cell_types_new %in% c(paste0("ILC",1:3),"ILC3(LTi-like)")) %>%
  tidyr::pivot_wider(id_cols = called_cell_types_new, 
                     names_from = best_clean,
                     values_from = c(n, proportion)) %>%
  gt()
  
```


```{r}
p1 <- obs_f1 %>% 
  filter(best_clean %in% c("C57BL_6NJ","NOD_ShiLtJ")) %>%
  group_by(best_clean, called_cell_types_new) %>%
  count() %>%
  group_by(best_clean) %>%
  mutate(proportion = n / sum(n)) %>%
  filter(called_cell_types_new %in% c(paste0("ILC",1:3),"ILC3(LTi-like)")) %>%
  ggplot(aes(x = called_cell_types_new, y = proportion, fill = best_clean)) +
    geom_bar(stat="identity", position="dodge") +
    labs(x = "Cell Type", 
         y = "Proportion", 
         fill = "Founder",
         title = "ILC Prevalence Comparison") +
    scale_fill_manual(values = c("C57BL_6NJ"="#555555","NOD_ShiLtJ"="#0072B2")) +
    theme_minimal()

p1

ggsave(filename = paste0(results_dir, "founders/NOD_B6_ILC_comparison.png"),
      plot = p1)
  
```


