---
title: "Untitled"
format: html
editor: visual
---

## Reviewer Comments

First, to clarify, in the analysis in Fig. 2 we showed DO mice. Each DO mouse has a different genetics and mixture as the mouse are outbred (not inbred strains) and each is its own “strain”.


```{r setup}
library(qtl2)
library(data.table)
library(readr)
library(dplyr)
library(ggplot2)

domice_dir <- "/workspace/fasi-domice/"
data_dir <- paste0(domice_dir, "data/")
results_dir <- paste0(domice_dir, "results/")
figure_dir <- paste0(domice_dir, "figures/")
```


```{r alleleprobs}
load(paste0(data_dir, "genotype/alleleprobs.Rdata"))
antibody <- read_csv(paste0(data_dir, "phenotype/pheno_batch4_processed.csv"))
cytokines <- readxl::read_excel(paste0(data_dir, "phenotype/cytokine production -steady state.xlsx"))

```

```{r alleleprobs_filtering}
aprob_rnames <- rownames(aprobs[[1]])

name_vec <- sapply(cytokines$mouse_id, function(x) grep(x, aprob_rnames, value=T))

name_map <- data.frame(full_id = sapply(name_vec, last)) %>% 
  tibble::rownames_to_column("mouse_id") %>% 
  dplyr::filter(!is.na(full_id))
```

## DO Mice Founder Allele Contributions

In the revised manuscript, we will provide a new analysis showing the percent of each founder's haplotype that contributes to each of the DO mice genetic makeup.


TODO: Edit the following code.  Count/Average allele prevalence of each founder strain and plot per DO mouse.

```{r mean_df}
aprob_chr10_markers <- intersect(dimnames(aprobs[["10"]])[[3]], marker_list)
out_df <- aprobs[["10"]][,,aprob_chr10_markers]

# Compute mean across all data frames
mean_df <- apply(out_df,c(1,2),mean) %>% as.data.frame() %>%
  tibble::rownames_to_column("full_id")
```



## Founder Mice ILC3 Comparisons

Second, we have profiled by scRNA-seq the 8 founder strains, and in the revised manuscript, we will show the difference between ILC3s from NOD and C57BL/6 mice.


TODO: Look at the scRNAseq founder data and the ILC3 cell type calls to quantify this.



